### 2017-06-12 methyl-valerate/methyl-valerate.tex
Add US NCM paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This work is licensed under the Creative Commons Attribution 4.0
% International License. To view a copy of this license, visit
% http://creativecommons.org/licenses/by/4.0/.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{ussci}

\usepackage{enumitem}
\setlist{noitemsep}

\usepackage{xmpincl}
\includexmp{CC_Attribution_4.0_International}

\hypersetup{
    pdftitle={Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine},%
    pdfauthor={Bryan W. Weber},%
}

\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\sisetup{separate-uncertainty=true}

\usepackage{graphicx}
\usepackage{pgf}
\usepgflibrary{fpu}

\usepackage[tableposition=top]{caption}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{microtype}

\newcommand*{\logten}{\mathop{\log_{10}}}

% Fix missing mu character from a reference
\usepackage{textcomp}
\DeclareUnicodeCharacter{2212}{\textendash}

\usepackage[capitalize]{cleveref}
%======================================================================
\addbibresource{methyl-valerate.bib}
%======================================================================
\newcommand\papertopic{Reaction Kinetics}
%======================================================================
\DeclareSIUnit\torr{torr}

\title{ Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }

\author[1*]{Bryan W.\ Weber}
\author[1]{Justin Bunnell}
\author[2]{Kamal Kumar}
\author[1]{Chih-Jen Sung}
\affil[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT, USA}
\affil[2]{Department of Mechanical Engineering, University of Idaho, Moscow, ID, USA}
\affil[*]{Corresponding Author: \email{bryan.weber@uconn.edu}}

\begin{document}
\maketitle

%====================================================================
\begin{abstract} % not to exceed 200 words
    Methyl valerate (\ce{C6H12O2}, methyl pentanoate) is a methyl ester and a
    relevant surrogate component for biodiesel. In this work, we present
    ignition delays of methyl valerate measured using a rapid compression
    machine at a range of engine-relevant temperature, pressure, and equivalence
    ratio conditions. The conditions we have studied include equivalence ratios
    from \numrange{0.25}{2.0}, temperatures between \SIlist{680;1050}{\K}, and
    pressures of \SIlist{15;30}{\bar}. The ignition delay data demonstrate a
    negative temperature coefficient region in the temperature range of
    \SIrange[range-phrase={--}]{720}{800}{\K} for both \(\phi=2.0\),
    \SI{15}{\bar} and \(\phi=1.0\), \SI{30}{\bar}, with two-stage ignition
    apparent over the narrower temperature ranges of
    \SIrange[range-phrase={--}]{720}{760}{\K} for the lower pressure and
    \SIrange[range-phrase={--}]{740}{760}{\K} at the higher pressure. In
    addition, the experimental ignition delay data are compared with simulations
    using an existing chemical kinetic model from the literature. The
    simulations with the literature model under-predict the data by factors
    between \numlist{2;10} over the entire range of the experimental data. To
    help determine the possible reasons for the discrepancy between simulations
    and experiments, a new chemical kinetic model is developed using the
    Reaction Mechanism Generator (RMG) software. The agreement between the
    experimental data and the RMG model is improved but still not satisfactory.
    Directions for future improvement of the methyl valerate model are
    discussed.
\end{abstract}

% (Provide 2-4 keywords describing your research. Only abbreviations firmly
% established in the field may be used. These keywords will be used for
% sessioning/indexing purposes.)
\begin{keyword}
    chemical kinetics\sep rapid compression machine\sep autoignition\sep methyl ester
\end{keyword}

\section{Introduction}\label{sec:introduction}

For transportation applications, biodiesel is an important constituent in
improving environmental friendliness of fuels, due to its renewability when
produced from sustainable agricultural crops and its ability to reduce emissions
relative to conventionally fueled engines~\autocite{Hoekman2012}. A recent
review paper summarizes the work on methyl esters relevant to biodiesel
combustion~\autocite{Coniglio2013}. Autoignition of methyl butanoate (MB,
\ce{C5H10O2}) has been well-studied in both shock tube and rapid compression
machine experiments, and readers are referred to the review of
\textcite{Coniglio2013} for further details. The prevalence of MB data in the
literature is largely due to the early identification of MB as a potential
surrogate fuel for biodiesel~\autocite{Fisher2000}. However, the experiments
have shown that MB may not be an appropriate surrogate for biodiesel, due to its
lack of negative temperature coefficient (NTC) behavior, a requirement for a
suitable biodiesel surrogate~\autocite{Coniglio2013}.

Larger methyl esters such as methyl valerate (MV, \ce{C6H12O2}, methyl
pentanoate) have also been studied as possible biodiesel surrogates.
\textcite{Hadj-Ali2009} used a rapid compression machine (RCM) to study the
autoignition of several methyl esters including MV. Although MV exhibited
two-stage ignition in this study, little additional research has been done on
its oxidation. \textcite{Korobeinichev2015} studied MV in premixed laminar
flames and extended a detailed high temperature chemical kinetic model to
include MV and methyl hexanoate. \textcite{Dmitriev2015} added MV to
n-heptane/toluene fuel blends to determine the resulting intermediate species in
premixed flames using a flat burner at \SI{1}{atm} and an equivalence ratio of
1.75. The addition of MV helped reduce soot forming intermediates including
benzene, cyclopentadienyl, acetylene, propargyl, and
vinylacetylene~\autocite{Dmitriev2015}. \textcite{Hayes2009} computationally
examined the peroxy radical isomerization reactions for MV to better understand
the low temperature reaction pathways. Finally, \textcite{Dievart2013} used
diffusion flames in the counterflow configuration to determine extinction limits
for a number of methyl esters, including MV, and validated a detailed kinetic
model with the experimental data.

This work provides additional data for the autoignition of MV. Data is collected
in a RCM under engine relevant conditions spanning
from \SIrange{15}{30}{\bar}, equivalence ratios from \numrange{0.25}{2.0}, and
temperatures from \SIrange{682}{1048}{\K}. The NTC region of MV is mapped out to
provide additional information on the fidelity of using MV as a biodiesel
surrogate.

\section{Experimental Methods}\label{sec:experimental-methods}

The RCM used in this study is a single piston arrangement and is pneumatically
driven and hydraulically stopped. The device has been described in detail
previously~\autocite{Mittal2007a} and will be described here briefly for
reference. The end of compression (EOC) temperature and pressure (\(T_C\) and
\(P_C\) respectively), are independently changed by varying the overall
compression ratio, initial pressure \((P_0)\), and initial temperature \((T_0)\)
of the experiments. The primary diagnostic on the RCM is the in-cylinder
pressure. The pressure data is processed by a Python package called UConnRCMPy~\autocite{uconnrcmpy}, which calculates \(P_C\), \(T_C\), and the ignition
delay(s). The definitions of the ignition delays are shown in
\cref{fig:ign-delay-def}. The time of the EOC is defined as the maximum of the
pressure trace prior to the start of ignition and the ignition delays are
defined as the time from the EOC until local maxima in the first time derivative
of the pressure. Each experimental condition is repeated at least five times to
ensure repeatability of the data. As there is some random scatter present in the
data, the standard deviation \((\sigma)\) of the ignition delays from the runs
at a given condition is computed. In all cases, \(\sigma\) is less than
\SI{10}{\percent} of the mean value of the overall ignition delay.

In addition to the reactive experiments, non-reactive experiments are conducted
to determine the influence of machine-specific behavior on the experimental
conditions and permit the calculation of the EOC temperature via the isentropic
relations between pressure and temperature~\autocite{Lee1998}. The EOC
temperature is calculated by the procedure described in
\cref{sec:computational-methods}.

\begin{figure}[htb]
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \resizebox{\linewidth}{!}{\input{figures/ignition_delay_definition.pgf}}
        \caption{Definition of the ignition delays used in this work. The
        experiment in this figure was conducted for a \(\phi=2.0\) mixture with
        \mbox{\(\ce{Ar}/(\ce{N2}+\ce{AR})=0.5\)}, \(P_0=\SI{0.7694}{\bar}\),
        \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.94}{\bar}\), \(T_C=\SI{723}{\K}\),
        \(\tau=\SI{27.44\pm0.99}{\ms}\), \(\tau_1=\SI{16.57\pm0.48}{\ms}\).}
        \label{fig:ign-delay-def}
    \end{minipage}\hfill%
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \resizebox{\linewidth}{!}{\input{figures/vapor_pressure.pgf}}
        \caption{Saturated vapor pressure of MV as a function of temperature,
        plotted using the Antoine equation, \cref{eq:antoine}, with
        \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
        \label{fig:vapor-pressure}
    \end{minipage}
\end{figure}

The RCM is equipped with heaters to control the initial temperature of the
mixture. After filling in the components to the mixing tanks, the heaters are
switched on and the system is allowed \SI{1.5}{\hour} to come to steady state.
The mixing tanks are also equipped with magnetic stir bars so the reactants are
well mixed for the duration of the experiments.

The initial temperature is chosen such that the saturated vapor pressure
\((P_{\text{sat}})\) of the fuel at the initial temperature is at least twice
the partial pressure of the fuel in the mixing tank. The Antoine equation
%
\begin{equation}\label{eq:antoine}
    \logten{P_{\text{sat}}} = A - \frac{B}{T - C}
\end{equation}
%
is used to model the saturated vapor pressure of MV as a function of
temperature, where \(A\), \(B\), and \(C\) are substance-specific coefficients.
The coefficients in \cref{eq:antoine} are determined by least squares fitting of
the data of \textcite{Ortega2003}, \textcite{vanGenderen2002}, and
\textcite{Verevkin2008} using the \verb|curve_fit()| function of
SciPy~\autocite{Jones2001} version 0.18.1. \Cref{fig:vapor-pressure} shows that
the coefficients fit with this procedure give good agreement with the
experimental data.

The mixtures considered in this study are shown in \cref{tab:mixtures}. Mixtures
are prepared in stainless steel mixing tanks. The proportions of reactants in
the mixture are determined by specifying the absolute mass of the fuel, the
equivalence ratio \((\phi)\), and the ratio of \(\ce{Ar}:\ce{N2}\) in the
oxidizer. Since MV is a liquid with a relatively small vapor pressure at room
temperature, it is injected into the mixing tank through a septum. Proportions
of \ce{O2}, \ce{Ar}, and \ce{N2} are added manometrically at room temperature.

\begin{table}[htb]
    \centering
    \caption{Mixtures considered in this work}
    \begin{tabular}{Sccccc}
        \toprule
        {\(\phi\)} & \multicolumn{4}{c}{Mole Fraction (purity)} & \ce{Ar}/(\ce{N2} + \ce{Ar}) \\
        \cmidrule{2-5}
         & \ce{MV} (\SI{100}{\percent}) & \ce{O2} (\SI{99.994}{\percent}) & \ce{Ar} (\SI{99.999}{\percent}) & \ce{N2} (\SI{99.999}{\percent}) &  \\
        \midrule
        0.25 & 0.0065 & 0.2087 & 0.7848 & 0.0000 & 1.0 \\
        0.5 & 0.0130 & 0.2074 & 0.7798 & 0.0000 & 1.0 \\
        1.0 & 0.0256 & 0.2047 & 0.7697 & 0.0000 & 1.0 \\
        1.0 & 0.0256 & 0.2047 & 0.3849 & 0.3848 & 0.5 \\
        2.0 & 0.0499 & 0.1996 & 0.0000 & 0.7505 & 0.0 \\
        2.0 & 0.0499 & 0.1996 & 0.3752 & 0.3753 & 0.5 \\
        \bottomrule
    \end{tabular}
    \label{tab:mixtures}
\end{table}

% \begin{table}
%     \centering
%     \caption{Antoine Equation coefficients computed in this work and from the
%     literature. The $2\sigma$ confidence interval is estimated by taking the
%     square root of the diagonals of the covariance matrix returned from
%     \texttt{curve\_fit()}}
%     \label{tab:antoine}
%     \begin{tabular}{cccccc}
%         \toprule
%         & $A$ & $B$ & $C$ & $T_{\text{min}}$, \si{\K} & $T_{\text{max}}$, \si{\K} \\
%         \midrule
%         This Work & 6.4030 & 1528.69 & 52.881 & 274.9 & 417.18 \\
%         $2\sigma$ Confidence Interval & 0.0919 & 53.47 & 4.934 & --- & --- \\
%         \textcite{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
%         \textcite{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
%         \textcite{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
%         \bottomrule
%     \end{tabular}
% \end{table}

\section{Computational Methods}\label{sec:computational-methods}
\subsection{RCM Modeling}\label{sec:experimental-modeling}

The Python 3.5 interface of Cantera~\autocite{cantera} version 2.2.1 is used for
all simulations in this work. Detailed descriptions of the use of Cantera for
these simulations can be found in the work of \textcite{Weber2016a} and
\textcite{Dames2016}; a brief overview is given here. As mentioned in
\cref{sec:experimental-methods}, non-reactive experiments are conducted to
characterize the machine-specific effects on the experimental conditions in the
RCM. This pressure trace is used to compute a volume trace by assuming that the
reactants undergo a reversible, adiabatic, constant composition (i.e.,
isentropic) compression during the compression stroke and an isentropic
expansion after the EOC. The volume trace is applied to a simulation conducted
in an \verb|IdealGasReactor| in Cantera~\autocite{cantera} using the CVODES
solver from the SUNDIALS suite~\autocite{Hindmarsh2005}. The ignition delay from
the simulations is defined in the same manner as in the experiments. The time
derivative of the pressure in the simulations is computed by second order
Lagrange polynomials, as discussed by \textcite{Chapra2010}.

To the best of our knowledge, there are three mechanisms for MV combustion
available in the literature. The first two, by \textcite{Korobeinichev2015} and
\textcite{Dmitriev2015}, were developed to simulate flames, and do not include
the low-temperature chemistry necessary to simulate the conditions in these
experiments. The third model was developed by \textcite{Dievart2013} and
includes low-temperature chemistry of MV, although it was only validated by
comparison with flame extinction limits.
% To use this mechanism in Cantera, we
% made a number of modifications, including removing duplicate species from the
% thermodynamic database and removing the excited \ce{OH*} and \ce{CH*} chemistry,
% which is not supported by Cantera 2.2.1. For most of the duplicate species in
% the thermodynamic database, the thermodynamic data is identical, so one copy of
% the duplicate species is simply removed from the database. However, some species
% are not exact duplicates. For these species, it is not clear from the
% thermodynamic database file which data set should be preferred and the default
% of Cantera (and CHEMKIN) to choose the first instance of a duplicate species is
% used.
The detailed \textcite{Dievart2013} model includes 1103 species and 7557
reactions.

\subsection{Reaction Mechanism Generator}\label{sec:reaction-mechanism-generator}

In addition to using a mechanism from the literature, we investigate the use of
an automatic mechanism generator, the open-source Reaction Mechanism Generator
(RMG)~\autocite{Allen2012} version 1.0.4. The Python version of RMG is used,
which requires Python 2.7, and version 1.10.0 of the RMG database is used. The
final RMG model contains 483 species and 19990 reactions. Note that the number
of species is much lower than the \textcite{Dievart2013} model because the RMG
model focuses on only one fuel (MV), but the number of reactions is
substantially higher.

\section{Experimental Results}\label{sec:experimental-results}

\Cref{fig:ignition-delays} shows the ignition delay results measured in this
study. Filled markers denote the overall ignition delay and hollow markers
indicate the first-stage ignition delay. Vertical error bars are drawn on the
symbols to represent the uncertainty in the ignition delay; for many of the
experiments, the uncertainty is approximately the same size as the data point,
so the error bar is hidden. Horizontal error bars are shown on the first and
last points of each equivalence ratio indicating the estimated uncertainty in
the EOC temperature of $\pm\SI{1}{\percent}$~\autocite{Weber2015}.
\cref{fig:ignition-delays}a shows the results for a compressed pressure of
\SI{15}{\bar}, while \cref{fig:ignition-delays}b shows the results for a
compressed pressure of \SI{30}{\bar}. Note that $\phi=2.0$ results were not
collected for \SI{30}{\bar}, so there are no red data points in
\cref{fig:ignition-delays}b.

It can be seen from \cref{fig:ignition-delays} that the ignition delays for the
\(\phi=0.25\text{ and } 0.5\) mixtures do not show an NTC region of the ignition
delay for both of the pressures studied in this work. However, the $\phi=1.0$
mixture shows an NTC region at $P_C=\SI[number-unit-product={\ }]{30}{\bar}$
between approximately \SIlist{720;800}{\K}, with measured first-stage ignition
delays at \SIlist{734;757}{\K}. In addition, the $\phi=2.0$ mixture shows an NTC
region of ignition delay at \SI{15}{\bar} from approximately
\SIrange{720}{775}{\K}, with measured first-stage ignition delays between
\SIlist{720;750}{\K}.

\begin{figure}[htb]
    \centering
    \input{figures/ignition-delays.pgf}
    \caption{Ignition delays of MV as a function of inverse temperature. Filled
    points are the overall ignition delays and hollow points are the first stage
    ignition delays. a) \SI{15}{\bar}. b) \SI{30}{\bar}}
    \label{fig:ignition-delays}
\end{figure}

\Cref{fig:pressure-traces} shows the pressure traces for selected experiments at
\(\phi=1.0\), \(P_C =\SI[number-unit-product={\ }]{30}{\bar}\). The three
reactive pressure traces shown are at the low-temperature end of the NTC (blue,
\SI{700}{\K}), one case with two-stage ignition (orange, \SI{734}{\K}), and one
case near the high-temperature limit of the NTC region (green, \SI{775}{\K}).
Also shown is the non-reactive pressure trace for the \SI{700}{\K} case (red).
By comparing the \SI{700}{\K} pressure trace with the non-reactive pressure
trace, it can be seen that there is substantial heat release prior to main
ignition as measured by the deviation of the reactive pressure trace from the
non-reactive trace. However, there is only one peak in the time derivative of
the pressure, so no first-stage ignition delay is defined for this case. It can
also be seen in \cref{fig:pressure-traces} that the \SI{775}{\K} case shows some
heat release prior to ignition, although again there is only one peak in the
time derivative of the pressure. Furthermore, the heat release at \SI{775}{\K}
appears to be more gradual than at the lowest temperature.

\begin{figure}[htb]
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \resizebox{\linewidth}{!}{\input{figures/pressure-traces.pgf}}
        \caption{Selected pressure traces around the NTC region of ignition delay for $\phi=1.0$}
        \label{fig:pressure-traces}
    \end{minipage}\hfill%
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \resizebox{\linewidth}{!}{\input{figures/simulation-comparison.pgf}}
        \caption{Comparison of experimental and simulated results for $\phi=1.0$}
        \label{fig:simulation-comparison}
    \end{minipage}
\end{figure}

\section{Computational Results}\label{sec:computational-results}

\Cref{fig:simulation-comparison} compares experimentally measured overall
ignition delays with ignition delays computed with the detailed model of
\textcite{Dievart2013} for the $\phi=1.0$ experiments. Results for the other
equivalence ratios are similar to these results, so are not shown here. It is
important to note that the model of \textcite{Dievart2013} was not validated for
MV ignition delays, only for extinction strain rates. At \SI{15}{\bar}, the
model tends to under-predict the ignition delay and predicts an NTC region that
is not present in the experiments. At \SI{30}{\bar}, the model predicts the
low-temperature ignition delays well, but does not predict the NTC region found
experimentally.

To understand the underlying reasons for the disagreement between the
\textcite{Dievart2013} model and the data, we constructed an additional model
using RMG (see \cref{sec:reaction-mechanism-generator}). As can be seen in
\cref{fig:simulation-comparison}, the agreement between the RMG model and the
experimental data is similar to the \textcite{Dievart2013} model for the
\SI{30}{\bar} data. At \SI{15}{\bar}, the RMG model predicts a somewhat longer
ignition delay than the model of \textcite{Dievart2013}, but still predicts an
NTC region where none is present in the experimental data.

In general, there could be three likely sources of error in the models: missing
reaction pathways, incorrect values of the reaction rates, and incorrect values
for thermodynamic properties of the species. We have noted in
\cref{sec:reaction-mechanism-generator} that the RMG model has many more
reactions than the \textcite{Dievart2013} model and the algorithm used in RMG
considers a substantial number of the possible pathways. This reduces the
possibility of missing reaction pathways affecting the model. Further detailed
studies are required to ensure that the RMG model includes all of the relevant
reaction pathways.

The second source of error may be incorrect reaction rate parameters, either
because the rates are specified incorrectly in the model (e.g., typos) or
because the rates are not well estimated by the typical analogy based-rules. It
should be noted that errors of this type may affect the model generated by
RMG---if the rates are not estimated correctly, reactions that are important in
reality may not be included in the model. Determining the accuracy of the
reaction rates used in the RMG and \textcite{Dievart2013} models requires
further detailed studies of the models. Another related source of error could be
incorrect estimation of the pressure dependence of the reaction rates, which may
be particularly important for the isomerization reactions prevalent in
low-temperature chemistry.

The third source of error may lie in the estimation of the thermodynamic
properties of the species, particularly their heats of formation. We have begun
to analyze the possibility of this source of error by conducting a reaction
pathway analysis to determine which radicals are formed from the breakdown of
the fuel. The following analysis is conducted for a constant volume simulation
at \SI{700}{\K}, \SI{30}{\bar}, where the rates of production of the species
have been integrated until the time of \SI{20}{\percent} fuel consumption. The
results of this analysis are shown in \cref{fig:mv-structure} and
\cref{tab:mv-radicals} for the two models. The percentages shown in the
\cref{tab:mv-radicals} are the percent of the fuel destroyed to form a
particular fuel radical by all the reactions that can form that radical.

At the relatively low temperature and high pressure condition of this analysis,
all of the fuel is destroyed by \ce{H}-atom abstractions to form the fuel
radicals shown. It can be seen that the two models have quite different
distributions of products from the first \ce{H}-abstraction reactions. The model
of \textcite{Dievart2013} predicts that \ce{H}-abstraction from the second
carbon is the most prevalent, followed closely by abstraction from the methyl
group. This is in line with the bond energies of the \ce{C-H} bonds for those
carbon atoms; we expect that the presence of the oxygen atoms will cause
hydrogen abstraction at the nearby carbons to be favored. However, the RMG model
predicts that radicals in the middle of the carbon chain will be primarily
formed. The cause of this discrepancy is under investigation, but it may be
caused by the estimation of thermodynamic properties of the radicals.

\begin{figure}
    \begin{minipage}[t][][b]{0.3\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/Methyl_pentanoate.png}
        \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:mv-radicals}}
        \label{fig:mv-structure}
    \end{minipage}\hfill%
    \begin{minipage}[t][][b]{0.68\textwidth}
        \centering
        \captionof{table}{Percent of MV destroyed to form fuel radical species with a hydrogen atom missing at the location indicated in the first column}
        \label{tab:mv-radicals}
        \begin{tabular}{cSS}
            \toprule
            Radical Site & {\textcite{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} \\
            \midrule
            2 & 29.3 & 7.4 \\
            3 & 17.5 & 36.0 \\
            4 & 17.5 & 41.1 \\
            5 & 9.4 & 3.7 \\
            M & 26.3 & 11.8 \\
            \bottomrule
        \end{tabular}
    \end{minipage}
\end{figure}

\section{Conclusions}\label{sec:conclusions}

In this study, we have measured ignition delays for methyl valerate over a wide
range of engine-relevant pressures, temperatures, and equivalence ratios. An NTC
region of the ignition delay and two-stage ignition were recorded for pressures
of \SI{15}{\bar} at \(\phi=2.0\) and \SI{30}{\bar} at \(\phi=1.0\). A detailed
chemical kinetic model available in the literature was unable to reproduce the
experimental results, so a new model was constructed using the Reaction
Mechanism Generator software. Although the new model contains many more
reactions than the literature model, it is still unable to predict the
experimental ignition delays satisfactorily. Possible reasons for the
discrepancy include missing reaction pathways, incorrect rate estimates, and
incorrect thermodynamic property estimates. Future work will include
investigation of the discrepancies between models and experiments to further
understand the autoignition kinetics of methyl valerate.

\printbibliography

\end{document}

### 2017-06-14 methyl-valerate.tex
Move files out of subdirectory

--- 2017-06-12 methyl-valerate/methyl-valerate.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This work is licensed under the Creative Commons Attribution 4.0
-% International License. To view a copy of this license, visit
-% http://creativecommons.org/licenses/by/4.0/.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[12pt]{ussci}
+\documentclass[letterpaper,review]{elsarticle}
+
+\journal{Fuel}
+
+\usepackage{lineno}
 
@@
 \setlist{noitemsep}
-
-\usepackage{xmpincl}
-\includexmp{CC_Attribution_4.0_International}
-
-\hypersetup{
-    pdftitle={Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine},%
-    pdfauthor={Bryan W. Weber},%
-}
 
@@
 \usepackage{mathtools}
-\usepackage{microtype}
 
@@
 % Fix missing mu character from a reference
-\usepackage{textcomp}
-\DeclareUnicodeCharacter{2212}{\textendash}
+% \usepackage{textcomp}
+% \DeclareUnicodeCharacter{2212}{\textendash}
 
 \usepackage[capitalize]{cleveref}
-%======================================================================
-\addbibresource{methyl-valerate.bib}
-%======================================================================
-\newcommand\papertopic{Reaction Kinetics}
-%======================================================================
 \DeclareSIUnit\torr{torr}
 
+\bibliographystyle{elsarticle-num}
+
+\begin{document}
+%======================================================================
+\begin{frontmatter}
 \title{ Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }
 
-\author[1*]{Bryan W.\ Weber}
-\author[1]{Justin Bunnell}
-\author[2]{Kamal Kumar}
-\author[1]{Chih-Jen Sung}
-\affil[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT, USA}
-\affil[2]{Department of Mechanical Engineering, University of Idaho, Moscow, ID, USA}
-\affil[*]{Corresponding Author: \email{bryan.weber@uconn.edu}}
-
-\begin{document}
-\maketitle
+\author[uconn]{Bryan W.\ Weber\corref{cor1}}
+\author[uconn]{Justin Bunnell}
+\author[idaho]{Kamal Kumar}
+\author[uconn]{Chih-Jen Sung}
+\address[uconn]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT, USA}
+\address[idaho]{Department of Mechanical Engineering, University of Idaho, Moscow, ID, USA}
+\cortext[cor1]{Corresponding Author: bryan.weber@uconn.edu}
 
@@
 \end{keyword}
+\end{frontmatter}
+%======================================================================
+
+\linenumbers
 
@@
 produced from sustainable agricultural crops and its ability to reduce emissions
-relative to conventionally fueled engines~\autocite{Hoekman2012}. A recent
+relative to conventionally fueled engines~\cite{Hoekman2012}. A recent
 review paper summarizes the work on methyl esters relevant to biodiesel
-combustion~\autocite{Coniglio2013}. Autoignition of methyl butanoate (MB,
+combustion~\cite{Coniglio2013}. Autoignition of methyl butanoate (MB,
 \ce{C5H10O2}) has been well-studied in both shock tube and rapid compression
 machine experiments, and readers are referred to the review of
-\textcite{Coniglio2013} for further details. The prevalence of MB data in the
+\cite{Coniglio2013} for further details. The prevalence of MB data in the
 literature is largely due to the early identification of MB as a potential
-surrogate fuel for biodiesel~\autocite{Fisher2000}. However, the experiments
+surrogate fuel for biodiesel~\cite{Fisher2000}. However, the experiments
 have shown that MB may not be an appropriate surrogate for biodiesel, due to its
 lack of negative temperature coefficient (NTC) behavior, a requirement for a
-suitable biodiesel surrogate~\autocite{Coniglio2013}.
+suitable biodiesel surrogate~\cite{Coniglio2013}.
 
@@
 pentanoate) have also been studied as possible biodiesel surrogates.
-\textcite{Hadj-Ali2009} used a rapid compression machine (RCM) to study the
+\cite{Hadj-Ali2009} used a rapid compression machine (RCM) to study the
 autoignition of several methyl esters including MV. Although MV exhibited
 two-stage ignition in this study, little additional research has been done on
-its oxidation. \textcite{Korobeinichev2015} studied MV in premixed laminar
+its oxidation. \cite{Korobeinichev2015} studied MV in premixed laminar
 flames and extended a detailed high temperature chemical kinetic model to
-include MV and methyl hexanoate. \textcite{Dmitriev2015} added MV to
+include MV and methyl hexanoate. \cite{Dmitriev2015} added MV to
 n-heptane/toluene fuel blends to determine the resulting intermediate species in
@@
 benzene, cyclopentadienyl, acetylene, propargyl, and
-vinylacetylene~\autocite{Dmitriev2015}. \textcite{Hayes2009} computationally
+vinylacetylene~\cite{Dmitriev2015}. \cite{Hayes2009} computationally
 examined the peroxy radical isomerization reactions for MV to better understand
-the low temperature reaction pathways. Finally, \textcite{Dievart2013} used
+the low temperature reaction pathways. Finally, \cite{Dievart2013} used
 diffusion flames in the counterflow configuration to determine extinction limits
@@
 driven and hydraulically stopped. The device has been described in detail
-previously~\autocite{Mittal2007a} and will be described here briefly for
+previously~\cite{Mittal2007a} and will be described here briefly for
 reference. The end of compression (EOC) temperature and pressure (\(T_C\) and
@@
 of the experiments. The primary diagnostic on the RCM is the in-cylinder
-pressure. The pressure data is processed by a Python package called UConnRCMPy~\autocite{uconnrcmpy}, which calculates \(P_C\), \(T_C\), and the ignition
+pressure. The pressure data is processed by a Python package called UConnRCMPy~\cite{uconnrcmpy}, which calculates \(P_C\), \(T_C\), and the ignition
 delay(s). The definitions of the ignition delays are shown in
@@
 conditions and permit the calculation of the EOC temperature via the isentropic
-relations between pressure and temperature~\autocite{Lee1998}. The EOC
+relations between pressure and temperature~\cite{Lee1998}. The EOC
 temperature is calculated by the procedure described in
@@
 The coefficients in \cref{eq:antoine} are determined by least squares fitting of
-the data of \textcite{Ortega2003}, \textcite{vanGenderen2002}, and
-\textcite{Verevkin2008} using the \verb|curve_fit()| function of
-SciPy~\autocite{Jones2001} version 0.18.1. \Cref{fig:vapor-pressure} shows that
+the data of \cite{Ortega2003}, \cite{vanGenderen2002}, and
+\cite{Verevkin2008} using the \verb|curve_fit()| function of
+SciPy~\cite{Jones2001} version 0.18.1. \Cref{fig:vapor-pressure} shows that
 the coefficients fit with this procedure give good agreement with the
@@
 %         $2\sigma$ Confidence Interval & 0.0919 & 53.47 & 4.934 & --- & --- \\
-%         \textcite{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
-%         \textcite{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
-%         \textcite{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
+%         \cite{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
+%         \cite{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
+%         \cite{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
 %         \bottomrule
@@
 
-The Python 3.5 interface of Cantera~\autocite{cantera} version 2.2.1 is used for
+The Python 3.5 interface of Cantera~\cite{cantera} version 2.2.1 is used for
 all simulations in this work. Detailed descriptions of the use of Cantera for
-these simulations can be found in the work of \textcite{Weber2016a} and
-\textcite{Dames2016}; a brief overview is given here. As mentioned in
+these simulations can be found in the work of \cite{Weber2016a} and
+\cite{Dames2016}; a brief overview is given here. As mentioned in
 \cref{sec:experimental-methods}, non-reactive experiments are conducted to
@@
 expansion after the EOC. The volume trace is applied to a simulation conducted
-in an \verb|IdealGasReactor| in Cantera~\autocite{cantera} using the CVODES
-solver from the SUNDIALS suite~\autocite{Hindmarsh2005}. The ignition delay from
+in an \verb|IdealGasReactor| in Cantera~\cite{cantera} using the CVODES
+solver from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delay from
 the simulations is defined in the same manner as in the experiments. The time
 derivative of the pressure in the simulations is computed by second order
-Lagrange polynomials, as discussed by \textcite{Chapra2010}.
+Lagrange polynomials, as discussed by \cite{Chapra2010}.
 
 To the best of our knowledge, there are three mechanisms for MV combustion
-available in the literature. The first two, by \textcite{Korobeinichev2015} and
-\textcite{Dmitriev2015}, were developed to simulate flames, and do not include
+available in the literature. The first two, by \cite{Korobeinichev2015} and
+\cite{Dmitriev2015}, were developed to simulate flames, and do not include
 the low-temperature chemistry necessary to simulate the conditions in these
-experiments. The third model was developed by \textcite{Dievart2013} and
+experiments. The third model was developed by \cite{Dievart2013} and
 includes low-temperature chemistry of MV, although it was only validated by
@@
 % used.
-The detailed \textcite{Dievart2013} model includes 1103 species and 7557
+The detailed \cite{Dievart2013} model includes 1103 species and 7557
 reactions.
@@
 an automatic mechanism generator, the open-source Reaction Mechanism Generator
-(RMG)~\autocite{Allen2012} version 1.0.4. The Python version of RMG is used,
+(RMG)~\cite{Allen2012} version 1.0.4. The Python version of RMG is used,
 which requires Python 2.7, and version 1.10.0 of the RMG database is used. The
 final RMG model contains 483 species and 19990 reactions. Note that the number
-of species is much lower than the \textcite{Dievart2013} model because the RMG
+of species is much lower than the \cite{Dievart2013} model because the RMG
 model focuses on only one fuel (MV), but the number of reactions is
@@
 last points of each equivalence ratio indicating the estimated uncertainty in
-the EOC temperature of $\pm\SI{1}{\percent}$~\autocite{Weber2015}.
+the EOC temperature of $\pm\SI{1}{\percent}$~\cite{Weber2015}.
 \cref{fig:ignition-delays}a shows the results for a compressed pressure of
@@
 ignition delays with ignition delays computed with the detailed model of
-\textcite{Dievart2013} for the $\phi=1.0$ experiments. Results for the other
+\cite{Dievart2013} for the $\phi=1.0$ experiments. Results for the other
 equivalence ratios are similar to these results, so are not shown here. It is
-important to note that the model of \textcite{Dievart2013} was not validated for
+important to note that the model of \cite{Dievart2013} was not validated for
 MV ignition delays, only for extinction strain rates. At \SI{15}{\bar}, the
@@
 To understand the underlying reasons for the disagreement between the
-\textcite{Dievart2013} model and the data, we constructed an additional model
+\cite{Dievart2013} model and the data, we constructed an additional model
 using RMG (see \cref{sec:reaction-mechanism-generator}). As can be seen in
 \cref{fig:simulation-comparison}, the agreement between the RMG model and the
-experimental data is similar to the \textcite{Dievart2013} model for the
+experimental data is similar to the \cite{Dievart2013} model for the
 \SI{30}{\bar} data. At \SI{15}{\bar}, the RMG model predicts a somewhat longer
-ignition delay than the model of \textcite{Dievart2013}, but still predicts an
+ignition delay than the model of \cite{Dievart2013}, but still predicts an
 NTC region where none is present in the experimental data.
@@
 \cref{sec:reaction-mechanism-generator} that the RMG model has many more
-reactions than the \textcite{Dievart2013} model and the algorithm used in RMG
+reactions than the \cite{Dievart2013} model and the algorithm used in RMG
 considers a substantial number of the possible pathways. This reduces the
@@
 reality may not be included in the model. Determining the accuracy of the
-reaction rates used in the RMG and \textcite{Dievart2013} models requires
+reaction rates used in the RMG and \cite{Dievart2013} models requires
 further detailed studies of the models. Another related source of error could be
@@
 distributions of products from the first \ce{H}-abstraction reactions. The model
-of \textcite{Dievart2013} predicts that \ce{H}-abstraction from the second
+of \cite{Dievart2013} predicts that \ce{H}-abstraction from the second
 carbon is the most prevalent, followed closely by abstraction from the methyl
@@
             \toprule
-            Radical Site & {\textcite{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} \\
+            Radical Site & {\cite{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} \\
             \midrule
@@
 
-\printbibliography
+\section*{References}
+\bibliography{methyl-valerate}
 

### 2017-06-15 methyl-valerate.tex
Add els-CNF bib style
Rework the first few sections

@@
 \journal{Fuel}
+
+\usepackage{todonotes}
 
@@
 
-% Fix missing mu character from a reference
-% \usepackage{textcomp}
-% \DeclareUnicodeCharacter{2212}{\textendash}
+\DeclareSIUnit\torr{torr}
+\DeclareSIUnit\atm{atm}
 
+\bibliographystyle{elsarticle-num-CNF}
+\usepackage{doi}
 \usepackage[capitalize]{cleveref}
-\DeclareSIUnit\torr{torr}
-
-\bibliographystyle{elsarticle-num}
-
 \begin{document}
@@
 
-For transportation applications, biodiesel is an important constituent in
-improving environmental friendliness of fuels, due to its renewability when
-produced from sustainable agricultural crops and its ability to reduce emissions
-relative to conventionally fueled engines~\cite{Hoekman2012}. A recent
-review paper summarizes the work on methyl esters relevant to biodiesel
-combustion~\cite{Coniglio2013}. Autoignition of methyl butanoate (MB,
-\ce{C5H10O2}) has been well-studied in both shock tube and rapid compression
-machine experiments, and readers are referred to the review of
-\cite{Coniglio2013} for further details. The prevalence of MB data in the
-literature is largely due to the early identification of MB as a potential
-surrogate fuel for biodiesel~\cite{Fisher2000}. However, the experiments
-have shown that MB may not be an appropriate surrogate for biodiesel, due to its
-lack of negative temperature coefficient (NTC) behavior, a requirement for a
-suitable biodiesel surrogate~\cite{Coniglio2013}.
+For transportation applications, biodiesel is an important constituent in improving environmental
+friendliness of fuels. This is due to its renewability when produced from sustainable agricultural
+crops and its ability to reduce emissions relative to conventionally fueled
+engines~\cite{Hoekman2012}. Biodiesel typically consists of long-chain methyl ester molecules, with
+typical compositions of \ce{C14} to \ce{C20}~\cite{Hoekman2012}. Recognizing that the large
+molecular size of the methyl esters within biodiesel fuel makes creating and using detailed chemical
+kinetic mechanisms challenging~\cite{Lai2011}, it is desired to study their combustion chemistry by
+studying simpler molecules.
 
-Larger methyl esters such as methyl valerate (MV, \ce{C6H12O2}, methyl
-pentanoate) have also been studied as possible biodiesel surrogates.
-\cite{Hadj-Ali2009} used a rapid compression machine (RCM) to study the
-autoignition of several methyl esters including MV. Although MV exhibited
-two-stage ignition in this study, little additional research has been done on
-its oxidation. \cite{Korobeinichev2015} studied MV in premixed laminar
-flames and extended a detailed high temperature chemical kinetic model to
-include MV and methyl hexanoate. \cite{Dmitriev2015} added MV to
-n-heptane/toluene fuel blends to determine the resulting intermediate species in
-premixed flames using a flat burner at \SI{1}{atm} and an equivalence ratio of
-1.75. The addition of MV helped reduce soot forming intermediates including
-benzene, cyclopentadienyl, acetylene, propargyl, and
-vinylacetylene~\cite{Dmitriev2015}. \cite{Hayes2009} computationally
-examined the peroxy radical isomerization reactions for MV to better understand
-the low temperature reaction pathways. Finally, \cite{Dievart2013} used
-diffusion flames in the counterflow configuration to determine extinction limits
-for a number of methyl esters, including MV, and validated a detailed kinetic
+A recent review paper summarizes the work on methyl esters relevant to biodiesel
+combustion~\cite{Coniglio2013}; the following summary focuses on ignition delay measurements, since
+these are the focus of this paper. Autoignition of methyl butanoate (MB, \ce{C5H10O2}) has been
+well-studied in both shock tube and rapid compression machine experiments~
+\cite{Metcalfe2007,Walton2009,Dooley2008,Akih-Kumgeh2010,Akih-Kumgeh2011,Hadj-Ali2009,Kumar2016}.
+The prevalence of MB data in the literature is largely due to the early identification of MB as a
+potential surrogate fuel for biodiesel~\cite{Fisher2000}. However, the experiments have shown that
+MB may not be an appropriate surrogate for biodiesel, due to its lack of negative temperature
+coefficient (NTC) behavior, a requirement for a suitable biodiesel surrogate~\cite{Coniglio2013}.
+
+Larger methyl esters such as methyl valerate (MV, \ce{C6H12O2}, methyl pentanoate) have also been
+studied as possible biodiesel surrogates. \citet{Hadj-Ali2009} used a rapid compression machine
+(RCM) to study the autoignition of several methyl esters including MV. Although MV exhibited
+two-stage ignition in this study, little additional research has been done on its oxidation.
+\citet{Korobeinichev2015} studied MV in premixed laminar flames and extended a detailed high
+temperature chemical kinetic model to include MV and methyl hexanoate. \citet{Dmitriev2015} added
+MV to n-heptane/toluene fuel blends to determine the resulting intermediate species in premixed
+flames using a flat burner at \SI{1}{atm} and an equivalence ratio of 1.75. The addition of MV
+helped reduce soot forming intermediates including benzene, cyclopentadienyl, acetylene, propargyl,
+and vinylacetylene~\cite{Dmitriev2015}. \citet{Hayes2009} computationally examined the peroxy
+radical isomerization reactions for MV to better understand the low temperature reaction pathways.
+Finally, \citet{Dievart2013} used diffusion flames in the counterflow configuration to determine
+extinction limits for a number of methyl esters, including MV, and validated a detailed kinetic
 model with the experimental data.
 
-This work provides additional data for the autoignition of MV. Data is collected
-in a RCM under engine relevant conditions spanning
-from \SIrange{15}{30}{\bar}, equivalence ratios from \numrange{0.25}{2.0}, and
-temperatures from \SIrange{682}{1048}{\K}. The NTC region of MV is mapped out to
-provide additional information on the fidelity of using MV as a biodiesel
-surrogate.
+This work provides additional data for the autoignition of MV. Data is collected in a RCM under
+engine relevant conditions spanning from \SIrange{15}{30}{\bar}, equivalence ratios from
+\numrange{0.25}{2.0}, and temperatures from \SIrange{682}{1048}{\K}. The NTC region of MV is mapped
+out to provide additional information on the fidelity of using MV as a biodiesel surrogate.
 
@@
 
-The RCM used in this study is a single piston arrangement and is pneumatically
-driven and hydraulically stopped. The device has been described in detail
-previously~\cite{Mittal2007a} and will be described here briefly for
-reference. The end of compression (EOC) temperature and pressure (\(T_C\) and
-\(P_C\) respectively), are independently changed by varying the overall
-compression ratio, initial pressure \((P_0)\), and initial temperature \((T_0)\)
-of the experiments. The primary diagnostic on the RCM is the in-cylinder
-pressure. The pressure data is processed by a Python package called UConnRCMPy~\cite{uconnrcmpy}, which calculates \(P_C\), \(T_C\), and the ignition
-delay(s). The definitions of the ignition delays are shown in
-\cref{fig:ign-delay-def}. The time of the EOC is defined as the maximum of the
-pressure trace prior to the start of ignition and the ignition delays are
-defined as the time from the EOC until local maxima in the first time derivative
-of the pressure. Each experimental condition is repeated at least five times to
-ensure repeatability of the data. As there is some random scatter present in the
-data, the standard deviation \((\sigma)\) of the ignition delays from the runs
-at a given condition is computed. In all cases, \(\sigma\) is less than
-\SI{10}{\percent} of the mean value of the overall ignition delay.
+The RCM used in this study is a single piston arrangement and is pneumatically driven and
+hydraulically stopped. The device has been described in detail previously~\cite{Mittal2007a} and
+will be described here briefly for reference. The end of compression (EOC) temperature and pressure
+(\(T_C\) and \(P_C\) respectively), are independently changed by varying the overall compression
+ratio, initial pressure \((P_0)\), and initial temperature \((T_0)\) of the experiments. The piston
+in the reaction chamber is machined with a specially designed crevice to suppress the roll-up vortex
+effect and promote homogeneous conditions in the reactor during and after
+compression~\cite{Mittal2006}.
 
-In addition to the reactive experiments, non-reactive experiments are conducted
-to determine the influence of machine-specific behavior on the experimental
-conditions and permit the calculation of the EOC temperature via the isentropic
-relations between pressure and temperature~\cite{Lee1998}. The EOC
-temperature is calculated by the procedure described in
-\cref{sec:computational-methods}.
+The primary diagnostic on the RCM is the in-cylinder pressure measured by a Kistler 6125C dynamic
+transducer that is compensated for thermal shock. The transducer is coupled to a Kistler 5010B
+charge amplifier. The voltage output of the charge amplifier is recorded by a National Instruments
+9125 analog input device connected to a cDAQ 9178 chassis. The voltage is sampled at a rate of
+either \todo{Check this} 50 kHz or 100 kHz by a LabView VI and processed by a Python package called
+UConnRCMPy~\cite{Weber2016a}. Version 3.0.0 of UConnRCMPy~\cite{uconnrcmpy}, 3.6.1 of Python, 2.3.0
+of Cantera~\cite{cantera}, 1.13 of NumPy~\cite{vanderWalt2011}, 0.19.0 of SciPy~\cite{Jones2001},
+and 2.0.1 of Matplotlib~\cite{Hunter2007} were used in the analysis in this paper.
 
-\begin{figure}[htb]
-    \begin{minipage}[t]{0.48\textwidth}
+The compression stroke of the RCM brings the fuel/oxidizer mixture to the EOC conditions, and for
+suitable thermodynamic states, the mixture will ignite after a delay period. The definitions of the
+ignition delays are shown in \cref{fig:ign-delay-def}. The time of the EOC is defined as the maximum
+of the pressure trace prior to the start of ignition and the ignition delays are defined as the time
+from the EOC until local maxima in the first time derivative of the pressure. Each experimental
+condition is repeated at least five times to ensure repeatability of the data. As there is some
+random scatter present in the data, the standard deviation \((\sigma)\) of the ignition delays from
+the runs at a given condition is computed. In all cases, \(\sigma\) is less than \SI{10}{\percent}
+of the mean value of the overall ignition delay.
+
+\begin{figure}
     \centering
@@
     \label{fig:ign-delay-def}
-    \end{minipage}\hfill%
-    \begin{minipage}[t]{0.48\textwidth}
-        \centering
-        \resizebox{\linewidth}{!}{\input{figures/vapor_pressure.pgf}}
-        \caption{Saturated vapor pressure of MV as a function of temperature,
-        plotted using the Antoine equation, \cref{eq:antoine}, with
-        \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
-        \label{fig:vapor-pressure}
-    \end{minipage}
 \end{figure}
 
-The RCM is equipped with heaters to control the initial temperature of the
-mixture. After filling in the components to the mixing tanks, the heaters are
-switched on and the system is allowed \SI{1.5}{\hour} to come to steady state.
-The mixing tanks are also equipped with magnetic stir bars so the reactants are
-well mixed for the duration of the experiments.
+In addition to the reactive experiments, non-reactive experiments are conducted to determine the
+influence of machine-specific behavior on the experimental conditions and permit the calculation of
+the EOC temperature via the isentropic relations between pressure and temperature~\cite{Lee1998}.
+The EOC temperature is calculated by the procedure described in \cref{sec:computational-methods}.
 
-The initial temperature is chosen such that the saturated vapor pressure
-\((P_{\text{sat}})\) of the fuel at the initial temperature is at least twice
-the partial pressure of the fuel in the mixing tank. The Antoine equation
-%
-\begin{equation}\label{eq:antoine}
-    \logten{P_{\text{sat}}} = A - \frac{B}{T - C}
-\end{equation}
-%
-is used to model the saturated vapor pressure of MV as a function of
-temperature, where \(A\), \(B\), and \(C\) are substance-specific coefficients.
-The coefficients in \cref{eq:antoine} are determined by least squares fitting of
-the data of \cite{Ortega2003}, \cite{vanGenderen2002}, and
-\cite{Verevkin2008} using the \verb|curve_fit()| function of
-SciPy~\cite{Jones2001} version 0.18.1. \Cref{fig:vapor-pressure} shows that
-the coefficients fit with this procedure give good agreement with the
-experimental data.
+The mixtures considered in this study are shown in \cref{tab:mixtures}. Four equivalence ratios of
+MV in ``air'' are considered. The ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer is varied to adjust
+the temperatures reached at the EOC for a given mixture. Two \(P_C\) conditions are studied in this
+work, \SIlist{15;30}{\bar}, representing engine-relevant conditions. For the \(\phi=2.0\) condition,
+only \(P_C = \SI{15}{\bar}\) is considered because we could not achieve \(T_C\) values low enough
+that the ignition was long enough to be measured in our apparatus (the typical lower limit of
+ignition delay on the present RCM is approximately \SI{5}{\ms}).
 
-The mixtures considered in this study are shown in \cref{tab:mixtures}. Mixtures
-are prepared in stainless steel mixing tanks. The proportions of reactants in
-the mixture are determined by specifying the absolute mass of the fuel, the
-equivalence ratio \((\phi)\), and the ratio of \(\ce{Ar}:\ce{N2}\) in the
-oxidizer. Since MV is a liquid with a relatively small vapor pressure at room
-temperature, it is injected into the mixing tank through a septum. Proportions
-of \ce{O2}, \ce{Ar}, and \ce{N2} are added manometrically at room temperature.
+Mixtures are prepared in stainless steel mixing tanks, approximately \SI{17}{\L} and \SI{15}{\L} in
+size. The proportions of reactants in the mixture are determined by specifying the absolute mass of
+the fuel, the equivalence ratio \((\phi)\), and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer.
+Mixtures are made by first vacuuming the mixing tanks to an ultimate pressure less than
+\SI{5}{\torr}. Since MV is a liquid with a relatively small vapor pressure at room temperature and
+pressure, it is measured gravimetrically in a syringe to within \SI{0.01}{\g} of the specified
+value. The fuel is injected into the mixing tank through a septum. Proportions of \ce{O2}, \ce{Ar},
+and \ce{N2} are added manometrically at room temperature and the total pressure is measured by an
+Omega Engineering MMA type static pressure transducer. The same transducer is used to measure the
+pressure of the reactants prior to an experiment.
 
-\begin{table}[htb]
+\begin{table}
     \centering
@@
 
-% \begin{table}
-%     \centering
-%     \caption{Antoine Equation coefficients computed in this work and from the
-%     literature. The $2\sigma$ confidence interval is estimated by taking the
-%     square root of the diagonals of the covariance matrix returned from
-%     \texttt{curve\_fit()}}
-%     \label{tab:antoine}
-%     \begin{tabular}{cccccc}
-%         \toprule
-%         & $A$ & $B$ & $C$ & $T_{\text{min}}$, \si{\K} & $T_{\text{max}}$, \si{\K} \\
-%         \midrule
-%         This Work & 6.4030 & 1528.69 & 52.881 & 274.9 & 417.18 \\
-%         $2\sigma$ Confidence Interval & 0.0919 & 53.47 & 4.934 & --- & --- \\
-%         \cite{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
-%         \cite{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
-%         \cite{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
-%         \bottomrule
-%     \end{tabular}
-% \end{table}
+The RCM is equipped with heaters to control the initial temperature of the mixture. After filling in
+the components to the mixing tanks, the heaters are switched on and the system is allowed
+\SI{1.5}{\hour} to come to steady state. The mixing tanks are also equipped with magnetic stir bars
+so the reactants are well mixed for the duration of the experiments.
+
+The initial temperature is chosen such that the saturated vapor pressure \((P_{\text{sat}})\) of the
+fuel at the initial temperature is at least twice the partial pressure of the fuel in the mixing
+tank. The Antoine equation
+%
+\begin{equation}\label{eq:antoine}
+    \logten{P_{\text{sat}}} = A - \frac{B}{T - C}
+\end{equation}
+%
+is used to model the saturated vapor pressure of MV as a function of temperature, where \(A\),
+\(B\), and \(C\) are substance-specific coefficients. Coefficients for \cref{eq:antoine} are given
+in the literature by~\citet{Ortega2003}, \citet{Camacho2007}, and \citet{Stephenson1987};
+unfortunately, the values of the coefficients are different among all three authors. Therefore,
+coefficients for use in \cref{eq:antoine} are determined in this work by least squares fitting of
+the data of \citet{Ortega2003}, \citet{vanGenderen2002}, and \citet{Verevkin2008} using the
+\verb|curve_fit()| function of SciPy~\cite{Jones2001} version 0.19.0. \Cref{fig:vapor-pressure}
+shows that the coefficients fit with this procedure give good agreement with the experimental data;
+values for the coefficients computed in this work and in the literature works are given in
+\cref{tab:antoine}. The data used to calculate the coefficients are provided in the Supplementary
+Material.
+
+\begin{figure}
+    \centering
+    \resizebox{\linewidth}{!}{\input{figures/vapor_pressure.pgf}}
+    \caption{Saturated vapor pressure of MV as a function of temperature,
+    plotted using the Antoine equation, \cref{eq:antoine}, with
+    \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
+    \label{fig:vapor-pressure}
+\end{figure}
+
+\begin{table}
+    \centering
+    \caption{Antoine Equation coefficients computed in this work and from the
+    literature. The $2\sigma$ confidence interval is estimated by taking the
+    square root of the diagonals of the covariance matrix returned from
+    \texttt{curve\_fit()}}
+    \label{tab:antoine}
+    \begin{tabular}{cccccc}
+        \toprule
+        & $A$ & $B$ & $C$ & $T_{\text{min}}$, \si{\K} & $T_{\text{max}}$, \si{\K} \\
+        \midrule
+        This Work & 6.4030 & 1528.69 & 52.881 & 274.9 & 417.18 \\
+        $2\sigma$ Confidence Interval & 0.0919 & 53.47 & 4.934 & --- & --- \\
+        \cite{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
+        \cite{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
+        \cite{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
+        \bottomrule
+    \end{tabular}
+\end{table}
 
@@
 
-The Python 3.5 interface of Cantera~\cite{cantera} version 2.2.1 is used for
-all simulations in this work. Detailed descriptions of the use of Cantera for
-these simulations can be found in the work of \cite{Weber2016a} and
-\cite{Dames2016}; a brief overview is given here. As mentioned in
-\cref{sec:experimental-methods}, non-reactive experiments are conducted to
-characterize the machine-specific effects on the experimental conditions in the
-RCM. This pressure trace is used to compute a volume trace by assuming that the
-reactants undergo a reversible, adiabatic, constant composition (i.e.,
-isentropic) compression during the compression stroke and an isentropic
-expansion after the EOC. The volume trace is applied to a simulation conducted
-in an \verb|IdealGasReactor| in Cantera~\cite{cantera} using the CVODES
-solver from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delay from
-the simulations is defined in the same manner as in the experiments. The time
-derivative of the pressure in the simulations is computed by second order
-Lagrange polynomials, as discussed by \cite{Chapra2010}.
+The Python 3.6 interface of Cantera~\cite{cantera} version 2.3.0 is used for all simulations in this
+work. Detailed descriptions of the use of Cantera for these simulations can be found in the work
+of~\citet{Weber2016a} and~ \citet{Dames2016}; a brief overview is given here. As mentioned in
+\cref{sec:experimental-methods}, non-reactive experiments are conducted to characterize the
+machine-specific effects on the experimental conditions in the RCM. This pressure trace is combined
+with the reactive pressure trace and used to compute a volume trace by assuming that the reactants
+undergo a reversible, adiabatic, constant composition (i.e., isentropic) compression during the
+compression stroke and an isentropic expansion after the EOC. The volume trace is applied to a
+simulation conducted in an \verb|IdealGasReactor| in Cantera~\cite{cantera} using the CVODES
+solver from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delay from the simulations is
+defined in the same manner as in the experiments. The time derivative of the pressure in the
+simulations is computed by second order Lagrange polynomials, as discussed by \citet{Chapra2010}.
 
-To the best of our knowledge, there are three mechanisms for MV combustion
-available in the literature. The first two, by \cite{Korobeinichev2015} and
-\cite{Dmitriev2015}, were developed to simulate flames, and do not include
-the low-temperature chemistry necessary to simulate the conditions in these
-experiments. The third model was developed by \cite{Dievart2013} and
-includes low-temperature chemistry of MV, although it was only validated by
-comparison with flame extinction limits.
-% To use this mechanism in Cantera, we
-% made a number of modifications, including removing duplicate species from the
-% thermodynamic database and removing the excited \ce{OH*} and \ce{CH*} chemistry,
-% which is not supported by Cantera 2.2.1. For most of the duplicate species in
-% the thermodynamic database, the thermodynamic data is identical, so one copy of
-% the duplicate species is simply removed from the database. However, some species
-% are not exact duplicates. For these species, it is not clear from the
-% thermodynamic database file which data set should be preferred and the default
-% of Cantera (and CHEMKIN) to choose the first instance of a duplicate species is
-% used.
-The detailed \cite{Dievart2013} model includes 1103 species and 7557
-reactions.
+To the best of our knowledge, there are three mechanisms for MV combustion available in the
+literature. The first two, by \cite{Korobeinichev2015} and \cite{Dmitriev2015}, were developed to
+simulate flames, and do not include the low-temperature chemistry necessary to simulate the
+conditions in these experiments. The third model was developed by \cite{Dievart2013} and includes
+low-temperature chemistry of MV, although it was only validated by comparison with flame extinction
+limits. In converting this mechanism for use in Cantera, we found that there were many species in the
+thermodynamic database with multiple data entries. For most of these species the thermodynamic data
+is identical. However, some species are not exact duplicates. For these species, it is not clear
+from the thermodynamic database file which data set should be preferred. Since Cantera (and CHEMKIN)
+choose the first instance of a duplicate species to be used, we retained the first entry for all
+duplicated species. The detailed \cite{Dievart2013} model includes \todo{check this}1103 species and 7557 reactions,
+and the CHEMKIN and Cantera formatted input files are available in the Supplementary Material.
 
@@
 
-In addition to using a mechanism from the literature, we investigate the use of
-an automatic mechanism generator, the open-source Reaction Mechanism Generator
-(RMG)~\cite{Allen2012} version 1.0.4. The Python version of RMG is used,
-which requires Python 2.7, and version 1.10.0 of the RMG database is used. The
-final RMG model contains 483 species and 19990 reactions. Note that the number
-of species is much lower than the \cite{Dievart2013} model because the RMG
-model focuses on only one fuel (MV), but the number of reactions is
-substantially higher.
+In addition to using a mechanism from the literature, we investigate the use of an automatic
+mechanism generator, the open-source Reaction Mechanism Generator (RMG)~\cite{Allen2012} version
+2.1.0. The Python version of RMG is used, which requires Python 2.7, and version 2.1.0 of the RMG
+database is used. The final RMG model contains 483 species and 19990 reactions. Note that the number
+of species is much lower than the \cite{Dievart2013} model because the RMG model focuses on only one
+fuel (MV), but the number of reactions is substantially higher. The input file used to generate the
+model is available in the Supplementary Material.
 

### 2017-06-27 methyl-valerate.tex
Add updates to paper
Update vapor pressure

@@
-\documentclass[letterpaper,review]{elsarticle}
+\documentclass[letterpaper, review]{elsarticle}
 
@@
 
-\begin{figure}
-    \centering
-    \resizebox{\linewidth}{!}{\input{figures/ignition_delay_definition.pgf}}
+\begin{center}
+    \captionsetup{type=figure}
+    \includegraphics{figures/ignition_delay_definition.pdf}
     \caption{Definition of the ignition delays used in this work. The
     experiment in this figure was conducted for a \(\phi=2.0\) mixture with
-    \mbox{\(\ce{Ar}/(\ce{N2}+\ce{AR})=0.5\)}, \(P_0=\SI{0.7694}{\bar}\),
-    \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.94}{\bar}\), \(T_C=\SI{723}{\K}\),
-    \(\tau=\SI{27.44\pm0.99}{\ms}\), \(\tau_1=\SI{16.57\pm0.48}{\ms}\).}
+    \mbox{\(\ce{Ar}/(\ce{N2}+\ce{AR})=0.5\)}, \(P_0=\SI{0.7806}{\bar}\),
+    \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.92}{\bar}\), \(T_C=\SI{720}{\K}\),
+    \(\tau=\SI{27.56\pm0.89}{\ms}\), \(\tau_1=\SI{16.60\pm0.46}{\ms}\).}
     \label{fig:ign-delay-def}
-\end{figure}
+\end{center}
 
@@
 
-\begin{table}
-    \centering
+\begin{center}
+    \captionsetup{type=table}
     \caption{Mixtures considered in this work}
@@
     \label{tab:mixtures}
-\end{table}
+\end{center}
 
@@
 \(B\), and \(C\) are substance-specific coefficients. Coefficients for \cref{eq:antoine} are given
-in the literature by~\citet{Ortega2003}, \citet{Camacho2007}, and \citet{Stephenson1987};
-unfortunately, the values of the coefficients are different among all three authors. Therefore,
-coefficients for use in \cref{eq:antoine} are determined in this work by least squares fitting of
-the data of \citet{Ortega2003}, \citet{vanGenderen2002}, and \citet{Verevkin2008} using the
-\verb|curve_fit()| function of SciPy~\cite{Jones2001} version 0.19.0. \Cref{fig:vapor-pressure}
-shows that the coefficients fit with this procedure give good agreement with the experimental data;
-values for the coefficients computed in this work and in the literature works are given in
-\cref{tab:antoine}. The data used to calculate the coefficients are provided in the Supplementary
-Material.
+in the literature by~\citet{Ortega2003}, \citet{Camacho2007}, and~\citet{Stephenson1987}.
+Unfortunately, the values of the coefficients are different among all three authors and, more
+importantly, the temperature ranges provided in those three fits do not cover the entire range of
+interest for this study. Therefore, coefficients for use in \cref{eq:antoine} are determined in this
+work by least squares fitting of the data of \citet{Ortega2003}, \citet{vanGenderen2002}, and
+\citet{Verevkin2008} using the \verb|curve_fit()| function of SciPy~\cite{Jones2001} version 0.19.0.
+\Cref{fig:vapor-pressure} shows that the coefficients fit with this procedure give good agreement
+with the experimental data; values for the coefficients computed in this work and in the literature
+works are given in \cref{tab:antoine}. The data used to calculate the coefficients are provided in
+the Supplementary Material.
 
-\begin{figure}
-    \centering
-    \resizebox{\linewidth}{!}{\input{figures/vapor_pressure.pgf}}
+\begin{center}
+    \captionsetup{type=figure}
+    \input{figures/vapor_pressure.pgf}
     \caption{Saturated vapor pressure of MV as a function of temperature,
@@
     \label{fig:vapor-pressure}
-\end{figure}
+\end{center}
 
-\begin{table}
-    \centering
+\begin{center}
+    \captionsetup{type=table}
     \caption{Antoine Equation coefficients computed in this work and from the
@@
         $2\sigma$ Confidence Interval & 0.0919 & 53.47 & 4.934 & --- & --- \\
-        \cite{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
-        \cite{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
-        \cite{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
+        \citet{Ortega2003} & 6.23175 & 1429.00 & 62.30 & 364.75 & 417.18 \\
+        \citet{Camacho2007} & 5.9644 & 1281.06 & 75.94 & 281 & 547 \\
+        \citet{Stephenson1987} & 6.62646 & 1658.4 & 42.09 & 297 & 411 \\
         \bottomrule
     \end{tabular}
-\end{table}
+\end{center}
 
@@
     \centering
-    \input{figures/ignition-delays.pgf}
+    \resizebox{\linewidth}{!}{\input{figures/ignition-delays.pgf}}
     \caption{Ignition delays of MV as a function of inverse temperature. Filled

### 2017-06-27 methyl-valerate.tex
Updates

@@
 choose the first instance of a duplicate species to be used, we retained the first entry for all
-duplicated species. The detailed \cite{Dievart2013} model includes \todo{check this}1103 species and 7557 reactions,
+duplicated species. The detailed \cite{Dievart2013} model includes 1105 species and 7141 reactions,
 and the CHEMKIN and Cantera formatted input files are available in the Supplementary Material.
@@
 2.1.0. The Python version of RMG is used, which requires Python 2.7, and version 2.1.0 of the RMG
-database is used. The final RMG model contains 483 species and 19990 reactions. Note that the number
-of species is much lower than the \cite{Dievart2013} model because the RMG model focuses on only one
-fuel (MV), but the number of reactions is substantially higher. The input file used to generate the
-model is available in the Supplementary Material.
+database is used. The final RMG model contains 427 species and 13640 reactions. Note that the number
+of species is much lower than the \citet{Dievart2013} model because the RMG model focuses on only
+one fuel (MV), but the number of reactions is substantially higher. The input file used to generate
+the model is available in the Supplementary Material.
 

### 2017-06-27 methyl-valerate.tex
Updates for ignition delay

@@
 
-\Cref{fig:ignition-delays} shows the ignition delay results measured in this
-study. Filled markers denote the overall ignition delay and hollow markers
-indicate the first-stage ignition delay. Vertical error bars are drawn on the
-symbols to represent the uncertainty in the ignition delay; for many of the
-experiments, the uncertainty is approximately the same size as the data point,
-so the error bar is hidden. Horizontal error bars are shown on the first and
-last points of each equivalence ratio indicating the estimated uncertainty in
-the EOC temperature of $\pm\SI{1}{\percent}$~\cite{Weber2015}.
-\cref{fig:ignition-delays}a shows the results for a compressed pressure of
-\SI{15}{\bar}, while \cref{fig:ignition-delays}b shows the results for a
-compressed pressure of \SI{30}{\bar}. Note that $\phi=2.0$ results were not
-collected for \SI{30}{\bar}, so there are no red data points in
-\cref{fig:ignition-delays}b.
+\Cref{fig:ignition-delays} shows the ignition delay results measured in this study. Filled markers
+denote the overall ignition delay and hollow markers indicate the first-stage ignition delay.
+Vertical error bars are drawn on the symbols to represent the uncertainty in the ignition delay; for
+many of the experiments, the uncertainty is approximately the same size as the data point, so the
+error bar is hidden. Horizontal error bars are shown on the first and last points of each
+equivalence ratio indicating the estimated uncertainty in the EOC temperature of
+$\pm\SI{1}{\percent}$~\cite{Weber2015}. \cref{fig:ignition-delays}a shows the results for a
+compressed pressure of \SI{15}{\bar}, while \cref{fig:ignition-delays}b shows the results for a
+compressed pressure of \SI{30}{\bar}. Note that $\phi=2.0$ results were not collected for
+\SI{30}{\bar}, so there are no red data points in \cref{fig:ignition-delays}b.
 
-It can be seen from \cref{fig:ignition-delays} that the ignition delays for the
-\(\phi=0.25\text{ and } 0.5\) mixtures do not show an NTC region of the ignition
-delay for both of the pressures studied in this work. However, the $\phi=1.0$
-mixture shows an NTC region at $P_C=\SI[number-unit-product={\ }]{30}{\bar}$
-between approximately \SIlist{720;800}{\K}, with measured first-stage ignition
-delays at \SIlist{734;757}{\K}. In addition, the $\phi=2.0$ mixture shows an NTC
-region of ignition delay at \SI{15}{\bar} from approximately
-\SIrange{720}{775}{\K}, with measured first-stage ignition delays between
-\SIlist{720;750}{\K}.
+It can be seen from \cref{fig:ignition-delays} that the ignition delays for the \(\phi=0.25\text{
+and } 0.5\) mixtures do not show an NTC region of the ignition delay for both of the pressures
+studied in this work. However, the $\phi=1.0$ mixture shows an NTC region at
+$P_C=\SI[number-unit-product={\ }]{30}{\bar}$ between approximately \SIlist{720;800}{\K}, with
+measured first-stage ignition delays at \SIlist{733;757}{\K}. In addition, the $\phi=2.0$ mixture
+shows an NTC region of ignition delay at \SI{15}{\bar} from approximately \SIrange{720}{780}{\K},
+with measured first-stage ignition delays between \SIlist{720;750}{\K}.
 
-\begin{figure}[htb]
-    \centering
-    \resizebox{\linewidth}{!}{\input{figures/ignition-delays.pgf}}
+\begin{center}
+    \captionsetup{type=figure}
+    \includegraphics{figures/ignition-delays.pdf}
     \caption{Ignition delays of MV as a function of inverse temperature. Filled
@@
     \label{fig:ignition-delays}
-\end{figure}
+\end{center}
 
-\Cref{fig:pressure-traces} shows the pressure traces for selected experiments at
-\(\phi=1.0\), \(P_C =\SI[number-unit-product={\ }]{30}{\bar}\). The three
-reactive pressure traces shown are at the low-temperature end of the NTC (blue,
-\SI{700}{\K}), one case with two-stage ignition (orange, \SI{734}{\K}), and one
-case near the high-temperature limit of the NTC region (green, \SI{775}{\K}).
-Also shown is the non-reactive pressure trace for the \SI{700}{\K} case (red).
-By comparing the \SI{700}{\K} pressure trace with the non-reactive pressure
-trace, it can be seen that there is substantial heat release prior to main
-ignition as measured by the deviation of the reactive pressure trace from the
-non-reactive trace. However, there is only one peak in the time derivative of
-the pressure, so no first-stage ignition delay is defined for this case. It can
-also be seen in \cref{fig:pressure-traces} that the \SI{775}{\K} case shows some
-heat release prior to ignition, although again there is only one peak in the
-time derivative of the pressure. Furthermore, the heat release at \SI{775}{\K}
-appears to be more gradual than at the lowest temperature.
+\Cref{fig:pressure-traces}a shows the pressure traces for selected experiments at \(\phi=1.0\),
+\(P_C =\SI[number-unit-product={\ }]{30}{\bar}\). The three reactive pressure traces shown are at
+the low-temperature end of the NTC (blue, \SI{700}{\K}), one case with two-stage ignition (orange,
+\SI{733}{\K}), and one case near the high-temperature limit of the NTC region (green, \SI{774}{\K}).
+Also shown is the non-reactive pressure trace for the \SI{700}{\K} case (red). By comparing the
+\SI{700}{\K} pressure trace with the non-reactive pressure trace, it can be seen that there is
+substantial heat release prior to main ignition as measured by the deviation of the reactive
+pressure trace from the non-reactive trace. However, there is only one peak in the time derivative
+of the pressure, so no first-stage ignition delay is defined for this case. It can also be seen in
+\cref{fig:pressure-traces}a that the \SI{775}{\K} case shows some heat release prior to ignition,
+although again there is only one peak in the time derivative of the pressure. Furthermore, the heat
+release at \SI{775}{\K} appears to be more gradual than at the lowest temperature.
 
-\begin{figure}[htb]
-    \begin{minipage}[t]{0.48\textwidth}
-        \centering
-        \resizebox{\linewidth}{!}{\input{figures/pressure-traces.pgf}}
-        \caption{Selected pressure traces around the NTC region of ignition delay for $\phi=1.0$}
+\begin{center}
+    \captionsetup{type=figure}
+    \includegraphics[width=\textwidth]{figures/pressure-traces.pdf}
+    \caption{Selected pressure traces around the NTC region of ignition delay.
+    a) \(\phi=1.0\) b) \(\phi=2.0\)}
     \label{fig:pressure-traces}
-    \end{minipage}\hfill%
-    \begin{minipage}[t]{0.48\textwidth}
-        \centering
-        \resizebox{\linewidth}{!}{\input{figures/simulation-comparison.pgf}}
-        \caption{Comparison of experimental and simulated results for $\phi=1.0$}
-        \label{fig:simulation-comparison}
-    \end{minipage}
-\end{figure}
+    % \begin{minipage}[t]{0.48\textwidth}
+    %     \centering
+    %     \resizebox{\linewidth}{!}{\input{figures/simulation-comparison.pgf}}
+    %     \caption{Comparison of experimental and simulated results for $\phi=1.0$}
+    %     \label{fig:simulation-comparison}
+    % \end{minipage}
+\end{center}
+
+A similar trend can be observed in \cref{fig:pressure-traces}b for \(\phi=2.0\) at \(
 

### 2017-06-27 methyl-valerate.tex
Pressure trace updates

@@
 
-A similar trend can be observed in \cref{fig:pressure-traces}b for \(\phi=2.0\) at \(
+A similar trend can be observed in \cref{fig:pressure-traces}b for \(\phi=2.0\) at
+\(P_C=\SI{15}{\bar}\), where pressure traces at several points around the NTC region are plotted. As
+in \cref{fig:pressure-traces}a, the three reactive pressure traces shown are at the low-temperature
+end of the NTC (blue, \SI{698}{\K}), one case with two-stage ignition (orange, \SI{720}{\K}), and
+one case near the high-temperature limit of the NTC region (green, \SI{782}{\K}). Also shown is the
+non-reactive pressure trace for the \SI{698}{\K} case (red). As for the \(\phi=1.0\) case, the
+pressure traces show significant heat release prior to the overall ignition, as judged by deviation
+from the non-reactive case.
 

### 2017-06-27 methyl-valerate.tex
Update simulation-comparison

@@
 
-\Cref{fig:simulation-comparison} compares experimentally measured overall
-ignition delays with ignition delays computed with the detailed model of
-\cite{Dievart2013} for the $\phi=1.0$ experiments. Results for the other
-equivalence ratios are similar to these results, so are not shown here. It is
-important to note that the model of \cite{Dievart2013} was not validated for
-MV ignition delays, only for extinction strain rates. At \SI{15}{\bar}, the
-model tends to under-predict the ignition delay and predicts an NTC region that
-is not present in the experiments. At \SI{30}{\bar}, the model predicts the
-low-temperature ignition delays well, but does not predict the NTC region found
-experimentally.
+\Cref{fig:simulation-comparison} compares experimentally measured overall ignition delays with
+ignition delays computed with the detailed model of \citet{Dievart2013} for the \(\phi=1.0\)
+experiments. Results for the other equivalence ratios are similar to these results, so are not shown
+here. It is important to note that the model of \citet{Dievart2013} was not validated for MV
+ignition delays, only for extinction strain rates. At \SI{15}{\bar}, the model tends to
+under-predict the ignition delay and predicts an NTC region that is not present in the experiments.
+At \SI{30}{\bar}, the model predicts the low-temperature ignition delays well, but does not predict
+the NTC region found experimentally.
 

### 2017-06-28 methyl-valerate.tex
Update simulation figure

--- 2017-06-27 methyl-valerate.tex
@@
     \label{fig:pressure-traces}
-    % \begin{minipage}[t]{0.48\textwidth}
-    %     \centering
-    %     \resizebox{\linewidth}{!}{\input{figures/simulation-comparison.pgf}}
-    %     \caption{Comparison of experimental and simulated results for $\phi=1.0$}
-    %     \label{fig:simulation-comparison}
-    % \end{minipage}
 \end{center}
@@
 
-\Cref{fig:simulation-comparison} compares experimentally measured overall
-ignition delays with ignition delays computed with the detailed model of
-\cite{Dievart2013} for the $\phi=1.0$ experiments. Results for the other
-equivalence ratios are similar to these results, so are not shown here. It is
-important to note that the model of \cite{Dievart2013} was not validated for
-MV ignition delays, only for extinction strain rates. At \SI{15}{\bar}, the
-model tends to under-predict the ignition delay and predicts an NTC region that
-is not present in the experiments. At \SI{30}{\bar}, the model predicts the
-low-temperature ignition delays well, but does not predict the NTC region found
-experimentally.
+\Cref{fig:simulation-comparison} compares experimentally measured overall ignition delays with
+ignition delays computed with the detailed model of \citet{Dievart2013} (solid lines).
+\Cref{fig:simulation-comparison}a shows results at \(P_C = \SI{15}{\bar}\), while
+\cref{fig:simulation-comparison}b shows results at \(P_C = \SI{30}{\bar}\). Only some equivalence
+ratios are shown for each pressure condition; data and simulated results are not shown for cases
+where the reactive simulated temperature at the EOC deviated substantially from the non-reactive
+temperature due to heat release during the compression stroke. Furthermore, it is important to note
+that the model of \citet{Dievart2013} was not validated for MV ignition delays, only for extinction
+strain rates.
 
-To understand the underlying reasons for the disagreement between the
-\cite{Dievart2013} model and the data, we constructed an additional model
-using RMG (see \cref{sec:reaction-mechanism-generator}). As can be seen in
-\cref{fig:simulation-comparison}, the agreement between the RMG model and the
-experimental data is similar to the \cite{Dievart2013} model for the
-\SI{30}{\bar} data. At \SI{15}{\bar}, the RMG model predicts a somewhat longer
-ignition delay than the model of \cite{Dievart2013}, but still predicts an
-NTC region where none is present in the experimental data.
+\begin{center}
+    \captionsetup{type=figure}
+    \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
+    \caption{Comparison of experimental and simulated results. a) \SI{15}{\bar}; b) \SI{30}{\bar}}
+    \label{fig:simulation-comparison}
+\end{center}
+
+At \SI{15}{\bar}, the experimental ignition delays are under-predicted by the \citet{Dievart2013}
+model for the three equivalence ratios shown. For the \(\phi =\) \numlist{0.25;0.5} conditions, the
+model appears to be predicting an NTC region of the ignition delays as the temperature decreases,
+although such a trend is not observed in the data. However, at \(\phi = 2.0\), the model does not
+predict the presence of an NTC region, although one is present in the experiments. Nonetheless, the
+agreement seems to be improving as the temperature is decreased. Comparing the \citet{Dievart2013}
+model to the data at \SI{30}{\bar}, we find a similar trend as the \(\phi=2.0,\ P_C=\SI{15}{\bar}\)
+data. The model does not predict the NTC region found experimentally for the \(\phi = 1.0\)
+experiments, but the agreement improves as the temperature decreases.
+
+To elucidate the underlying reasons for the disagreement between the \citet{Dievart2013} model and
+the data, we constructed an additional model using RMG (see
+\cref{sec:reaction-mechanism-generator}). As can be seen in \cref{fig:simulation-comparison}a, the
+agreement between the RMG model (dashed lines) and the experimental data is similar to the
+\citet{Dievart2013} model for the \SI{15}{\bar}, \(\phi =\) \numlist{0.25; 0.5} data. Moreover, the
+RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar},
+\(\phi = 2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays, and
+does not predict the NTC region found experimentally. As before, the trend at \SI{30}{\bar},
+\(\phi=1.0\) is similar to the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the
+low-temperature ignition delays and does not predict the experimental NTC region.
 

### 2017-06-28 methyl-valerate.tex
Update ignition delay section

@@
 \section{Experimental Results}\label{sec:experimental-results}
+\subsection{Ignition Delays}\label{sec:ignition-delays}
 
@@
 and } 0.5\) mixtures do not show an NTC region of the ignition delay for both of the pressures
-studied in this work. However, the $\phi=1.0$ mixture shows an NTC region at
-$P_C=\SI[number-unit-product={\ }]{30}{\bar}$ between approximately \SIlist{720;800}{\K}, with
-measured first-stage ignition delays at \SIlist{733;757}{\K}. In addition, the $\phi=2.0$ mixture
+studied in this work. However, the \(\phi=1.0\) mixture shows an NTC region at
+\(P_C=\SI[number-unit-product={\ }]{30}{\bar}\) between approximately \SIlist{720;800}{\K}, with
+measured first-stage ignition delays at \SIlist{733;757}{\K}. In addition, the \(\phi=2.0\) mixture
 shows an NTC region of ignition delay at \SI{15}{\bar} from approximately \SIrange{720}{780}{\K},
@@
 \end{center}
+
+\citet{Hadj-Ali2009} also observed two-stage ignition of MV in stoichiometric mixtures, stating that
+``[m]ethyl pentanoate... was more reactive [than methyl butanoate] with a limit below which
+autoignition no longer occurs observed at \(T_c = \SI{670}{\K}\) and \(P_c = \SI{11.4}{\bar}\). At
+this temperature, the autoignition occurred in two stages with a clearly identified cool flame
+event.''\cite{Hadj-Ali2009} However, we do not find two stage ignition for the similar pressure of
+\(P_C = \SI{15}{\bar}\) in this study. We note that the stated temperature of the experiment from
+the work of \citet{Hadj-Ali2009} (\SI{670}{\K}) is much lower than the lowest temperature we
+considered in this work at \SI{15}{\bar}, \(\phi=1.0\) (\SI{828}{\K}). We did not conduct
+experiments at lower temperatures because the work of \citet{Mittal2006} showed that the temperature
+field in the RCM reaction chamber was uniform for approximately \SI{100}{\ms} after the EOC, and our
+measured ignition delay at \SI{15}{\bar}, \(\phi=1.0\) \SI{828}{\K} is \SI{92.14}{\ms}.
+
+However, we note NTC behavior of the ignition delay and two-stage ignition at the higher pressure of
+\SI{30}{\bar}, and at higher temperatures than those reported for two-stage ignition in the study of
+\citet{Hadj-Ali2009}. The trend of NTC behavior shifting to higher temperatures with increasing
+pressure can be seen in other classes of fuels. \citet{Kukkadapu2012} found a similar trend in
+gasoline composed of iso-alkanes, n-alkanes, cyclo-alkanes, aromatics, and olefins.
+\citet{Kukkadapu2012} attributed the NTC shift to the reactions between the hydroperoxyalkyl radical
+(\ce{QOOH}) and \ce{O2} becoming more dominant than the unimolecular decomposition of \ce{QOOH} at
+higher pressures. Similar trends could occur for the hydroperoxyalkyl radicals of MV.
+
+To further understand the effect of the methyl ester functional group on the NTC region of ignition
+delay, we compare with the alkane and alcohol with 5-carbon alkyl chains, n-pentane and n-pentanol.
+n-Pentane and MV have the same fuel mole percentage for stoichiometric mixtures in air
+(\SI{2.56}{\percent}), while n-pentanol has a fuel mole percentage of \SI{2.72}{\percent} for
+stoichiometric conditions. \citet{Ribaucour1998} and \citet{Bugler2015} found the NTC region for
+n-pentane to be between \SIlist{760;910}{\K} at pressures near \SI{10}{\atm}. As we will compare
+with our MV data at \SI{30}{\bar}, we note that increasing the pressure tends to shift the NTC
+to higher temperatures, as mentioned previously \cite{Kukkadapu2012}. \citet{Heufer2013} found NTC
+behavior for n-pentanol in the range of \SIrange{770}{900}{\K} at \SI{30}{\bar}. In this study, we
+find the NTC window for MV at \SI{30}{\bar} to be between \SIlist{720;800}{\K}. Therefore, it
+appears that the methyl ester functional group shifts the NTC range to lower temperature as compared
+to alkanes and alcohols with similar alkyl chain lengths. This result was also noted by
+\citet{Hadj-Ali2009} for MHEX as the fuel.
+
+\subsection{Pressure Traces}\label{sec:pressure-traces}
 

### 2017-06-29 methyl-valerate.tex
Updates

@@
 
-In general, there could be three likely sources of error in the models: missing
-reaction pathways, incorrect values of the reaction rates, and incorrect values
-for thermodynamic properties of the species. We have noted in
-\cref{sec:reaction-mechanism-generator} that the RMG model has many more
-reactions than the \cite{Dievart2013} model and the algorithm used in RMG
-considers a substantial number of the possible pathways. This reduces the
-possibility of missing reaction pathways affecting the model. Further detailed
-studies are required to ensure that the RMG model includes all of the relevant
-reaction pathways.
+It is clear that neither model is able to predict the ignition delays of MV particularly well. In
+addition to the poor agreement shown in \cref{fig:simulation-comparison}, the simulations for
+\(P_C = \SI{15}{\bar}\), \(\phi=1.0\) and \(P_C = \SI{30}{\bar}\), \(\phi=\) \numlist{0.25;0.5;2.0}
+showed substantial heat release during the compression stroke (i.e., the simulations are much too
+reactive), and so these conditions aren't compared in \cref{fig:simulation-comparison}. We note
+again that the model by \citet{Dievart2013} was validated for MV combustion only by comparison to
+flame extinction limits, so the disagreement is not wholly surprising.
 
-The second source of error may be incorrect reaction rate parameters, either
-because the rates are specified incorrectly in the model (e.g., typos) or
-because the rates are not well estimated by the typical analogy based-rules. It
-should be noted that errors of this type may affect the model generated by
-RMG---if the rates are not estimated correctly, reactions that are important in
-reality may not be included in the model. Determining the accuracy of the
-reaction rates used in the RMG and \cite{Dievart2013} models requires
-further detailed studies of the models. Another related source of error could be
-incorrect estimation of the pressure dependence of the reaction rates, which may
-be particularly important for the isomerization reactions prevalent in
-low-temperature chemistry.
+In general, there could be three likely sources of error in the models: missing reaction pathways,
+incorrect values of the reaction rates, and incorrect values for thermodynamic properties of the
+species. We have noted in \cref{sec:reaction-mechanism-generator} that the RMG model has many more
+reactions than the \citet{Dievart2013} model and the algorithm used in RMG considers a substantial
+number of the possible pathways. This reduces the possibility of missing reaction pathways affecting
+the model. Further detailed studies are required to ensure that the RMG model includes all of the
+relevant reaction pathways, which are outside the scope of this work.
 
-The third source of error may lie in the estimation of the thermodynamic
-properties of the species, particularly their heats of formation. We have begun
-to analyze the possibility of this source of error by conducting a reaction
-pathway analysis to determine which radicals are formed from the breakdown of
-the fuel. The following analysis is conducted for a constant volume simulation
-at \SI{700}{\K}, \SI{30}{\bar}, where the rates of production of the species
-have been integrated until the time of \SI{20}{\percent} fuel consumption. The
-results of this analysis are shown in \cref{fig:mv-structure} and
-\cref{tab:mv-radicals} for the two models. The percentages shown in the
-\cref{tab:mv-radicals} are the percent of the fuel destroyed to form a
-particular fuel radical by all the reactions that can form that radical.
+The second source of error may be incorrect reaction rate parameters, either because the rates are
+specified incorrectly in the model (e.g., typos) or because the rates are not well estimated by the
+typical analogy based-rules. It should be noted that errors of this type may affect the model
+generated by RMG---if the rates are not estimated correctly, reactions that are important in reality
+may not be included in the model. Determining the accuracy of the reaction rates used in the RMG and
+\citet{Dievart2013} models requires further detailed studies of the models, which are also outside
+the scope of this work. Another, related, source of error could be incorrect estimation of the
+pressure dependence of the reaction rates, which may be particularly important for the isomerization
+reactions prevalent in low-temperature chemistry.
+
+The third source of error may lie in the estimation of the thermodynamic properties of the species,
+particularly the fuel radicals. By conducting a reaction pathway analysis to determine which
+radicals are formed from the breakdown of the fuel, we can analyze the proportion of each radical
+formed as the fuel breaks down during the autoignition process. The following analysis is conducted
+for a constant volume simulation at \SI{700}{\K}, \SI{30}{\bar}, and the stoichiometric equivalence
+ratio, where the rates of production of the species have been integrated until the time of
+\SI{20}{\percent} fuel consumption. The results of this analysis are shown in
+\cref{fig:mv-structure} and \cref{tab:mv-radicals} for the two models. The percentages shown in the
+\cref{tab:mv-radicals} are the percent of the fuel destroyed to form a particular fuel radical by
+all the reactions that can form that radical.
 
@@
 
-\begin{figure}
-    \begin{minipage}[t][][b]{0.3\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{figures/Methyl_pentanoate.png}
+\begin{center}
+    \captionsetup{type=figure}
+    \includegraphics[width=90mm]{figures/Methyl_pentanoate.png}
     \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:mv-radicals}}
     \label{fig:mv-structure}
-    \end{minipage}\hfill%
-    \begin{minipage}[t][][b]{0.68\textwidth}
-        \centering
-        \captionof{table}{Percent of MV destroyed to form fuel radical species with a hydrogen atom missing at the location indicated in the first column}
+\end{center}
+
+\begin{center}
+    \captionsetup{type=table}
+    \caption{Percent of MV destroyed to form fuel radical species with a hydrogen atom missing at
+    the location indicated in the first column and \cref{fig:mv-structure}}
     \label{tab:mv-radicals}
-        \begin{tabular}{cSS}
+    \begin{tabular}{cSSSS}
         \toprule
-            Radical Site & {\cite{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} \\
+        Radical Site & {\citet{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} & & \\
         \midrule
-            2 & 29.3 & 7.4 \\
-            3 & 17.5 & 36.0 \\
-            4 & 17.5 & 41.1 \\
-            5 & 9.4 & 3.7 \\
-            M & 26.3 & 11.8 \\
+        2 & 29.3 & 12.5 & 27.7 & 15.6\\
+        3 & 17.5 & 12.2 & 17.8 & 20.1\\
+        4 & 17.5 & 50.6 & 17.8 & 23.2 \\
+        5 & 9.4 & 3.9 & 10.1 & 2.7 \\
+        M & 26.3 & 20.8 & 26.6 & 38.4\\
         \bottomrule
     \end{tabular}
-    \end{minipage}
-\end{figure}
+\end{center}
 
@@
 
-In this study, we have measured ignition delays for methyl valerate over a wide
-range of engine-relevant pressures, temperatures, and equivalence ratios. An NTC
-region of the ignition delay and two-stage ignition were recorded for pressures
-of \SI{15}{\bar} at \(\phi=2.0\) and \SI{30}{\bar} at \(\phi=1.0\). A detailed
-chemical kinetic model available in the literature was unable to reproduce the
-experimental results, so a new model was constructed using the Reaction
-Mechanism Generator software. Although the new model contains many more
-reactions than the literature model, it is still unable to predict the
-experimental ignition delays satisfactorily. Possible reasons for the
-discrepancy include missing reaction pathways, incorrect rate estimates, and
-incorrect thermodynamic property estimates. Future work will include
-investigation of the discrepancies between models and experiments to further
-understand the autoignition kinetics of methyl valerate.
+In this study, we have measured ignition delays for methyl valerate over a wide range of
+engine-relevant pressures, temperatures, and equivalence ratios. An NTC region of the ignition delay
+and two-stage ignition were recorded for pressures of \SI{15}{\bar} at \(\phi=2.0\) and
+\SI{30}{\bar} at \(\phi=1.0\). A detailed chemical kinetic model available in the literature was
+unable to reproduce the experimental results, so a new model was constructed using the Reaction
+Mechanism Generator software. Although the new model contains many more reactions than the
+literature model, it is still unable to predict the experimental ignition delays satisfactorily.
+Possible reasons for the discrepancy include missing reaction pathways, incorrect rate estimates,
+and incorrect thermodynamic property estimates. Future work will include investigation of the
+discrepancies between models and experiments to further understand the autoignition kinetics of
+methyl valerate.
 

### 2017-07-10 methyl-valerate.tex
Analysis

@@
 
-\usepackage{todonotes}
+\usepackage{multirow}
 
@@
 \usepackage{siunitx}
-\sisetup{separate-uncertainty=true}
+\sisetup{
+    separate-uncertainty=true,
+    per-mode=symbol
+}
 
@@
 9125 analog input device connected to a cDAQ 9178 chassis. The voltage is sampled at a rate of
-either \todo{Check this} 50 kHz or 100 kHz by a LabView VI and processed by a Python package called
+either 50 kHz or 100 kHz by a LabView VI and processed by a Python package called
 UConnRCMPy~\cite{Weber2016a}. Version 3.0.0 of UConnRCMPy~\cite{uconnrcmpy}, 3.6.1 of Python, 2.3.0
@@
 The third source of error may lie in the estimation of the thermodynamic properties of the species,
-particularly the fuel radicals. By conducting a reaction pathway analysis to determine which
-radicals are formed from the breakdown of the fuel, we can analyze the proportion of each radical
-formed as the fuel breaks down during the autoignition process. The following analysis is conducted
-for a constant volume simulation at \SI{700}{\K}, \SI{30}{\bar}, and the stoichiometric equivalence
-ratio, where the rates of production of the species have been integrated until the time of
-\SI{20}{\percent} fuel consumption. The results of this analysis are shown in
-\cref{fig:mv-structure} and \cref{tab:mv-radicals} for the two models. The percentages shown in the
-\cref{tab:mv-radicals} are the percent of the fuel destroyed to form a particular fuel radical by
-all the reactions that can form that radical.
-
-At the relatively low temperature and high pressure condition of this analysis,
-all of the fuel is destroyed by \ce{H}-atom abstractions to form the fuel
-radicals shown. It can be seen that the two models have quite different
-distributions of products from the first \ce{H}-abstraction reactions. The model
-of \cite{Dievart2013} predicts that \ce{H}-abstraction from the second
-carbon is the most prevalent, followed closely by abstraction from the methyl
-group. This is in line with the bond energies of the \ce{C-H} bonds for those
-carbon atoms; we expect that the presence of the oxygen atoms will cause
-hydrogen abstraction at the nearby carbons to be favored. However, the RMG model
-predicts that radicals in the middle of the carbon chain will be primarily
-formed. The cause of this discrepancy is under investigation, but it may be
-caused by the estimation of thermodynamic properties of the radicals.
+particularly the fuel radicals. In the work of \citet{Dievart2013}, the program THERM
+\cite{Ritter1991} is used to estimate thermodynamic values using the group additivity method. In the
+RMG model constructed in this work, RMG itself estimates the thermodynamic properties of the
+molecules also using the group additivity method. Nonetheless, the two models have differing
+predictions of the thermodynamic properties of the species in the model, particularly the fuel and
+its radicals. The values of the heats of formation of the fuel and its \ce{H}-atom abstraction
+radicals are shown in \cref{tab:heat-of-formation}; the radicals are labeled according to the
+convention shown in \cref{fig:mv-structure}.
 
@@
     \captionsetup{type=figure}
-    \includegraphics[width=90mm]{figures/Methyl_pentanoate.png}
-    \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:mv-radicals}}
+    \includegraphics[width=60mm]{figures/Methyl_pentanoate.png}
+    \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:heat-of-formation} and \cref{tab:mv-radicals}}
     \label{fig:mv-structure}
 \end{center}
+
+\begin{center}
+    \captionsetup{type=table}
+    \caption{Heats of formation of MV and its radicals, labeled according to the convention used in \cref{fig:mv-structure}}
+    \label{tab:heat-of-formation}
+    \begin{tabular}{cSSSS}
+        \toprule
+        \multirow{2}{*}{Radical Site} & \multicolumn{2}{c}{\citet{Dievart2013}} & \multicolumn{2}{c}{RMG Model} \\
+        \cmidrule(lr){2-3} \cmidrule(lr){4-5}
+        & {[\si{\kJ\per\mole}]} & {[\si{\kilo\calorie\per\mole}]} & {[\si{\kJ\per\mole}]} & {[\si{\kilo\calorie\per\mole}]} \\
+        \midrule
+        MV & -470.98 & -112.57 & -472.53 & -112.94 \\
+        2 & -297.16 & -71.02 & -273.63 & -65.40 \\
+        3 & -277.03 & -66.21 & -273.63 & -65.40 \\
+        4 & -277.03 & -66.21 & -278.61 & -66.59 \\
+        5 & -265.94 & -63.56 & -267.53 & -63.94 \\
+        M & -270.51 & -64.65 & -270.12 & -64.56 \\
+        \bottomrule
+    \end{tabular}
+\end{center}
+
+\Cref{tab:heat-of-formation} shows that the heats of formation of the fuel and radicals 3, 4, 5, and
+M are quite similar between the two mechanisms. However, the heat of formation of the second
+radical, the one closest to the methyl ester group, has a significantly lower heat of formation in
+the model by \citet{Dievart2013} than in the RMG model. Note that it is expected that the second
+radical will be somewhat more stable than the other radicals, due to the influence of the methyl
+ester group on the adjacent carbon atom.
+
+This difference in heats of formation affects the pathways that consume the fuel. By conducting a
+reaction pathway analysis to determine which radicals are formed from the breakdown of the fuel, we
+can analyze the proportion of each radical formed as the fuel breaks down during the autoignition
+process. The following analysis is conducted for a constant volume, adiabatic simulation with
+initial temperature and pressure of \SI{700}{\K}, \SI{30}{\bar}, and for the stoichiometric
+equivalence ratio. The rates of production of the species have been integrated until the time of
+\SI{20}{\percent} fuel consumption. The results of this analysis are shown in \cref{tab:mv-radicals}
+for the two models. The percentages shown in the \cref{tab:mv-radicals} are the percent of the fuel
+destroyed to form a particular fuel radical by all the reactions that can form that radical, and the
+radicals are labeled according to the convention in \cref{fig:mv-structure}.
 
@@
     \label{tab:mv-radicals}
-    \begin{tabular}{cSSSS}
+    \begin{tabular}{cSSS}
         \toprule
-        Radical Site & {\citet{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} & & \\
+        Radical Site & {\citet{Dievart2013} [\si{\percent}]} & {RMG Model [\si{\percent}]} & {RMG switched [\si{\percent}]}\\
         \midrule
-        2 & 29.3 & 12.5 & 27.7 & 15.6\\
-        3 & 17.5 & 12.2 & 17.8 & 20.1\\
-        4 & 17.5 & 50.6 & 17.8 & 23.2 \\
-        5 & 9.4 & 3.9 & 10.1 & 2.7 \\
-        M & 26.3 & 20.8 & 26.6 & 38.4\\
+        2 & 29.2 & 12.5 & 11.0 \\
+        3 & 17.5 & 12.2 & 11.1 \\
+        4 & 17.5 & 50.6 & 56.6 \\
+        5 & 9.5 & 3.9 & 4.3 \\
+        M & 26.3 & 20.8 & 16.9 \\
         \bottomrule
@@
 \end{center}
+
+At the relatively low temperature and high pressure condition of this analysis, all of the fuel is
+destroyed by \ce{H}-atom abstractions to form the fuel radicals shown. It can be seen that the two
+models have quite different distributions of products from the first \ce{H}-abstraction reactions.
+The model of \cite{Dievart2013} predicts that \ce{H}-abstraction from the second carbon is the most
+prevalent, while the RMG model predicts that the radical on the fourth carbon in the chain will be
+primarily formed. This is in line with the heats of formation in \cref{tab:heat-of-formation}, where
+the most stable radical (i.e., the radical with the smallest heat of formation) is most likely to be
+formed in each model.
+
+To further compare the models with each other, the NASA polynomials representing the thermodynamic
+properties of MV and the 5 fuel radicals from the model of \citet{Dievart2013} and are used to
+replace the equivalent molecules in the RMG model. The results of a path analysis at the same
+condition as the other analysis is shown in \cref{tab:mv-radicals} in the ``RMG switch'' column.
+This analysis shows that the radical on the fourth carbon atom is still the most prevalant, despite
+changing the heats of formation of the fuel and its radicals.
+
+Taken together, these results show that the errors in a given model cannot be attributed to a single
+source. There is a strong interaction between the thermodynamics of the species and the kinetics of
+the reactions, requiring further detailed study of the methyl ester system to accurately predict the
+low temperature ignition delays of methyl valerate.
 
@@
 literature model, it is still unable to predict the experimental ignition delays satisfactorily.
-Possible reasons for the discrepancy include missing reaction pathways, incorrect rate estimates,
-and incorrect thermodynamic property estimates. Future work will include investigation of the
-discrepancies between models and experiments to further understand the autoignition kinetics of
-methyl valerate.
+Both models predict an NTC region of the ignition delay under conditions where none was found in the
+experiments, and fail to predict the NTC region of ignition delay that is present in the
+experiments. Possible reasons for the discrepancy include missing reaction pathways, incorrect rate
+estimates, and incorrect thermodynamic property estimates. Comparative analysis of the two models
+failed to identify a single source of the error, and further detailed studies are required to
+improve predictions of the ignition delay at these engine-relevant conditions.
 

### 2017-07-10 methyl-valerate.tex
Kamal updates

@@
 \begin{frontmatter}
-\title{ Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }
+\title{ Experiments and Modeling of Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }
 
@@
 \begin{abstract} % not to exceed 200 words
-    Methyl valerate (\ce{C6H12O2}, methyl pentanoate) is a methyl ester and a
-    relevant surrogate component for biodiesel. In this work, we present
-    ignition delays of methyl valerate measured using a rapid compression
-    machine at a range of engine-relevant temperature, pressure, and equivalence
-    ratio conditions. The conditions we have studied include equivalence ratios
-    from \numrange{0.25}{2.0}, temperatures between \SIlist{680;1050}{\K}, and
-    pressures of \SIlist{15;30}{\bar}. The ignition delay data demonstrate a
-    negative temperature coefficient region in the temperature range of
-    \SIrange[range-phrase={--}]{720}{800}{\K} for both \(\phi=2.0\),
-    \SI{15}{\bar} and \(\phi=1.0\), \SI{30}{\bar}, with two-stage ignition
-    apparent over the narrower temperature ranges of
-    \SIrange[range-phrase={--}]{720}{760}{\K} for the lower pressure and
-    \SIrange[range-phrase={--}]{740}{760}{\K} at the higher pressure. In
-    addition, the experimental ignition delay data are compared with simulations
-    using an existing chemical kinetic model from the literature. The
-    simulations with the literature model under-predict the data by factors
-    between \numlist{2;10} over the entire range of the experimental data. To
-    help determine the possible reasons for the discrepancy between simulations
-    and experiments, a new chemical kinetic model is developed using the
-    Reaction Mechanism Generator (RMG) software. The agreement between the
-    experimental data and the RMG model is improved but still not satisfactory.
-    Directions for future improvement of the methyl valerate model are
-    discussed.
+    Methyl valerate (\ce{C6H12O2}, methyl pentanoate) is a methyl ester and a relevant surrogate
+    component for biodiesel. In this work, we present ignition delays of methyl valerate measured
+    using a rapid compression machine at a range of engine-relevant temperature, pressure, and
+    equivalence ratio conditions. The conditions we have studied include equivalence ratios from
+    \numrange{0.25}{2.0}, temperatures between \SIlist{680;1050}{\K}, and pressures of
+    \SIlist{15;30}{\bar}. The ignition delay data demonstrate a negative temperature coefficient
+    region in the temperature range of \SIrange[range-phrase={--}]{720}{800}{\K} for both
+    \(\phi=2.0\), \SI{15}{\bar} and \(\phi=1.0\), \SI{30}{\bar}, with two-stage ignition apparent
+    over the narrower temperature ranges of \SIrange[range-phrase={--}]{720}{760}{\K} for the lower
+    pressure and \SIrange[range-phrase={--}]{740}{760}{\K} at the higher pressure. In addition, the
+    experimental ignition delay data are compared with simulations using an existing chemical
+    kinetic model from the literature. The simulations with the literature model under-predict the
+    data by factors between \numlist{2;10} over the entire range of the experimental data. In
+    addition, a new chemical kinetic model is developed using the Reaction Mechanism Generator (RMG)
+    software. The agreement between the experimental data and the RMG model is also not
+    satisfactory. To help determine the possible reasons for the disagreement, a path analysis of
+    both models is completed. It is found that improvements to both the reaction pathways and
+    thermodynamic properties is required. Further directions for future improvement of the methyl
+    valerate model are discussed.
 \end{abstract}
@@
 kinetic mechanisms challenging~\cite{Lai2011}, it is desired to study their combustion chemistry by
-studying simpler molecules.
+studying simpler methyl ester molecules.
 
@@
 (\(T_C\) and \(P_C\) respectively), are independently changed by varying the overall compression
-ratio, initial pressure \((P_0)\), and initial temperature \((T_0)\) of the experiments. The piston
-in the reaction chamber is machined with a specially designed crevice to suppress the roll-up vortex
-effect and promote homogeneous conditions in the reactor during and after
-compression~\cite{Mittal2006}.
+ratio, initial pressure \((P_0)\), initial temperature \((T_0)\), and specific heat ratio of the
+experiments. The piston in the reaction chamber is machined with a specially designed crevice to
+suppress the roll-up vortex effect and promote homogeneous conditions in the reactor during and
+after compression~\cite{Mittal2006}.
 
@@
 
-Mixtures are prepared in stainless steel mixing tanks, approximately \SI{17}{\L} and \SI{15}{\L} in
-size. The proportions of reactants in the mixture are determined by specifying the absolute mass of
-the fuel, the equivalence ratio \((\phi)\), and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer.
-Mixtures are made by first vacuuming the mixing tanks to an ultimate pressure less than
-\SI{5}{\torr}. Since MV is a liquid with a relatively small vapor pressure at room temperature and
-pressure, it is measured gravimetrically in a syringe to within \SI{0.01}{\g} of the specified
-value. The fuel is injected into the mixing tank through a septum. Proportions of \ce{O2}, \ce{Ar},
-and \ce{N2} are added manometrically at room temperature and the total pressure is measured by an
-Omega Engineering MMA type static pressure transducer. The same transducer is used to measure the
-pressure of the reactants prior to an experiment.
+Mixtures are prepared in stainless steel mixing tanks, \SI{17}{\L} and \SI{15}{\L} in size. The
+proportions of reactants in the mixture are determined by specifying the absolute mass of the fuel,
+the equivalence ratio \((\phi)\), and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer. Mixtures are
+made by first vacuuming the mixing tanks to an ultimate pressure less than \SI{5}{\torr}. Since MV
+is a liquid with a relatively small vapor pressure at room temperature and pressure, it is measured
+gravimetrically to within \SI{0.01}{\g} of the specified value. The fuel is injected into the mixing
+tank through a septum. Proportions of \ce{O2}, \ce{Ar}, and \ce{N2} are added manometrically at room
+temperature and the total pressure is measured by an Omega Engineering MMA100V10T2D0T4A6 type static
+pressure transducer. The same transducer is used to measure the pressure of the reactants prior to
+an experiment.
 
@@
 compressed pressure of \SI{30}{\bar}. Note that $\phi=2.0$ results were not collected for
-\SI{30}{\bar}, so there are no red data points in \cref{fig:ignition-delays}b.
+\SI{30}{\bar}, so there are no red data points in \cref{fig:ignition-delays}b. All the data are
+available in comma-separated value files and ChemKED-format \cite{Weber2017} files in the
+Supplementary Material.
 
@@
 field in the RCM reaction chamber was uniform for approximately \SI{100}{\ms} after the EOC, and our
-measured ignition delay at \SI{15}{\bar}, \(\phi=1.0\) \SI{828}{\K} is \SI{92.14}{\ms}.
+measured ignition delay at \SI{15}{\bar}, \(\phi=1.0\), and \SI{828}{\K} is \SI{92.14}{\ms}.
 
@@
 The second source of error may be incorrect reaction rate parameters, either because the rates are
-specified incorrectly in the model (e.g., typos) or because the rates are not well estimated by the
+specified incorrectly in the model or because the rates are not well estimated by the
 typical analogy based-rules. It should be noted that errors of this type may affect the model
@@
 for the two models. The percentages shown in the \cref{tab:mv-radicals} are the percent of the fuel
-destroyed to form a particular fuel radical by all the reactions that can form that radical, and the
+consumed to form a particular fuel radical by all the reactions that can form that radical, and the
 radicals are labeled according to the convention in \cref{fig:mv-structure}.
@@
     \captionsetup{type=table}
-    \caption{Percent of MV destroyed to form fuel radical species with a hydrogen atom missing at
+    \caption{Percent of MV consumed to form fuel radical species with a hydrogen atom missing at
     the location indicated in the first column and \cref{fig:mv-structure}}
@@
 At the relatively low temperature and high pressure condition of this analysis, all of the fuel is
-destroyed by \ce{H}-atom abstractions to form the fuel radicals shown. It can be seen that the two
+consumed by \ce{H}-atom abstractions to form the fuel radicals shown. It can be seen that the two
 models have quite different distributions of products from the first \ce{H}-abstraction reactions.
@@
 replace the equivalent molecules in the RMG model. The results of a path analysis at the same
-condition as the other analysis is shown in \cref{tab:mv-radicals} in the ``RMG switch'' column.
+condition as the other analysis is shown in \cref{tab:mv-radicals} in the ``RMG switched'' column.
 This analysis shows that the radical on the fourth carbon atom is still the most prevalant, despite

### 2017-07-14 methyl-valerate.tex
Updates

@@
 data. The model does not predict the NTC region found experimentally for the \(\phi = 1.0\)
-experiments, but the agreement improves as the temperature decreases.
+experiments, but the agreement improves as the temperature decreases. Interestingly, two-stage
+ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
+\cref{fig:simulation-comparison}. However, the first-stage ignition delays are
+\SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
+\cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
+delay.
 
@@
 \citet{Dievart2013} model for the \SI{15}{\bar}, \(\phi =\) \numlist{0.25; 0.5} data. Moreover, the
-RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar},
-\(\phi = 2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays, and
-does not predict the NTC region found experimentally. As before, the trend at \SI{30}{\bar},
-\(\phi=1.0\) is similar to the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the
-low-temperature ignition delays and does not predict the experimental NTC region.
+RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar}, \(\phi =
+2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays, and does not
+predict the NTC region found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is
+similar to the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature
+ignition delays and does not predict the experimental NTC region. Finally, as in the
+\citet{Dievart2013} model, two-stage ignition is predicted for all of the \(\phi=1.0\) and
+\(\phi=2.0\) data shown in \cref{fig:simulation-comparison}. However, the first-stage ignition
+delays are \SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
+\cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
+delay.
 

### 2017-07-14 methyl-valerate.tex
Updates

@@
-\documentclass[letterpaper, review]{elsarticle}
+\documentclass[letterpaper, review, sort&compress]{elsarticle}
 
@@
 
-\DeclareSIUnit\torr{torr}
+\DeclareSIUnit\torr{Torr}
 \DeclareSIUnit\atm{atm}
@@
 \author[uconn]{Bryan W.\ Weber\corref{cor1}}
-\author[uconn]{Justin Bunnell}
+\author[uconn]{Justin A.\ Bunnell}
 \author[idaho]{Kamal Kumar}
@@
 \begin{abstract} % not to exceed 200 words
+
     Methyl valerate (\ce{C6H12O2}, methyl pentanoate) is a methyl ester and a relevant surrogate
@@
     using a rapid compression machine at a range of engine-relevant temperature, pressure, and
-    equivalence ratio conditions. The conditions we have studied include equivalence ratios from
-    \numrange{0.25}{2.0}, temperatures between \SIlist{680;1050}{\K}, and pressures of
-    \SIlist{15;30}{\bar}. The ignition delay data demonstrate a negative temperature coefficient
+    equivalence ratio conditions. The conditions we have studied include equivalence ratios
+    \((\phi)\) from \numrange{0.25}{2.0}, temperatures between \SIlist{680;1050}{\K}, and pressures
+    of \SIlist{15;30}{\bar}. The ignition delay data demonstrate a negative temperature coefficient
     region in the temperature range of \SIrange[range-phrase={--}]{720}{800}{\K} for both
     \(\phi=2.0\), \SI{15}{\bar} and \(\phi=1.0\), \SI{30}{\bar}, with two-stage ignition apparent
-    over the narrower temperature ranges of \SIrange[range-phrase={--}]{720}{760}{\K} for the lower
-    pressure and \SIrange[range-phrase={--}]{740}{760}{\K} at the higher pressure. In addition, the
+    over the narrower temperature ranges of \SIrange[range-phrase={--}]{720}{760}{\K} for
+    \SI{15}{\bar} and \SIrange[range-phrase={--}]{740}{760}{\K} at \SI{30}{\bar}. In addition, the
     experimental ignition delay data are compared with simulations using an existing chemical
@@
     both models is completed. It is found that improvements to both the reaction pathways and
-    thermodynamic properties is required. Further directions for future improvement of the methyl
+    thermodynamic properties are required. Further directions for future improvement of the methyl
     valerate model are discussed.
+
 \end{abstract}
@@
 \begin{keyword}
-    chemical kinetics\sep rapid compression machine\sep autoignition\sep methyl ester
+    chemical kinetics\sep rapid compression machine\sep autoignition\sep methyl ester\sep methyl valerate\sep methyl pentanoate
 \end{keyword}
@@
 friendliness of fuels. This is due to its renewability when produced from sustainable agricultural
-crops and its ability to reduce emissions relative to conventionally fueled
-engines~\cite{Hoekman2012}. Biodiesel typically consists of long-chain methyl ester molecules, with
-typical compositions of \ce{C14} to \ce{C20}~\cite{Hoekman2012}. Recognizing that the large
-molecular size of the methyl esters within biodiesel fuel makes creating and using detailed chemical
-kinetic mechanisms challenging~\cite{Lai2011}, it is desired to study their combustion chemistry by
-studying simpler methyl ester molecules.
+crops and its ability to reduce emissions relative to petroleum-derived fuels~\cite{Hoekman2012}.
+Biodiesel typically consists of long-chain methyl ester molecules, with typical compositions of
+\ce{C14} to \ce{C20}~\cite{Hoekman2012}. Recognizing that the large molecular size of the methyl
+esters within biodiesel fuel makes creating and using detailed chemical kinetic models
+challenging~\cite{Lai2011}, it is desired to study their combustion chemistry by studying simpler
+methyl ester molecules.
 
@@
 
-Larger methyl esters such as methyl valerate (MV, \ce{C6H12O2}, methyl pentanoate) have also been
-studied as possible biodiesel surrogates. \citet{Hadj-Ali2009} used a rapid compression machine
-(RCM) to study the autoignition of several methyl esters including MV. Although MV exhibited
-two-stage ignition in this study, little additional research has been done on its oxidation.
-\citet{Korobeinichev2015} studied MV in premixed laminar flames and extended a detailed high
-temperature chemical kinetic model to include MV and methyl hexanoate. \citet{Dmitriev2015} added
-MV to n-heptane/toluene fuel blends to determine the resulting intermediate species in premixed
-flames using a flat burner at \SI{1}{atm} and an equivalence ratio of 1.75. The addition of MV
-helped reduce soot forming intermediates including benzene, cyclopentadienyl, acetylene, propargyl,
-and vinylacetylene~\cite{Dmitriev2015}. \citet{Hayes2009} computationally examined the peroxy
-radical isomerization reactions for MV to better understand the low temperature reaction pathways.
-Finally, \citet{Dievart2013} used diffusion flames in the counterflow configuration to determine
-extinction limits for a number of methyl esters, including MV, and validated a detailed kinetic
-model with the experimental data.
+Methyl esters larger than MB, such as methyl valerate (MV, \ce{C6H12O2}, methyl pentanoate), have
+also been studied as possible biodiesel surrogates. \citet{Hadj-Ali2009} used a rapid compression
+machine (RCM) to study the autoignition of several methyl esters including MV. Although MV exhibited
+two-stage ignition in this study, little additional research has been done on its low-temperature
+chemistry. \citet{Korobeinichev2015} studied MV in premixed laminar flames and extended a detailed
+high temperature chemical kinetic model to include MV and methyl hexanoate. \citet{Dmitriev2015}
+added MV to n-heptane/toluene fuel blends to determine the resulting intermediate species in
+premixed flames using a flat burner at \SI{1}{atm} and an equivalence ratio of 1.75. The addition of
+MV helped reduce soot forming intermediates including benzene, cyclopentadienyl, acetylene,
+propargyl, and vinylacetylene~\cite{Dmitriev2015}. \citet{Hayes2009} computationally examined the
+peroxy radical isomerization reactions for MV to better understand the low temperature reaction
+pathways. Finally, \citet{Dievart2013} used diffusion flames in the counterflow configuration to
+determine extinction limits for a number of methyl esters, including MV, and validated a detailed
+kinetic model with the experimental data.
 
 This work provides additional data for the autoignition of MV. Data is collected in a RCM under
-engine relevant conditions spanning from \SIrange{15}{30}{\bar}, equivalence ratios from
-\numrange{0.25}{2.0}, and temperatures from \SIrange{682}{1048}{\K}. The NTC region of MV is mapped
+engine relevant conditions spanning from \SIrange{15}{30}{\bar}, equivalence ratios \((\phi)\) from
+\numrange{0.25}{2.0}, and temperatures from \SIrange{680}{1050}{\K}. The NTC region of MV is mapped
 out to provide additional information on the fidelity of using MV as a biodiesel surrogate.
@@
 either 50 kHz or 100 kHz by a LabView VI and processed by a Python package called
-UConnRCMPy~\cite{Weber2016a}. Version 3.0.0 of UConnRCMPy~\cite{uconnrcmpy}, 3.6.1 of Python, 2.3.0
+UConnRCMPy~\cite{Weber2016a}. Version 3.0.5 of UConnRCMPy~\cite{uconnrcmpy}, 3.6 of Python, 2.3.0
 of Cantera~\cite{cantera}, 1.13 of NumPy~\cite{vanderWalt2011}, 0.19.0 of SciPy~\cite{Jones2001},
-and 2.0.1 of Matplotlib~\cite{Hunter2007} were used in the analysis in this paper.
+and 2.0.1 of Matplotlib~\cite{Hunter2007} are used in the analysis in this paper.
+
+\begin{center}
+    \captionsetup{type=figure}
+    \includegraphics{figures/ignition_delay_definition.pdf}
+    \caption{Definition of the ignition delays used in this work. The
+    experiment in this figure is conducted for a \(\phi=2.0\) mixture with
+    \mbox{\(\ce{Ar}/(\ce{N2}+\ce{Ar})=0.5\)}, \(P_0=\SI{0.7806}{\bar}\),
+    \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.92}{\bar}\), \(T_C=\SI{720}{\K}\),
+    \(\tau=\SI{27.56\pm0.89}{\ms}\), and \(\tau_1=\SI{16.60\pm0.46}{\ms}\).}
+    \label{fig:ign-delay-def}
+\end{center}
 
@@
 random scatter present in the data, the standard deviation \((\sigma)\) of the ignition delays from
-the runs at a given condition is computed. In all cases, \(\sigma\) is less than \SI{10}{\percent}
-of the mean value of the overall ignition delay.
+the runs at a given condition is computed. Typically, \(\sigma\) is less than \SI{10}{\percent} of
+the mean values of the overall ignition delay \((\tau)\) and the first stage ignition delay
+\((\tau_1)\).
 
-\begin{center}
-    \captionsetup{type=figure}
-    \includegraphics{figures/ignition_delay_definition.pdf}
-    \caption{Definition of the ignition delays used in this work. The
-    experiment in this figure was conducted for a \(\phi=2.0\) mixture with
-    \mbox{\(\ce{Ar}/(\ce{N2}+\ce{AR})=0.5\)}, \(P_0=\SI{0.7806}{\bar}\),
-    \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.92}{\bar}\), \(T_C=\SI{720}{\K}\),
-    \(\tau=\SI{27.56\pm0.89}{\ms}\), \(\tau_1=\SI{16.60\pm0.46}{\ms}\).}
-    \label{fig:ign-delay-def}
-\end{center}
-
-In addition to the reactive experiments, non-reactive experiments are conducted to determine the
-influence of machine-specific behavior on the experimental conditions and permit the calculation of
-the EOC temperature via the isentropic relations between pressure and temperature~\cite{Lee1998}.
-The EOC temperature is calculated by the procedure described in \cref{sec:computational-methods}.
+In addition to the reactive experiments, non-reactive experiments are conducted by replacing \ce{O2}
+with \ce{N2} to determine the influence of machine-specific behavior on the experimental conditions
+and permit the calculation of the EOC temperature via the isentropic relations between pressure and
+temperature~\cite{Lee1998}. The EOC temperature is calculated by the procedure described in
+\cref{sec:computational-methods}.
 
 The mixtures considered in this study are shown in \cref{tab:mixtures}. Four equivalence ratios of
-MV in ``air'' are considered. The ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer is varied to adjust
-the temperatures reached at the EOC for a given mixture. Two \(P_C\) conditions are studied in this
-work, \SIlist{15;30}{\bar}, representing engine-relevant conditions. For the \(\phi=2.0\) condition,
-only \(P_C = \SI{15}{\bar}\) is considered because we could not achieve \(T_C\) values low enough
-that the ignition was long enough to be measured in our apparatus (the typical lower limit of
-ignition delay on the present RCM is approximately \SI{5}{\ms}).
-
-Mixtures are prepared in stainless steel mixing tanks, \SI{17}{\L} and \SI{15}{\L} in size. The
-proportions of reactants in the mixture are determined by specifying the absolute mass of the fuel,
-the equivalence ratio \((\phi)\), and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer. Mixtures are
-made by first vacuuming the mixing tanks to an ultimate pressure less than \SI{5}{\torr}. Since MV
-is a liquid with a relatively small vapor pressure at room temperature and pressure, it is measured
-gravimetrically to within \SI{0.01}{\g} of the specified value. The fuel is injected into the mixing
-tank through a septum. Proportions of \ce{O2}, \ce{Ar}, and \ce{N2} are added manometrically at room
-temperature and the total pressure is measured by an Omega Engineering MMA100V10T2D0T4A6 type static
-pressure transducer. The same transducer is used to measure the pressure of the reactants prior to
-an experiment.
+MV in ``air'' are considered. While \ce{O2} is kept at \SI{21}{\percent} by mole in the oxidizer,
+the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer is varied to adjust the temperatures reached at the
+EOC. Two \(P_C\) conditions are studied in this work, \SIlist{15;30}{\bar}, representing
+engine-relevant conditions. For the \(\phi=2.0\) condition, only \(P_C = \SI{15}{\bar}\) is
+considered because we could not achieve \(T_C\) values low enough that the ignition during the
+compression stroke can be prevented.
 
@@
         0.25 & 0.0065 & 0.2087 & 0.7848 & 0.0000 & 1.0 \\
-        0.5 & 0.0130 & 0.2074 & 0.7798 & 0.0000 & 1.0 \\
+        0.5 & 0.0130 & 0.2074 & 0.7796 & 0.0000 & 1.0 \\
         1.0 & 0.0256 & 0.2047 & 0.7697 & 0.0000 & 1.0 \\
@@
 
+Mixtures are prepared in stainless steel mixing tanks, \SI{17}{\L} and \SI{15}{\L} in size. The
+proportions of reactants in the mixture are determined by specifying the absolute mass of the fuel,
+the equivalence ratio, and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer. Mixtures are
+made by first vacuuming the mixing tanks to an ultimate pressure less than \SI{5}{\torr}. Since MV
+is a liquid with a relatively small vapor pressure at room temperature and pressure, it is measured
+gravimetrically to within \SI{0.01}{\g} of the specified value. The fuel is injected into the mixing
+tank through a septum. Proportions of \ce{O2}, \ce{Ar}, and \ce{N2} are added manometrically at room
+temperature and the total pressure is measured by an Omega Engineering MMA100V10T2D0T4A6 type static
+pressure transducer. The same transducer is used to measure the pressure of the reactants prior to
+an experiment.
+
 The RCM is equipped with heaters to control the initial temperature of the mixture. After filling in
@@
 so the reactants are well mixed for the duration of the experiments.
+
+\begin{center}
+    \captionsetup{type=figure}
+    \input{figures/vapor_pressure.pgf}
+    \caption{Saturated vapor pressure of MV as a function of temperature,
+    plotted using the Antoine equation, \cref{eq:antoine}, with
+    \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
+    \label{fig:vapor-pressure}
+\end{center}
 
@@
 %
-is used to model the saturated vapor pressure of MV as a function of temperature, where \(A\),
-\(B\), and \(C\) are substance-specific coefficients. Coefficients for \cref{eq:antoine} are given
-in the literature by~\citet{Ortega2003}, \citet{Camacho2007}, and~\citet{Stephenson1987}.
-Unfortunately, the values of the coefficients are different among all three authors and, more
-importantly, the temperature ranges provided in those three fits do not cover the entire range of
-interest for this study. Therefore, coefficients for use in \cref{eq:antoine} are determined in this
-work by least squares fitting of the data of \citet{Ortega2003}, \citet{vanGenderen2002}, and
-\citet{Verevkin2008} using the \verb|curve_fit()| function of SciPy~\cite{Jones2001} version 0.19.0.
-\Cref{fig:vapor-pressure} shows that the coefficients fit with this procedure give good agreement
-with the experimental data; values for the coefficients computed in this work and in the literature
-works are given in \cref{tab:antoine}. The data used to calculate the coefficients are provided in
-the Supplementary Material.
-
-\begin{center}
-    \captionsetup{type=figure}
-    \input{figures/vapor_pressure.pgf}
-    \caption{Saturated vapor pressure of MV as a function of temperature,
-    plotted using the Antoine equation, \cref{eq:antoine}, with
-    \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
-    \label{fig:vapor-pressure}
-\end{center}
+is used to model the saturated vapor pressure of MV as a function of temperature \((T)\), where
+\(A\), \(B\), and \(C\) are substance-specific coefficients. Coefficients for \cref{eq:antoine} are
+given in the literature by~\citet{Ortega2003}, \citet{Camacho2007}, and~\citet{Stephenson1987}.
+Unfortunately, the values of the coefficients are different among all three references. Therefore,
+coefficients for use in \cref{eq:antoine} are determined in this work by least squares fitting of
+the data of \citet{Ortega2003}, \citet{vanGenderen2002}, and \citet{Verevkin2008} using the
+\verb|curve_fit()| function of SciPy~\cite{Jones2001} version 0.19.0. \Cref{fig:vapor-pressure}
+shows that the coefficients fitted with this procedure give good agreement with the experimental
+data; values for the coefficients computed in this work and in the literature works are given in
+\cref{tab:antoine}. The data used to calculate the coefficients are provided in the Supplementary
+Material.
 
@@
 simulation conducted in an \verb|IdealGasReactor| in Cantera~\cite{cantera} using the CVODES
-solver from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delay from the simulations is
+solver from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delays from the simulations are
 defined in the same manner as in the experiments. The time derivative of the pressure in the
-simulations is computed by second order Lagrange polynomials, as discussed by \citet{Chapra2010}.
+simulation is computed by second order Lagrange polynomials, as discussed by \citet{Chapra2010}.
 
 To the best of our knowledge, there are three mechanisms for MV combustion available in the
-literature. The first two, by \cite{Korobeinichev2015} and \cite{Dmitriev2015}, were developed to
+literature. The first two, by \citet{Korobeinichev2015} and \citet{Dmitriev2015}, were developed to
 simulate flames, and do not include the low-temperature chemistry necessary to simulate the
-conditions in these experiments. The third model was developed by \cite{Dievart2013} and includes
+conditions in these experiments. The third model was developed by \citet{Dievart2013} and includes
 low-temperature chemistry of MV, although it was only validated by comparison with flame extinction
-limits. In converting this mechanism for use in Cantera, we found that there were many species in the
-thermodynamic database with multiple data entries. For most of these species the thermodynamic data
-is identical. However, some species are not exact duplicates. For these species, it is not clear
-from the thermodynamic database file which data set should be preferred. Since Cantera (and CHEMKIN)
-choose the first instance of a duplicate species to be used, we retained the first entry for all
-duplicated species. The detailed \cite{Dievart2013} model includes 1105 species and 7141 reactions,
-and the CHEMKIN and Cantera formatted input files are available in the Supplementary Material.
+limits. In converting this mechanism for use in Cantera, we found that there were many species in
+the thermodynamic database with multiple data entries. For most of these species the thermodynamic
+data is identical. However, some species are not exact duplicates. For these species, it is not
+clear from the thermodynamic database file which data set should be preferred. Since Cantera (and
+CHEMKIN) choose the first instance of a duplicate species to be used, we retained the first entry
+for all duplicated species. The detailed model of \citet{Dievart2013} includes 1105 species and 7141
+reactions, and the Cantera formatted input file is available in the Supplementary Material.
 
@@
 one fuel (MV), but the number of reactions is substantially higher. The input file used to generate
-the model is available in the Supplementary Material.
+the model is available in the Supplementary Material. In addition, the CHEMKIN and Cantera formatted
+input files for the RMG model are available in the Supplementary Material.
 
@@
 denote the overall ignition delay and hollow markers indicate the first-stage ignition delay.
-Vertical error bars are drawn on the symbols to represent the uncertainty in the ignition delay; for
-many of the experiments, the uncertainty is approximately the same size as the data point, so the
-error bar is hidden. Horizontal error bars are shown on the first and last points of each
-equivalence ratio indicating the estimated uncertainty in the EOC temperature of
+Vertical error bars are drawn on the symbols to represent the \(2\sigma\) uncertainty in the
+ignition delay; for many of the experiments, the uncertainty is approximately the same size as the
+data point, so the error bar is hidden. Horizontal error bars are shown on the first and last points
+of each equivalence ratio indicating the estimated uncertainty in the EOC temperature of
 $\pm\SI{1}{\percent}$~\cite{Weber2015}. \cref{fig:ignition-delays}a shows the results for a
@@
 compressed pressure of \SI{30}{\bar}. Note that $\phi=2.0$ results were not collected for
-\SI{30}{\bar}, so there are no red data points in \cref{fig:ignition-delays}b. All the data are
-available in comma-separated value files and ChemKED-format \cite{Weber2017} files in the
+\SI{30}{\bar}, so there are no red triangle data points in \cref{fig:ignition-delays}b. All the data
+are available in comma-separated value files and ChemKED-format \cite{Weber2017} files in the
 Supplementary Material.
@@
     \includegraphics{figures/ignition-delays.pdf}
-    \caption{Ignition delays of MV as a function of inverse temperature. Filled
-    points are the overall ignition delays and hollow points are the first stage
-    ignition delays. a) \SI{15}{\bar}. b) \SI{30}{\bar}}
+    \caption{Ignition delays of MV as a function of inverse temperature for varying equivalence
+    ratios. Filled points are the overall ignition delays and hollow points are the first stage
+    ignition delays. a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
     \label{fig:ignition-delays}
@@
 this temperature, the autoignition occurred in two stages with a clearly identified cool flame
-event.''\cite{Hadj-Ali2009} However, we do not find two stage ignition for the similar pressure of
+event.'' However, we do not find two stage ignition for the similar pressure of
 \(P_C = \SI{15}{\bar}\) in this study. We note that the stated temperature of the experiment from
@@
 gasoline composed of iso-alkanes, n-alkanes, cyclo-alkanes, aromatics, and olefins.
-\citet{Kukkadapu2012} attributed the NTC shift to the reactions between the hydroperoxyalkyl radical
-(\ce{QOOH}) and \ce{O2} becoming more dominant than the unimolecular decomposition of \ce{QOOH} at
-higher pressures. Similar trends could occur for the hydroperoxyalkyl radicals of MV.
+\citet{Kukkadapu2012} attributed the shift of the NTC region to the reactions between the
+hydroperoxyalkyl radical (\ce{QOOH}) and \ce{O2} becoming more dominant than the unimolecular
+decomposition of \ce{QOOH} at higher pressures. Similar trends could occur for the hydroperoxyalkyl
+radicals of MV.
 
@@
 n-pentane to be between \SIlist{760;910}{\K} at pressures near \SI{10}{\atm}. As we will compare
-with our MV data at \SI{30}{\bar}, we note that increasing the pressure tends to shift the NTC
-to higher temperatures, as mentioned previously \cite{Kukkadapu2012}. \citet{Heufer2013} found NTC
+with our MV data at \SI{30}{\bar}, we note that increasing the pressure tends to shift the NTC to
+higher temperatures, as mentioned previously \cite{Kukkadapu2012}. \citet{Heufer2013} found NTC
 behavior for n-pentanol in the range of \SIrange{770}{900}{\K} at \SI{30}{\bar}. In this study, we
 find the NTC window for MV at \SI{30}{\bar} to be between \SIlist{720;800}{\K}. Therefore, it
-appears that the methyl ester functional group shifts the NTC range to lower temperature as compared
-to alkanes and alcohols with similar alkyl chain lengths. This result was also noted by
-\citet{Hadj-Ali2009} for MHEX as the fuel.
+appears that the methyl ester functional group causes the NTC range to occur at lower temperature as
+compared to alkanes and alcohols with similar alkyl chain lengths. This result was also noted by
+\citet{Hadj-Ali2009} for methyl hexanoate as the fuel.
 
@@
 of the pressure, so no first-stage ignition delay is defined for this case. It can also be seen in
-\cref{fig:pressure-traces}a that the \SI{775}{\K} case shows some heat release prior to ignition,
+\cref{fig:pressure-traces}a that the \SI{774}{\K} case shows some heat release prior to ignition,
 although again there is only one peak in the time derivative of the pressure. Furthermore, the heat
-release at \SI{775}{\K} appears to be more gradual than at the lowest temperature.
+release at \SI{774}{\K} appears to be more gradual than at \SI{700}{\K}.
 
@@
     \caption{Selected pressure traces around the NTC region of ignition delay.
-    a) \(\phi=1.0\) b) \(\phi=2.0\)}
+    a) \(\phi=1.0,\ P_C=\SI{30}{\bar}\) b) \(\phi=2.0,\ P_C=\SI{15}{\bar}\)}
     \label{fig:pressure-traces}
@@
 agreement seems to be improving as the temperature is decreased. Comparing the \citet{Dievart2013}
-model to the data at \SI{30}{\bar}, we find a similar trend as the \(\phi=2.0,\ P_C=\SI{15}{\bar}\)
-data. The model does not predict the NTC region found experimentally for the \(\phi = 1.0\)
-experiments, but the agreement improves as the temperature decreases. Interestingly, two-stage
-ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
+model to the stoichiometric data at \SI{30}{\bar}, we find a similar trend as the \(\phi=2.0,\
+P_C=\SI{15}{\bar}\) data. The model does not predict the NTC region found experimentally for the
+\(\phi = 1.0\) experiments, but the agreement improves as the temperature decreases. Interestingly,
+two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
 \cref{fig:simulation-comparison}. However, the first-stage ignition delays are
@@
 \cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
-delay.
+delay. While the model of \citet{Dievart2013} over-predicts the first-stage ignition delay, it also
+over-predicts the first-stage pressure rise, thereby leading to rapid overall ignition.
 
@@
 \cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
-delay.
+delay. Again, the over-prediction of the first-stage ignition delay and pressure rise requires
+further investigation and understanding of the complex low-temperature chemistry.
 
@@
         \toprule
-        \multirow{2}{*}{Radical Site} & \multicolumn{2}{c}{\citet{Dievart2013}} & \multicolumn{2}{c}{RMG Model} \\
+        \multirow{2}{*}{Radical Site} & \multicolumn{2}{c}{\citet{Dievart2013}} & \multicolumn{2}{c}{RMG Model (this work)} \\
         \cmidrule(lr){2-3} \cmidrule(lr){4-5}
@@
 
-Taken together, these results show that the errors in a given model cannot be attributed to a single
-source. There is a strong interaction between the thermodynamics of the species and the kinetics of
-the reactions, requiring further detailed study of the methyl ester system to accurately predict the
-low temperature ignition delays of methyl valerate.
+Taken together, these results show that the poor performance in a given model cannot be attributed
+to a single source. There is a strong interaction between the thermodynamics of the species and the
+kinetics of the reactions, requiring further detailed study of the methyl ester system to accurately
+predict the low temperature ignition delays of methyl valerate.
 
@@
 engine-relevant pressures, temperatures, and equivalence ratios. An NTC region of the ignition delay
-and two-stage ignition were recorded for pressures of \SI{15}{\bar} at \(\phi=2.0\) and
-\SI{30}{\bar} at \(\phi=1.0\). A detailed chemical kinetic model available in the literature was
-unable to reproduce the experimental results, so a new model was constructed using the Reaction
-Mechanism Generator software. Although the new model contains many more reactions than the
-literature model, it is still unable to predict the experimental ignition delays satisfactorily.
-Both models predict an NTC region of the ignition delay under conditions where none was found in the
-experiments, and fail to predict the NTC region of ignition delay that is present in the
-experiments. Possible reasons for the discrepancy include missing reaction pathways, incorrect rate
-estimates, and incorrect thermodynamic property estimates. Comparative analysis of the two models
-failed to identify a single source of the error, and further detailed studies are required to
-improve predictions of the ignition delay at these engine-relevant conditions.
+and two-stage ignition are recorded for pressures of \SI{15}{\bar} at \(\phi=2.0\) and \SI{30}{\bar}
+at \(\phi=1.0\). A detailed chemical kinetic model available in the literature is unable to
+reproduce the experimental results, so a new model is constructed using the Reaction Mechanism
+Generator software. Although the new model contains many more reactions than the literature model,
+it is still unable to predict the experimental ignition delays satisfactorily. Both models predict
+an NTC region of the ignition delay under conditions where none is found in the experiments, and
+fail to predict the NTC region of ignition delay that is present in the experiments. Possible
+reasons for the discrepancy include missing reaction pathways, incorrect rate estimates, and
+incorrect thermodynamic property estimates. Comparative analysis of the two models failed to
+identify a single source of the error, and further detailed studies are required to improve
+predictions of the ignition delay at these engine-relevant conditions.
 

### 2017-09-22 revision-1/response/response.tex
Revision 1 sent to Jackie

% !TEX program: xelatex
\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[numbers]{natbib}
\usepackage{xcite}
\externalcitedocument{../revised-paper/methyl-valerate}
\usepackage{xr}
\externaldocument{../revised-paper/methyl-valerate}
\usepackage[capitalize]{cleveref}
\crefname{appendix}{}{}
\usepackage{siunitx}
\sisetup{
    separate-uncertainty=true,
    per-mode=symbol
}
\usepackage{url}

\newenvironment{reviewer}{\vspace{0.5\baselineskip}\begingroup\itshape\textbf{Reviewer:}}{\endgroup\vspace{0.5\baselineskip}}
\newenvironment{response}{\vspace{0.5\baselineskip}\textbf{Our Response:}}{\vspace{0.5\baselineskip}}
\newcommand{\review}[1]{{\itshape#1}}
\newcommand{\respond}[1]{\textbf{Our Response:} #1}

\begin{document}

\noindent Experiments and Modeling of the Autoignition of Methyl Pentanoate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine \\
Ms.\ Ref.\ No.:  JFUE-D-17-02705

We would like to thank the reviewers for their thoughtful comments.
Below are detailed responses to each of the reviewers' comments.
In the attached marked revision, deleted text is marked in red and added text is marked in blue.

\vspace{\baselineskip}

\noindent \textbf{Reviewer \#1:}

\begin{reviewer}
%
    The authors study ignition of methyl pentanoate, a small methyl ester, in the current study. The
    data from the current study is helpful in establishing the reaction rates which can be
    systematically used for developing the kinetic models for large esters and hence very valuable
    for developing kinetic models of biodiesel. The study also shows the poor performance of the
    literature models in estimating the ignition delays underscoring the need for such novel
    experimental datasets and hence would recommend this work for publication. However, I think the
    manuscript needs to be revised before it can be published. Below are the specific comments which
    I think need to addressed before it can be published.
%
\end{reviewer}

\begin{response}
    Thank you for your comments. Detailed responses to each comment are below.
\end{response}

\begin{reviewer}
%
    Please improving the quality of figures. For eg: First stage and total ignition delay Symbols
    in Figs 5a \& 5b are very hard to differentiate. Figure 1 could also be improved.
%
\end{reviewer}

\begin{response}
    %
    We have increased the mark size on \cref{fig:simulation-comparison} to better distinguish the
    overall ignition delay and the first-stage ignition delay. On \cref{fig:ign-delay-def}, we made
    several changes to improve the clarity, including swapping the colors of reactive and
    non-reactive pressure traces, changing the non-reactive pressure to a dashed line style, and
    making the black lines thinner.
    %
\end{response}

\begin{reviewer}
    %
    Regarding the statement "The NTC region of MV is mapped out to provide additional information on
    the fidelity of using MV as a biodiesel surrogate." I am not convinced that MV could be a
    biodiesel surrogate, in my opinion it helps in understanding the chemistry of large esters.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that it is possible that MV will not be a good surrogate for
    biodiesel. Nonetheless, to our knowledge, there has been no study to suggest that it is not a
    good surrogate, probably because very few studies of MV have been conducted, so there is not
    enough information to judge. However, we think that judging the suitability (or lack thereof)
    for MV as a surrogate of biodiesel is outside the scope of this work, so we are simply noting
    that this study provides part of the necessary experimental information so that others may judge
    for themselves. We also agree with the reviewer that the scope of this work is actually wider
    than we had noted previously, so we have added a note that this study provides further
    information on the autoignition chemistry of large methyl esters near line 41:

    \begin{quote}
        %
        ``The NTC region of MV is mapped out to provide additional information on the fidelity of
        using MV as a biodiesel surrogate and on the autoignition chemistry of large methyl
        esters.''
        %
    \end{quote}
    %
\end{response}

\begin{reviewer}
    Specify the units of coefficients B, C  and, temperature T in equation (1)
\end{reviewer}

\begin{response}
    The coefficients are in \si{kPa} and \si{\kelvin}. This has been added to the text near line
    112 and in the caption of \cref{tab:antoine}.
    \begin{quote}
        ``... where \(A\), \(B\), and \(C\) are substance-specific coefficients, given in units of
        \si{\kelvin} and \si{\kPa}.''
    \end{quote}
\end{response}

\begin{reviewer}
    %
    The authors report an ignition delay of 100 ms at 640 K f=2, MV/air mixture at 15 bar. So at 30
    bar I would expect an ignition delay which is for sure will be more than 10 ms. So I am
    surprised by the statement "For the phi = 2:0 condition, only  PC = 15 bar is considered because
    we could not achieve TC values low enough that the ignition during the compression stroke can be
    prevented'. Can the authors comment on this.
    %
\end{reviewer}

\begin{response}
    %
    The reason for this is because of physical limitations of the experimental apparatus and the
    fuel. In particular, achieving sufficiently low temperatures on our apparatus requires
    increasing the reaction chamber volume at the end of compression, which in turn requires
    substantially higher initial pressures to reach the EOC pressure of \SI{30}{\bar}. The initial
    pressure in the RCM reaction chamber is limited by the pressure in the mixing tank, and the
    total pressure in the mixing tank is limited by the vapor pressure of MV at the preheat
    temperature. The combination of these factors meant that we could not prepare a mixture with
    sufficiently high tank pressure to enable repeated experiments with the EOC volume required to
    reach temperatures where the ignition delay would be measurable. However, we feel that this
    explanation would only confuse the manuscript, so we omit it.
    %
\end{response}

\begin{reviewer}
    %
    The comparison of ignition delays of MV from current work and that of Hadj-Ali et al .[9] would
    be helpful.
    %
\end{reviewer}

\begin{response}
    %
    The ignition delays of MV measured in the work of \citet{Hadj-Ali2009} were conducted at several
    pressures and roughly a single EOC temperature (near \SI{815}{\K}). However, none of the
    pressures considered in that study match the pressures considered in our study, so we feel that
    including the data on one of our existing plots would be very confusing. Moreover, the influence
    of facility-specific effects make a direct comparison of ignition delays from different RCMs
    somewhat challenging, particularly without an accurate chemical kinetic model to capture the
    facility effects. Therefore, we have not added a figure with this comparison.
    %
\end{response}

\begin{reviewer}
    %
    On discussion related to Dievart model in figure 5, it is surprising that phi=0.25 and phi=0.5
    simulations show NTC but phi=1 and phi=2 simulations do not. The trend is counter intuitive as
    the cool flame chemistry and NTC response are expected to increase with increase in equivalence
    ratio. This needs to be clarified. My guess is that the location of the NTC is not captured
    accurately by model. NTC region predicted by model is shifted to high temperatures where
    ignition is observed in RCM simulations. Showing Adiabatic constant volume simulations to
    understand the location of NTC at different equivalence ratios.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that the location of the NTC region with respect to inverse
    temperature is not well captured by the model. It does seem as though the NTC has been shifted
    to higher temperatures with respect to the data, in addition to the model being too reactive.
    This is supported by the newly added \cref{fig:conv-comparison}, which compares adiabatic,
    constant volume simulations with the RCM simulations from \cref{fig:simulation-comparison}. It
    is apparent from this figure that the predicted ignition delays in the $P_C = \SI{15}{\bar}$,
    $\phi = \num{2.0}$ and the $P_C = \SI{30}{\bar}$, $\phi = \num{1.0}$ simulations fall on the low
    temperature side of the NTC, while the $P_C = \SI{15}{\bar}$, $\phi =$ \numlist{0.25;0.50}
    simulated ignition delays are on the high-temperature side of the NTC, but approaching the NTC
    range such that they start to assume the characteristic curvature as the temperature decreases.
    In addition to the figure, we have added some text near line 290 to clarify this point.
    %
\end{response}

\begin{reviewer}
    %
    Figure 5 caption could also be revised.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the text of the caption of \cref{fig:simulation-comparison} to read:

    \begin{quote}
        %
        ``Comparison of experimental and simulated ignition delays computed using the procedure
        described in \cref{sec:experimental-modeling}. a) \SI{15}{\bar}, b) \SI{30}{\bar}.''
        %
    \end{quote}
    %
\end{response}

\begin{reviewer}
    %
    From Figure 5a for phi=2, 15 bar and  fig. 5b phi=1, 30 bar the  Dievart et al. model
    consistently under predicts the ignition delays which contradicts the statement "While the
    model of Dievart et al. [15] over-predicts the first-stage ignition delay, it also over-predicts
    the first-stage pressure rise, thereby leading  to rapid overall ignition." This needs to be
    revised.
    %
\end{reviewer}

\begin{response}
    %
    This statement has been removed from the manuscript.
    %
\end{response}

\begin{reviewer}
    %
    On discussion related to RMG model from Figure 5, the authors state " model over-predicts the
    low-temperature ignition delays and does not predict.." define low temperature ignition delays."
    The ignition delays from RMG model and experiments cross over around 720 K so the statement that
    low temp. ignition delays are overpredicted is not completely accurate.
    %
\end{reviewer}

\begin{response}
    %
    In this context, we are defining low temperature ignition delays as those to the right of the
    experimental NTC region on the Arrhenius plot. We have added the following text near line 308:

    \begin{quote}
        ``... the RMG model tends to over-predict the low-temperature ignition delays (i.e., those to the right of the experimental NTC region on the Arrhenius plot)...''
    \end{quote}
    %
\end{response}

\begin{reviewer}
    %
    Regarding the heat of formation shown in Table 3, it is interesting that the heat of formation
    of radical 3 and 4 estimated by RMG are different. I would request the authors check their
    numbers.
    %
\end{reviewer}

\begin{response}
    %
    We have checked the values for the estimates in the mechanism, and they are identical to the
    values in \cref{tab:heat-of-formation} and further identical to the values available from the
    RMG website at \url{http://rmg.mit.edu}. As such, no changes have been made to the manuscript in
    response to this comment.
    %
\end{response}

\begin{reviewer}
    %
    The authors use the branching ratios of fuel decomposition to fuel radicals to state the
    importance of fuel chemistry in the current work. The relative consumption shown in Table 4
    demonstrates that the branching ratios do not exhibit strong sensitive to thermochemistry which
    is supported by the fact that RMG and RMG switched show nearly identical branching ratios. I am
    not convinced that the importance of the thermochemistry is elucidated by the current discussion
    related to Table 4 and the discussion in my opinion does not provide any kinetic insights. May
    be the discussion could be removed.
    %
\end{reviewer}

\begin{response}
    %
    We have clarified the discussion of this point. In particular, we have emphasized that the
    results show that both the thermochemistry and reaction pathways are important, and cannot
    easily be disentangled from each other, requiring a thorough, detailed investigation of the
    methyl valerate reaction system which we consider outside the scope of the present work. In
    addition, we have noted that improper prediction of the thermochemistry of species may affect
    the RMG model in particular, due to the rate-based algorithm used in RMG, which may miss
    important reactions if the thermochemistry is not accurate. We have modified the text beginning
    near line 394, but do not copy the modification here for space.
    %
\end{response}

\begin{reviewer}
    %
    Alternatively, it would be interesting to compare the thermochemistry of the low temperature
    chemistry related reactions (RO2 chemistry) from models and ab-initio works like those Hayes and
    Burgess and also try adopting the reaction rates from ab-initio calculations and see to if the
    mechanism performance improves.
    %
\end{reviewer}

\begin{response}
    %
    We feel that the poor performance of both mechanisms indicates that a thorough and detailed
    study of the entire methyl valerate system is warranted, and small modifications, such as the
    replacement of only a few rates related to low-temperature chemistry, will not be sufficient to
    significantly improve the performance of the model. In addition, we feel that such ad-hoc
    replacement of rates without detailed study, even if it should improve the performance, could
    not be shown to be anything more than fortuitous without further study, which is outside the
    scope of the present work. Finally, in a separate work, Di\'{e}vart et al. (Combustion and
    Flame, 159 (5) 1793--1805, 2012) note that ``preliminary tests with up-to-date kinetic
    parameters [including the work of Hayes and Burgess] for the generic low temperature scheme have
    shown an enhancement of the fuel reactivity at the end of the NTC region.'' Since the model
    already over-predicts the reactivity of the RCM data, particularly at the end of the NTC region,
    further enhancement is not likely to improve agreement.

    Therefore, we have not included these modifications in the manuscript; we have, however, added a
    note to the end of the discussion that references the work of \citet{Hayes2009} to indicate that
    some of the necessary detailed work has been started.
    %
\end{response}

\noindent \textbf{Reviewer \#2:}

\begin{reviewer}
    %
    The measurement of Methyl Valerate at these temperature and pressure conditions constitutes
    important work and the manuscript of well laid out and has potential to be widely circulated.
    The arguments in the article are sound and the figures are clear.
    %
\end{reviewer}

\begin{response}
    Thank you for your comments. A detailed response is below.
\end{response}

\begin{reviewer}
    %
    I would like to ask the authors to check one item definitively before the article is published.
    I am concerned that at these higher pressures and equivalence ratios close to 1, there would be
    ignition which occurs during the compression stroke. This would not be noticed if some of the
    fuel condenses during the charging process and thereby you are dealing with lean mixtures than
    those cited in the paper. The authors are encouraged to make sure this is not the case by
    possibly sampling the charge after filling through a GC or by a quantitative absorption method
    as that used routinely by Stanford.
    %
\end{reviewer}

\begin{response}
    %
    With regards to ignition during the compression stroke, we have compared each reactive pressure
    trace to its corresponding non-reactive trace and found they are in good agreement until the end
    of compression, indicating that there is no heat release during the compression stroke. With
    regards to the filling process, we have thoroughly checked the temperatures throughout the
    apparatus and found them to be quite consistent, including the mixing tanks, the piping
    connections, and the reaction chamber itself. Moreover, we have conducted several studies in the
    past utilizing GC to determine mole fractions of components. All of these studies determined
    that the vaporization procedure used on this apparatus is sufficient to produce a homogeneous
    mixture of fuel and oxidizer, even for fuels with relatively low vapor pressures. We have added
    references to these prior studies in the present manuscript near line 105:

    \begin{quote}
        %
        ``Previous work has shown this procedure to completely vaporize the fuel and prevent fuel
        cracking during the heating process \cite{Weber2011a,Kumar2009,Das2012}.''
        %
    \end{quote}
    %
\end{response}

\begin{reviewer}
    %
    In the future, a possible direct charging method might be encouraged instead of a mix vessel.
    %
\end{reviewer}

\begin{response}
    %
    As we do not think it is appropriate to digress to specific discussion of various mixture
    preparation methods in this response or in the manuscript, we simply note in passing that each
    method has its own advantages and disadvantages, and each method should be used where its
    advantages provide the most benefit. We feel that such discussion in the manuscript would be
    beyond the scope of the current work, and we have not made any changes in response to this
    comment.
    %
\end{response}

\begin{reviewer}
    %
    It would also be good to include some of the pressure profiles so that others in the community
    can check.
    %
\end{reviewer}

\begin{response}
    %
    Pressure profiles are available in the data files that will be posted on the Combustion
    Diagnostics Laboratory website and on FigShare, with URLs given in the manuscript. We have
    endeavored to provide the community with all of the information necessary to reproduce these
    experiments, including the pressure traces and versions of software packages used in the
    processing. Moreover, we include several pressure traces with discussion in
    \cref{sec:pressure-traces} of the manuscript. As such, no changes have been made to the
    manuscript in response to this comment.
    %
\end{response}

\noindent\textbf{Reviewer \#3:}

\begin{reviewer}
    %
    The authors perform a shock tube study measuring ignition delay times of methyl pentanoate, and
    compare with a kinetic model taken from the literature and one they generated using a Reaction
    Mechanism Generator, showing that neither model can predict the ignition data satisfactorily;
    they investigate and discuss possible reasons, but (very reasonably) leave solving this
    discrepancy as an open challenge to the combustion modeling community.
    %
\end{reviewer}

\begin{response}
    Thank you for your comments. Detailed responses to each comment are presented below.
\end{response}

\begin{reviewer}
    %
    I would suggest using the preferred IUPAC name "Methyl pentanoate" in the title (and mentioning
    the common name "Methyl valerate" in the abstract and a keyword).
    %
\end{reviewer}

\begin{response}
    %
    We have modified the title to read ``Experiments and Modeling of the Autoignition of Methyl
    Pentanoate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression
    Machine'', as the reviewer suggests. Methyl valerate and methyl pentanoate are already keywords.
    We have opted to retain methyl valerate as the primary name used for the compound throughout the
    text because of the convenient MV acronym (MP is ambiguous, because it could mean methyl
    propanoate as well).
    %
\end{response}

\begin{reviewer}
    %
    Regarding: > "available on the web at https://combdiaglab.engr.uconn.edu/database/rcm-database
    and on figshare at https://doi.org/10.6084/m9.figshare.5213341. In addition, ChemKED-format [33]
    files are available in the main ChemKED database repository at
    https://github.com/pr-omethe-us/ChemKED-database."

    None of the links work or contain the relevant data. Presumably the authors embargoed it and
    plan to post it upon acceptance of the manuscript.

    But I am pleased to see this sharing of data, as well as the extensive supplementary materials
    aiding in reproducible science. Bravo.
    %
\end{reviewer}

\begin{response}
    %
    This is indeed the case, and the relevant files will be posted after acceptance. We have double
    checked that the URLs are correct, just in case. Thank you for your kind words.
    %
\end{response}

\begin{reviewer}
    %
    Regarding: > "At 15 bar, the experimental ignition delays are under-predicted by the Di\'{e}vart et
    al. [15] model for the three equivalence ratios shown. For the <phi> = 0.25 and 0.5 conditions,
    the model appears to be predicting an NTC region of the ignition delays as the temperature
    decreases"

    Figure 5 does not show, to my eye, any NTC behaviour in the simulations, as asserted in this
    sentence. Perhaps if the simulations were plotted at a wider range of temperatures (they're
    simulations, so no experimental reason not to?) this may show up, but at present I feel the
    figure does not support this statement. For <phi>=0.5 simulations are not even plotted over the
    whole range of the experimental data, let alone extrapolated to lower temperatures where NTC may
    be predicted.
    %
\end{reviewer}

\begin{response}
    %
    We interpret the increasing curvature of the model response for the $P_C = \SI{15}{\bar}$, $\phi =$
    \numlist{0.25;0.5} conditions as the beginning of an NTC region. The RCM simulations are not
    extrapolated because they use the volume profile associated with a single experiment. However,
    the simulations are plotted over the whole range of the $\phi = \num{0.5}$ experiments; the
    large discrepancy between the simulations and experiments causes the illusion that the whole
    range is not covered.

    In addition, we have added a new \cref{fig:conv-comparison} that compares adiabatic, constant
    volume simulations with the model of \citet{Dievart2013} with the RCM simulations presented in
    \cref{fig:simulation-comparison}. The adiabatic, constant volume simulations can be extrapolated
    over a wide range of temperatures. From \cref{fig:conv-comparison}, it can be seen that the
    curvature in the RCM simulations at $P_C = \SI{15}{\bar}$, $\phi =$ \numlist{0.25;0.5} is
    related to the NTC region of the ignition delay, while the lack of curvature in the $P_C =
    \SI{15}{\bar}$, $\phi = \num{2.0}$ and the $P_C = \SI{30}{\bar}$, $\phi = \num{1.0}$ simulations
    is because those lie on the low-temperature side of the predicted NTC.
    %
\end{response}

\begin{reviewer}
    %
    I have other questions about possible causes of the discrepancies, but answering them is beyond
    the reasonable scope of this paper, and the fact that I have them is a good sign regarding the
    potential impact of this paper. I think people will find it interesting.
    %
\end{reviewer}

\begin{response}
    Thank you again for your comments. We are also interested in the resolution of these questions,
    but we agree that they are out of scope for this work.
\end{response}

\end{document}

### 2017-09-22 revision-1/marked-revisions/methyl-valerate_old.tex
Revision 1 sent to Jackie

--- 2017-07-14 methyl-valerate.tex
@@
 \DeclareSIUnit\atm{atm}
+\DeclareSIUnit\calorie{cal}
 
@@
 \begin{frontmatter}
-\title{ Experiments and Modeling of Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }
+\title{ Experiments and Modeling of the Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }
 
@@
 UConnRCMPy~\cite{Weber2016a}. Version 3.0.5 of UConnRCMPy~\cite{uconnrcmpy}, 3.6 of Python, 2.3.0
-of Cantera~\cite{cantera}, 1.13 of NumPy~\cite{vanderWalt2011}, 0.19.0 of SciPy~\cite{Jones2001},
+of Cantera~\cite{cantera}, 1.13.0 of NumPy~\cite{vanderWalt2011}, 0.19.0 of SciPy~\cite{Jones2001},
 and 2.0.1 of Matplotlib~\cite{Hunter2007} are used in the analysis in this paper.
@@
 data; values for the coefficients computed in this work and in the literature works are given in
-\cref{tab:antoine}. The data used to calculate the coefficients are provided in the Supplementary
-Material.
+\cref{tab:antoine}. The data and code used to calculate the coefficients are provided in the
+Supplementary Material.
 
@@
 compression stroke and an isentropic expansion after the EOC. The volume trace is applied to a
-simulation conducted in an \verb|IdealGasReactor| in Cantera~\cite{cantera} using the CVODES
-solver from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delays from the simulations are
-defined in the same manner as in the experiments. The time derivative of the pressure in the
-simulation is computed by second order Lagrange polynomials, as discussed by \citet{Chapra2010}.
+simulation conducted in an \verb|IdealGasReactor| in Cantera~\cite{cantera} using the CVODES solver
+from the SUNDIALS suite~\cite{Hindmarsh2005}. The ignition delays from the simulations are defined
+in the same manner as in the experiments. The time derivative of the pressure in the simulation is
+computed by second order Lagrange polynomials, as discussed by \citet{Chapra2010}. The volume trace
+files, the corresponding pressure traces, and \texttt{volume-trace.yaml} files suitable for use with
+UConnRCMPy v3.0.5~\cite{uconnrcmpy} are available on the web at
+\url{https://combdiaglab.engr.uconn.edu/database/rcm-database} and on figshare at
+\url{https://doi.org/10.6084/m9.figshare.5213341}. In addition, ChemKED-format~\cite{Weber2017}
+files are available in the main ChemKED database repository at
+\url{https://github.com/pr-omethe-us/ChemKED-database}.
 
@@
 compressed pressure of \SI{30}{\bar}. Note that $\phi=2.0$ results were not collected for
-\SI{30}{\bar}, so there are no red triangle data points in \cref{fig:ignition-delays}b. All the data
-are available in comma-separated value files and ChemKED-format \cite{Weber2017} files in the
-Supplementary Material.
+\SI{30}{\bar}, so there are no red triangle data points in \cref{fig:ignition-delays}b. A summary of
+the ignition delay data is available as a comma-separated value file in the Supplementary Material.
 
@@
     \caption{Selected pressure traces around the NTC region of ignition delay.
-    a) \(\phi=1.0,\ P_C=\SI{30}{\bar}\) b) \(\phi=2.0,\ P_C=\SI{15}{\bar}\)}
+    a) \(\phi=1.0,\ P_C=\SI{30}{\bar}\), b) \(\phi=2.0,\ P_C=\SI{15}{\bar}\)}.
     \label{fig:pressure-traces}
@@
     \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
-    \caption{Comparison of experimental and simulated results. a) \SI{15}{\bar}; b) \SI{30}{\bar}}
+    \caption{Comparison of experimental and simulated results. a) \SI{15}{\bar}, b) \SI{30}{\bar}}.
     \label{fig:simulation-comparison}
@@
     \includegraphics[width=60mm]{figures/Methyl_pentanoate.png}
-    \caption{Structure of MV with carbon atoms labeled according to the convention used in \cref{tab:heat-of-formation} and \cref{tab:mv-radicals}}
+    \caption{Structure of MV with carbon atoms labeled according to the convention used in
+    \cref{tab:heat-of-formation} and \cref{tab:mv-radicals}}
     \label{fig:mv-structure}
@@
     \captionsetup{type=table}
-    \caption{Heats of formation of MV and its radicals, labeled according to the convention used in \cref{fig:mv-structure}}
+    \caption{Heats of formation of MV and its radicals, labeled according to the convention used in
+    \cref{fig:mv-structure}}
     \label{tab:heat-of-formation}
@@
 
+\section{Acknowledgments}\label{sec:acknowledgments}
+
+The authors acknowledge support from the Combustion Energy Frontier Research Center, an Energy
+Frontier Research Center funded by the U.S. Department of Energy, Office of Science, Office of Basic
+Energy Sciences, under award number DE-SC0001198.
+
 \section*{References}

### 2017-09-22 revision-1/revised-paper/methyl-valerate.tex
Revision 1 sent to Jackie

--- 2017-09-22 revision-1/marked-revisions/methyl-valerate_old.tex
@@
 \begin{frontmatter}
-\title{ Experiments and Modeling of the Autoignition of Methyl Valerate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }
+\title{ Experiments and Modeling of the Autoignition of Methyl Pentanoate at Low to Intermediate Temperatures and Elevated Pressures in a Rapid Compression Machine }
 
@@
 \numrange{0.25}{2.0}, and temperatures from \SIrange{680}{1050}{\K}. The NTC region of MV is mapped
-out to provide additional information on the fidelity of using MV as a biodiesel surrogate.
+out to provide additional information on the fidelity of using MV as a biodiesel surrogate and on
+the autoignition chemistry of large methyl esters.
 
@@
 \SI{1.5}{\hour} to come to steady state. The mixing tanks are also equipped with magnetic stir bars
-so the reactants are well mixed for the duration of the experiments.
+so the reactants are well mixed for the duration of the experiments. Previous work has shown this
+procedure to completely vaporize the fuel and prevent fuel cracking during the heating process
+\cite{Weber2011a,Kumar2009,Das2012}.
 
@@
 is used to model the saturated vapor pressure of MV as a function of temperature \((T)\), where
-\(A\), \(B\), and \(C\) are substance-specific coefficients. Coefficients for \cref{eq:antoine} are
-given in the literature by~\citet{Ortega2003}, \citet{Camacho2007}, and~\citet{Stephenson1987}.
-Unfortunately, the values of the coefficients are different among all three references. Therefore,
-coefficients for use in \cref{eq:antoine} are determined in this work by least squares fitting of
-the data of \citet{Ortega2003}, \citet{vanGenderen2002}, and \citet{Verevkin2008} using the
-\verb|curve_fit()| function of SciPy~\cite{Jones2001} version 0.19.0. \Cref{fig:vapor-pressure}
-shows that the coefficients fitted with this procedure give good agreement with the experimental
-data; values for the coefficients computed in this work and in the literature works are given in
-\cref{tab:antoine}. The data and code used to calculate the coefficients are provided in the
-Supplementary Material.
+\(A\), \(B\), and \(C\) are substance-specific coefficients, given in units of \si{\kelvin} and
+\si{\kPa}. Coefficients for \cref{eq:antoine} are given in the literature by~\citet{Ortega2003},
+\citet{Camacho2007}, and~\citet{Stephenson1987}. Unfortunately, the values of the coefficients are
+different among all three references. Therefore, coefficients for use in \cref{eq:antoine} are
+determined in this work by least squares fitting of the data of \citet{Ortega2003},
+\citet{vanGenderen2002}, and \citet{Verevkin2008} using the \verb|curve_fit()| function of
+SciPy~\cite{Jones2001} version 0.19.0. \Cref{fig:vapor-pressure} shows that the coefficients fitted
+with this procedure give good agreement with the experimental data; values for the coefficients
+computed in this work and in the literature works are given in \cref{tab:antoine}. The data and code
+used to calculate the coefficients are provided in the Supplementary Material.
 
@@
     \caption{Antoine Equation coefficients computed in this work and from the
-    literature. The $2\sigma$ confidence interval is estimated by taking the
-    square root of the diagonals of the covariance matrix returned from
+    literature, in units of \si{\kelvin} and \si{\kPa}. The $2\sigma$ confidence interval is
+    estimated by taking the square root of the diagonals of the covariance matrix returned from
     \texttt{curve\_fit()}}
@@
     \caption{Selected pressure traces around the NTC region of ignition delay.
-    a) \(\phi=1.0,\ P_C=\SI{30}{\bar}\), b) \(\phi=2.0,\ P_C=\SI{15}{\bar}\)}.
+    a) \(\phi=1.0,\ P_C=\SI{30}{\bar}\), b) \(\phi=2.0,\ P_C=\SI{15}{\bar}\).}
     \label{fig:pressure-traces}
@@
     \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
-    \caption{Comparison of experimental and simulated results. a) \SI{15}{\bar}, b) \SI{30}{\bar}}.
+    \caption{Comparison of experimental and simulated ignition delays computed using the procedure
+    described in \cref{sec:experimental-modeling}. a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
     \label{fig:simulation-comparison}
@@
 model for the three equivalence ratios shown. For the \(\phi =\) \numlist{0.25;0.5} conditions, the
-model appears to be predicting an NTC region of the ignition delays as the temperature decreases,
-although such a trend is not observed in the data. However, at \(\phi = 2.0\), the model does not
-predict the presence of an NTC region, although one is present in the experiments. Nonetheless, the
-agreement seems to be improving as the temperature is decreased. Comparing the \citet{Dievart2013}
-model to the stoichiometric data at \SI{30}{\bar}, we find a similar trend as the \(\phi=2.0,\
-P_C=\SI{15}{\bar}\) data. The model does not predict the NTC region found experimentally for the
-\(\phi = 1.0\) experiments, but the agreement improves as the temperature decreases. Interestingly,
-two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
-\cref{fig:simulation-comparison}. However, the first-stage ignition delays are
-\SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
-\cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
-delay. While the model of \citet{Dievart2013} over-predicts the first-stage ignition delay, it also
-over-predicts the first-stage pressure rise, thereby leading to rapid overall ignition.
+model appears to be predicting an NTC region of the ignition delays as the temperature decreases, as
+judged by the increasing curvature of the simulations, although such a trend is not observed in the
+data. However, at \(\phi = 2.0\), the model does not predict the presence of an NTC region, although
+one is present in the experiments. Nonetheless, the agreement seems to be improving as the
+temperature is decreased. Comparing the \citet{Dievart2013} model to the stoichiometric data at
+\SI{30}{\bar}, we find a similar trend as the \(\phi=2.0,\ P_C=\SI{15}{\bar}\) data. The model does
+not predict the NTC region found experimentally for the \(\phi = 1.0\) experiments, but the
+agreement improves as the temperature decreases. Interestingly, two-stage ignition is predicted for
+all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in \cref{fig:simulation-comparison}. However,
+the first-stage ignition delays are \SIrange{0.1}{0.5}{\ms} less than the overall ignition delays,
+and are not shown on \cref{fig:simulation-comparison} because they are nearly indistinguishable from
+the overall ignition delay.
+
+\begin{center}
+    \captionsetup{type=figure}
+    \includegraphics[width=\textwidth]{figures/constant-volume-comparison.pdf}
+    \caption{Comparison of simulated ignition delays computed in an adiabatic, constant volume
+    system (dotted lines) and computed using the procedure described in
+    \cref{sec:experimental-modeling} (solid lines). a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
+    \label{fig:conv-comparison}
+\end{center}
+
+To further understand the model of \citet{Dievart2013}, we have conducted adiabatic, constant volume
+simulations (called CONV simulations), as these simulations are not linked to a particular
+experiment by the volume trace and can be conducted over a wide range of temperatures.
+\Cref{fig:conv-comparison} compares the CONV simulations to the RCM simulations shown in
+\cref{fig:simulation-comparison}. From \cref{fig:conv-comparison}, it can be seen that the curvature
+in the RCM simulations at $P_C = \SI{15}{\bar}$, $\phi =$ \numlist{0.25;0.5} is related to the NTC
+region of the ignition delay, while the lack of curvature in the $P_C = \SI{15}{\bar}$, $\phi =
+\num{2.0}$ and the $P_C = \SI{30}{\bar}$, $\phi = \num{1.0}$ simulations is because those lie on the
+low-temperature side of the predicted NTC. It is clear from \cref{fig:conv-comparison} that the
+model of \citet{Dievart2013} predicts the NTC to occur at too high a temperature relative to the
+experiments.
 
@@
 RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar}, \(\phi =
-2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays, and does not
-predict the NTC region found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is
-similar to the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature
-ignition delays and does not predict the experimental NTC region. Finally, as in the
-\citet{Dievart2013} model, two-stage ignition is predicted for all of the \(\phi=1.0\) and
-\(\phi=2.0\) data shown in \cref{fig:simulation-comparison}. However, the first-stage ignition
-delays are \SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
+2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays (i.e., those to
+the right of the experimental NTC region on the Arrhenius plot), and does not predict the NTC region
+found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is similar to the
+\SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature ignition delays
+and does not predict the experimental NTC region. Finally, as in the \citet{Dievart2013} model,
+two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
+\cref{fig:simulation-comparison}. However, the first-stage ignition delays are
+\SIrange{0.1}{0.5}{\ms} less than the overall ignition delays, and are not shown on
 \cref{fig:simulation-comparison} because they are nearly indistinguishable from the overall ignition
-delay. Again, the over-prediction of the first-stage ignition delay and pressure rise requires
-further investigation and understanding of the complex low-temperature chemistry.
+delay.
 
@@
 \SI{20}{\percent} fuel consumption. The results of this analysis are shown in \cref{tab:mv-radicals}
-for the two models. The percentages shown in the \cref{tab:mv-radicals} are the percent of the fuel
+for the two models. The percentages shown in \cref{tab:mv-radicals} are the percent of the fuel
 consumed to form a particular fuel radical by all the reactions that can form that radical, and the
@@
 models have quite different distributions of products from the first \ce{H}-abstraction reactions.
-The model of \cite{Dievart2013} predicts that \ce{H}-abstraction from the second carbon is the most
+The model of \citet{Dievart2013} predicts that \ce{H}-abstraction from the second carbon is the most
 prevalent, while the RMG model predicts that the radical on the fourth carbon in the chain will be
@@
 condition as the other analysis is shown in \cref{tab:mv-radicals} in the ``RMG switched'' column.
-This analysis shows that the radical on the fourth carbon atom is still the most prevalant, despite
-changing the heats of formation of the fuel and its radicals.
+The results of the analysis of the ``RMG switched'' model show that the radical on the fourth carbon
+is still the most prevalent, despite the heats of formation for the radicals in the ``RMG switched''
+model being identical to the \citet{Dievart2013} model. This suggests that the reaction pathways
+have a substantial impact on the simulation, in addition to the influence of the thermochemistry, as
+discussed previously. Moreover, since the thermochemistry of the species in a reaction controls the
+reverse reaction rate of a reaction, the RMG algorithm may miss important pathways due to improperly
+estimated thermochemistry.
 
 Taken together, these results show that the poor performance in a given model cannot be attributed
-to a single source. There is a strong interaction between the thermodynamics of the species and the
-kinetics of the reactions, requiring further detailed study of the methyl ester system to accurately
-predict the low temperature ignition delays of methyl valerate.
+to a single source. Separating the influence of thermochemistry and kinetics requires further
+detailed study of the methyl valerate system specifically, and methyl ester systems more generally.
+Although such detailed work has begun, for example, with the work of \citet{Hayes2009}, further work
+is required to accurately predict the low temperature ignition delays of methyl valerate.
 

### 2017-09-22 revision-1/marked-revisions/figures/vapor_pressure/vapor_pressure_fit.py

@@ -1,0 +1,52 @@
+"""
+Calculate Antoine Equation Coefficients for the vapor pressure of
+methyl valerate.
+
+This code prints the best-fit coefficients of the Antoine Equation utilizing
+the data files available in this folder. The data have been collected from:
+
+Juan Ortega, Fernando Espiau, Jose Tojo, Jose Canosa, and Ana Rodriguez.
+"Isobaric Vapor-Liquid Equilibria and Excess Properties for the Binary
+Systems of Methyl Esters + Heptane." Journal of Chemical & Engineering
+Data 48, no. 5 (September 2003): 1183-1190. doi:10.1021/je030117d.
+
+Sergey P. Verevkin, and Vladimir N. Emel'yanenko. "Transpiration Method:
+Vapor Pressures and Enthalpies of Vaporization of Some Low-Boiling Esters."
+Fluid Phase Equilibria 266, no. 1-2 (April 2008): 64-75.
+doi:10.1016/j.fluid.2008.02.001.
+
+Aad C.G. van Genderen, J.Cees van Miltenburg, Jacobus G. Blok,
+Mark J. van Bommel, Paul J. van Ekeren, Gerrit J.K. van den Berg, and
+Harry A.J. Oonk. "Liquid-vapour Equilibria of the Methyl Esters of
+Alkanoic Acids: Vapour Pressures as a Function of Temperature and Standard
+Thermodynamic Function Changes." Fluid Phase Equilibria 202, no. 1
+(October 2002): 109-120. doi:10.1016/S0378-3812(02)00097-3.
+
+The data were transcribed from the original references by Bryan W. Weber.
+This file is distributed as part of the work "Experiments and Modeling of
+the Autoignition of Methyl Valerate at Low to Intermediate Temperatures
+and Elevated Pressures in a Rapid Compression Machine" by Bryan W. Weber,
+Justin A. Bunnell, Kamal Kumar, and Chih-Jen Sung.
+
+To the extent possible under law, the authors waive all copyright and related
+or neighboring rights to this file. This work is licensed under the CC0 public
+domain license, available at http://creativecommons.org/publicdomain/zero/1.0/
+"""
+import numpy as np
+from scipy.optimize import curve_fit
+
+
+def antoine(T, A, B, C):
+    return A - B/(T - C)
+
+
+ortega_data = np.loadtxt('ortega.csv', delimiter=',')
[…]

### 2017-09-25 revision-1/response/response.tex
Fix file modes

@@
     \begin{quote}
-        ``... the RMG model tends to over-predict the low-temperature ignition delays (i.e., those to the right of the experimental NTC region on the Arrhenius plot)...''
+        %
+        ``... the RMG model tends to over-predict the low-temperature overall ignition delays (i.e.,
+        those to the right of the experimental NTC region on the Arrhenius plot)...''
+        %
     \end{quote}
@@
 
+\newpage
+
 \noindent \textbf{Reviewer \#2:}
@@
     With regards to ignition during the compression stroke, we have compared each reactive pressure
-    trace to its corresponding non-reactive trace and found they are in good agreement until the end
-    of compression, indicating that there is no heat release during the compression stroke. With
-    regards to the filling process, we have thoroughly checked the temperatures throughout the
-    apparatus and found them to be quite consistent, including the mixing tanks, the piping
-    connections, and the reaction chamber itself. Moreover, we have conducted several studies in the
-    past utilizing GC to determine mole fractions of components. All of these studies determined
-    that the vaporization procedure used on this apparatus is sufficient to produce a homogeneous
-    mixture of fuel and oxidizer, even for fuels with relatively low vapor pressures. We have added
-    references to these prior studies in the present manuscript near line 105:
+    trace to its corresponding non-reactive trace and found they are in good agreement from the
+    beginning to the end of compression, indicating that there is no heat release during the
+    compression stroke. With regards to the filling process, we have thoroughly checked the
+    temperatures throughout the apparatus and found them to be quite consistent, including the
+    mixing tanks, the piping connections, and the reaction chamber itself. Moreover, we have
+    conducted several studies in the past utilizing GC to determine mole fractions of components.
+    All of these studies determined that the vaporization procedure used on this apparatus is
+    sufficient to produce a homogeneous mixture of fuel and oxidizer, even for fuels with relatively
+    low vapor pressures. We have added references to these prior studies in the present manuscript
+    near line 105:
 
@@
     %
-    As we do not think it is appropriate to digress to specific discussion of various mixture
-    preparation methods in this response or in the manuscript, we simply note in passing that each
-    method has its own advantages and disadvantages, and each method should be used where its
-    advantages provide the most benefit. We feel that such discussion in the manuscript would be
-    beyond the scope of the current work, and we have not made any changes in response to this
-    comment.
+    We first note that direct charging methods have their own set of disadvantages, and are still
+    required to address the issues of fuel condensation and fuel-oxidizer mixing. However, we feel
+    that adding a detailed discussion of the advantages and disadvantages of each mixing method to
+    the manuscript would be beyond the scope of the current work, and we have not made any changes
+    in response to this comment.
     %

### 2017-09-25 revision-1/response/response.tex
Start Jackie's revisions

@@
         ``The NTC region of MV is mapped out to provide additional information on the fidelity of
-        using MV as a biodiesel surrogate and on the autoignition chemistry of large methyl
-        esters.''
+        using MV as a biodiesel surrogate and insights into the autoignition chemistry of large
+        methyl esters.''
         %
@@
     of facility-specific effects make a direct comparison of ignition delays from different RCMs
-    somewhat challenging, particularly without an accurate chemical kinetic model to capture the
-    facility effects. Therefore, we have not added a figure with this comparison.
+    somewhat challenging, particularly without experimental pressure traces of MV from the work of
+    \citet{Hadj-Ali2009} to compare the facility effects. Therefore, we have not added a figure with
+    this comparison.
     %
@@
         %
-        ``Comparison of experimental and simulated ignition delays computed using the procedure
-        described in \cref{sec:experimental-modeling}. a) \SI{15}{\bar}, b) \SI{30}{\bar}.''
+        ``Comparison of experimental (\(\tau\) and \(\tau_1\)) and simulated (\(\tau\)) ignition delays
+        computed using the procedure described in \cref{sec:experimental-modeling}. a)
+        \SI{15}{\bar}, b) \SI{30}{\bar}.''
         %

### 2017-09-25 revision-1/revised-paper/methyl-valerate.tex
Start Jackie's revisions

@@
 \numrange{0.25}{2.0}, and temperatures from \SIrange{680}{1050}{\K}. The NTC region of MV is mapped
-out to provide additional information on the fidelity of using MV as a biodiesel surrogate and on
-the autoignition chemistry of large methyl esters.
+out to provide additional information on the fidelity of using MV as a biodiesel surrogate and
+insights into the autoignition chemistry of large methyl esters.
 
@@
     \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
-    \caption{Comparison of experimental and simulated ignition delays computed using the procedure
-    described in \cref{sec:experimental-modeling}. a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
+    \caption{Comparison of experimental (\(\tau\) and \(\tau_1\)) and simulated (\(\tau\)) ignition
+    delays computed using the procedure described in \cref{sec:experimental-modeling}. a)
+    \SI{15}{\bar}, b) \SI{30}{\bar}.}
     \label{fig:simulation-comparison}
@@
 RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar}, \(\phi =
-2.0\) data, the RMG model tends to over-predict the low-temperature ignition delays (i.e., those to
-the right of the experimental NTC region on the Arrhenius plot), and does not predict the NTC region
-found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is similar to the
-\SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature ignition delays
-and does not predict the experimental NTC region. Finally, as in the \citet{Dievart2013} model,
-two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
+2.0\) data, the RMG model tends to over-predict the low-temperature overall ignition delays (i.e.,
+those to the right of the experimental NTC region on the Arrhenius plot), and does not predict the
+NTC region found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is similar to
+the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature ignition
+delays and does not predict the experimental NTC region. Finally, as in the \citet{Dievart2013}
+model, two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
 \cref{fig:simulation-comparison}. However, the first-stage ignition delays are

### 2017-09-26 revision-1/revised-paper/methyl-valerate.tex
Finish Jackie's reviewer comments

@@
 \numrange{0.25}{2.0}, and temperatures from \SIrange{680}{1050}{\K}. The NTC region of MV is mapped
-out to provide additional information on the fidelity of using MV as a biodiesel surrogate and
-insights into the autoignition chemistry of large methyl esters.
+out to provide additional information on the fidelity of using MV as a biodiesel surrogate and on
+the autoignition chemistry of large methyl esters.
 
@@
     \resizebox{\textwidth}{!}{\input{figures/simulation-comparison.pgf}}
-    \caption{Comparison of experimental (\(\tau\) and \(\tau_1\)) and simulated (\(\tau\)) ignition
-    delays computed using the procedure described in \cref{sec:experimental-modeling}. a)
-    \SI{15}{\bar}, b) \SI{30}{\bar}.}
+    \caption{Comparison of experimental and simulated ignition delays computed using the procedure
+    described in \cref{sec:experimental-modeling}. a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
     \label{fig:simulation-comparison}

### 2017-09-26 revision-1/response/response.tex
Paper updates from Jackie

@@
         ``Previous work has shown this procedure to completely vaporize the fuel and prevent fuel
-        cracking during the heating process \cite{Weber2011a,Kumar2009,Das2012}.''
+        cracking during the heating process~\cite{Weber2011a,Kumar2009,Das2012}.''
         %

### 2017-09-26 revision-1/revised-paper/methyl-valerate.tex
Paper updates from Jackie

--- 2017-09-25 revision-1/revised-paper/methyl-valerate.tex
@@
 The prevalence of MB data in the literature is largely due to the early identification of MB as a
-potential surrogate fuel for biodiesel~\cite{Fisher2000}. However, the experiments have shown that
-MB may not be an appropriate surrogate for biodiesel, due to its lack of negative temperature
-coefficient (NTC) behavior, a requirement for a suitable biodiesel surrogate~\cite{Coniglio2013}.
+potential surrogate fuel for biodiesel~\cite{Fisher2000}. However, the literature experiments have
+shown that MB may not be an appropriate surrogate for biodiesel, due to its lack of negative
+temperature coefficient (NTC) behavior, a requirement for a suitable biodiesel
+surrogate~\cite{Coniglio2013}.
 
@@
     \includegraphics{figures/ignition_delay_definition.pdf}
-    \caption{Definition of the ignition delays used in this work. The
-    experiment in this figure is conducted for a \(\phi=2.0\) mixture with
-    \mbox{\(\ce{Ar}/(\ce{N2}+\ce{Ar})=0.5\)}, \(P_0=\SI{0.7806}{\bar}\),
-    \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.92}{\bar}\), \(T_C=\SI{720}{\K}\),
-    \(\tau=\SI{27.56\pm0.89}{\ms}\), and \(\tau_1=\SI{16.60\pm0.46}{\ms}\).}
+    \caption{Definition of the ignition delays used in this work. The experiment in this figure is
+    conducted for a \(\phi=2.0\) mixture with \mbox{\(\ce{Ar}/(\ce{N2}+\ce{Ar})=0.5\)},
+    \(P_0=\SI{0.7806}{\bar}\), \(T_0=\SI{373}{\K}\), \(P_C=\SI{14.92}{\bar}\), \(T_C=\SI{720}{\K}\),
+    \(\tau=\SI{27.56\pm0.89}{\ms}\), and \(\tau_1=\SI{16.60\pm0.46}{\ms}\). The non-reacting
+    pressure traces by replacing \ce{O2} with \ce{N2} is also shown for reference.}
     \label{fig:ign-delay-def}
@@
 with \ce{N2} to determine the influence of machine-specific behavior on the experimental conditions
-and permit the calculation of the EOC temperature via the isentropic relations between pressure and
-temperature~\cite{Lee1998}. The EOC temperature is calculated by the procedure described in
-\cref{sec:computational-methods}.
+(see \cref{fig:ign-delay-def}) and permit the calculation of the EOC temperature via the isentropic
+relations between pressure and temperature~\cite{Lee1998}. The EOC temperature is calculated by the
+procedure described in \cref{sec:computational-methods}.
 
@@
 proportions of reactants in the mixture are determined by specifying the absolute mass of the fuel,
-the equivalence ratio, and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer. Mixtures are
-made by first vacuuming the mixing tanks to an ultimate pressure less than \SI{5}{\torr}. Since MV
-is a liquid with a relatively small vapor pressure at room temperature and pressure, it is measured
+the equivalence ratio, and the ratio of \(\ce{Ar}:\ce{N2}\) in the oxidizer. Mixtures are made by
+first vacuuming the mixing tanks to an ultimate pressure less than \SI{5}{\torr}. Since MV is a
+liquid with a relatively small vapor pressure at room temperature and pressure, it is measured
 gravimetrically to within \SI{0.01}{\g} of the specified value. The fuel is injected into the mixing
@@
 temperature and the total pressure is measured by an Omega Engineering MMA100V10T2D0T4A6 type static
-pressure transducer. The same transducer is used to measure the pressure of the reactants prior to
-an experiment.
+pressure transducer. The same transducer is used to measure the pressure of the reactant mixture
+prior to an experiment.
 
@@
 so the reactants are well mixed for the duration of the experiments. Previous work has shown this
-procedure to completely vaporize the fuel and prevent fuel cracking during the heating process
-\cite{Weber2011a,Kumar2009,Das2012}.
-
-\begin{center}
-    \captionsetup{type=figure}
-    \input{figures/vapor_pressure.pgf}
-    \caption{Saturated vapor pressure of MV as a function of temperature,
-    plotted using the Antoine equation, \cref{eq:antoine}, with
-    \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
-    \label{fig:vapor-pressure}
-\end{center}
+procedure to completely vaporize the fuel and prevent fuel cracking during the heating
+process~\cite{Weber2011a,Kumar2009,Das2012}.
 
@@
 with this procedure give good agreement with the experimental data; values for the coefficients
-computed in this work and in the literature works are given in \cref{tab:antoine}. The data and code
-used to calculate the coefficients are provided in the Supplementary Material.
+computed in this work and reported in the literature works are given in \cref{tab:antoine}. The data
+and code used to calculate the coefficients are provided in the Supplementary Material.
+
+\begin{center}
+    \captionsetup{type=figure}
+    \input{figures/vapor_pressure.pgf}
+    \caption{Saturated vapor pressure of MV as a function of temperature,
+    plotted using the Antoine equation, \cref{eq:antoine}, with
+    \(A=6.4030\), \(B=1528.69\), and \(C=52.881\).}
+    \label{fig:vapor-pressure}
+\end{center}
 
@@
     \captionsetup{type=table}
-    \caption{Antoine Equation coefficients computed in this work and from the
-    literature, in units of \si{\kelvin} and \si{\kPa}. The $2\sigma$ confidence interval is
-    estimated by taking the square root of the diagonals of the covariance matrix returned from
-    \texttt{curve\_fit()}}
+    \caption{Antoine Equation coefficients computed in this work and obtained from the literature,
+    in units of \si{\kelvin} and \si{\kPa}. The $2\sigma$ confidence interval is estimated by taking
+    the square root of the diagonals of the covariance matrix returned from \texttt{curve\_fit()}}
     \label{tab:antoine}
@@
 simulate flames, and do not include the low-temperature chemistry necessary to simulate the
-conditions in these experiments. The third model was developed by \citet{Dievart2013} and includes
-low-temperature chemistry of MV, although it was only validated by comparison with flame extinction
-limits. In converting this mechanism for use in Cantera, we found that there were many species in
-the thermodynamic database with multiple data entries. For most of these species the thermodynamic
-data is identical. However, some species are not exact duplicates. For these species, it is not
-clear from the thermodynamic database file which data set should be preferred. Since Cantera (and
-CHEMKIN) choose the first instance of a duplicate species to be used, we retained the first entry
-for all duplicated species. The detailed model of \citet{Dievart2013} includes 1105 species and 7141
-reactions, and the Cantera formatted input file is available in the Supplementary Material.
+conditions in the current RCM experiments. The third model was developed by \citet{Dievart2013} and
+includes low-temperature chemistry of MV, although it was only validated by comparison with flame
+extinction limits. In converting this mechanism for use in Cantera, we found that there were many
+species in the thermodynamic database with multiple data entries. For most of these species the
+thermodynamic data is identical. However, some species are not exact duplicates. For these species,
+it is not clear from the thermodynamic database file which data set should be preferred. Since
+Cantera (and CHEMKIN) choose the first instance of a duplicate species to be used, we retained the
+first entry for all duplicated species. The detailed model of \citet{Dievart2013} includes 1105
+species and 7141 reactions, and the Cantera formatted input file is available in the Supplementary
+Material.
 
@@
 the model is available in the Supplementary Material. In addition, the CHEMKIN and Cantera formatted
-input files for the RMG model are available in the Supplementary Material.
+input files for the RMG-generated model are available in the Supplementary Material.
 
@@
 \SI{30}{\bar}, so there are no red triangle data points in \cref{fig:ignition-delays}b. A summary of
-the ignition delay data is available as a comma-separated value file in the Supplementary Material.
-
-It can be seen from \cref{fig:ignition-delays} that the ignition delays for the \(\phi=0.25\text{
-and } 0.5\) mixtures do not show an NTC region of the ignition delay for both of the pressures
-studied in this work. However, the \(\phi=1.0\) mixture shows an NTC region at
-\(P_C=\SI[number-unit-product={\ }]{30}{\bar}\) between approximately \SIlist{720;800}{\K}, with
-measured first-stage ignition delays at \SIlist{733;757}{\K}. In addition, the \(\phi=2.0\) mixture
-shows an NTC region of ignition delay at \SI{15}{\bar} from approximately \SIrange{720}{780}{\K},
-with measured first-stage ignition delays between \SIlist{720;750}{\K}.
+the ignition delay data is available as a comma-separated value file in the Supplementary Material,
+including the mixture conditions for each experiment, the initial conditions, the compressed
+conditions, and the ignition delays and their associated errors.
 
@@
 
+It can be seen from \cref{fig:ignition-delays} that the ignition delays for the \(\phi=0.25\text{
+and } 0.5\) mixtures do not show an NTC region of the ignition delay for both of the pressures
+studied in this work. However, the \(\phi=1.0\) mixture shows an NTC region at
+\(P_C=\SI[number-unit-product={\ }]{30}{\bar}\) between approximately \SIlist{720;800}{\K}, with
+measured first-stage ignition delays at \SIlist{733;757}{\K}. In addition, the \(\phi=2.0\) mixture
+shows an NTC region of ignition delay at \SI{15}{\bar} from approximately \SIrange{720}{780}{\K},
+with measured first-stage ignition delays between \SIlist{720;750}{\K}.
+
 \citet{Hadj-Ali2009} also observed two-stage ignition of MV in stoichiometric mixtures, stating that
@@
 this temperature, the autoignition occurred in two stages with a clearly identified cool flame
-event.'' However, we do not find two stage ignition for the similar pressure of
+event.'' However, we do not find two-stage ignition for the similar pressure of
 \(P_C = \SI{15}{\bar}\) in this study. We note that the stated temperature of the experiment from
@@
 
-However, we note NTC behavior of the ignition delay and two-stage ignition at the higher pressure of
-\SI{30}{\bar}, and at higher temperatures than those reported for two-stage ignition in the study of
-\citet{Hadj-Ali2009}. The trend of NTC behavior shifting to higher temperatures with increasing
-pressure can be seen in other classes of fuels. \citet{Kukkadapu2012} found a similar trend in
-gasoline composed of iso-alkanes, n-alkanes, cyclo-alkanes, aromatics, and olefins.
+However, we find NTC behavior of the overall ignition delay and two-stage ignition at the higher
+pressure of \SI{30}{\bar}, and at higher temperatures than those reported for two-stage ignition in
+the study of \citet{Hadj-Ali2009}. The trend of NTC behavior shifting to higher temperatures with
+increasing pressure can be seen in other classes of fuels. \citet{Kukkadapu2012} found a similar
+trend in gasoline composed of iso-alkanes, n-alkanes, cyclo-alkanes, aromatics, and olefins.
 \citet{Kukkadapu2012} attributed the shift of the NTC region to the reactions between the
@@
 with our MV data at \SI{30}{\bar}, we note that increasing the pressure tends to shift the NTC to
-higher temperatures, as mentioned previously \cite{Kukkadapu2012}. \citet{Heufer2013} found NTC
+higher temperatures, as mentioned previously~\cite{Kukkadapu2012}. \citet{Heufer2013} found NTC
 behavior for n-pentanol in the range of \SIrange{770}{900}{\K} at \SI{30}{\bar}. In this study, we
@@
     \includegraphics[width=\textwidth]{figures/pressure-traces.pdf}
-    \caption{Selected pressure traces around the NTC region of ignition delay.
-    a) \(\phi=1.0,\ P_C=\SI{30}{\bar}\), b) \(\phi=2.0,\ P_C=\SI{15}{\bar}\).}
+    \caption{Selected pressure traces around the NTC region of ignition delay.  a) \(\phi=1.0\),
+    \(P_C=\SI{30}{\bar}\), b) \(\phi=2.0\), \(P_C=\SI{15}{\bar}\). The corresponding non-reactive
+    pressure traces are also included for reference.}
     \label{fig:pressure-traces}
@@
 
-At \SI{15}{\bar}, the experimental ignition delays are under-predicted by the \citet{Dievart2013}
-model for the three equivalence ratios shown. For the \(\phi =\) \numlist{0.25;0.5} conditions, the
-model appears to be predicting an NTC region of the ignition delays as the temperature decreases, as
-judged by the increasing curvature of the simulations, although such a trend is not observed in the
-data. However, at \(\phi = 2.0\), the model does not predict the presence of an NTC region, although
-one is present in the experiments. Nonetheless, the agreement seems to be improving as the
-temperature is decreased. Comparing the \citet{Dievart2013} model to the stoichiometric data at
-\SI{30}{\bar}, we find a similar trend as the \(\phi=2.0,\ P_C=\SI{15}{\bar}\) data. The model does
-not predict the NTC region found experimentally for the \(\phi = 1.0\) experiments, but the
+At \SI{15}{\bar}, the experimental overall ignition delays are under-predicted by the
+\citet{Dievart2013} model for the three equivalence ratios shown. For the \(\phi =\)
+\numlist{0.25;0.5} conditions, the model appears to be predicting an NTC region of the overall
+ignition delays as the temperature decreases, as judged by the increasing curvature of the
+simulations, although such a trend is not observed apparent in the figure. However, at \(\phi =
+2.0\), the model does not predict the presence of an NTC region, although one is present in the
+experiments. Nonetheless, the agreement seems to be improving as the temperature is decreased.
+Comparing the \citet{Dievart2013} model to the stoichiometric data at \SI{30}{\bar}, we find a
+similar trend as the \(\phi=2.0\), \(P_C=\SI{15}{\bar}\) data. The model does not predict the NTC
+region found experimentally for the \(\phi = 1.0\), \(P_C=\SI{30}{\bar}\) experiments, but the
 agreement improves as the temperature decreases. Interestingly, two-stage ignition is predicted for
@@
 simulations (called CONV simulations), as these simulations are not linked to a particular
-experiment by the volume trace and can be conducted over a wide range of temperatures.
-\Cref{fig:conv-comparison} compares the CONV simulations to the RCM simulations shown in
-\cref{fig:simulation-comparison}. From \cref{fig:conv-comparison}, it can be seen that the curvature
-in the RCM simulations at $P_C = \SI{15}{\bar}$, $\phi =$ \numlist{0.25;0.5} is related to the NTC
-region of the ignition delay, while the lack of curvature in the $P_C = \SI{15}{\bar}$, $\phi =
-\num{2.0}$ and the $P_C = \SI{30}{\bar}$, $\phi = \num{1.0}$ simulations is because those lie on the
-low-temperature side of the predicted NTC. It is clear from \cref{fig:conv-comparison} that the
-model of \citet{Dievart2013} predicts the NTC to occur at too high a temperature relative to the
-experiments.
+experiment by the volume trace and can be conducted over a wide range of temperatures. In the CONV
+simulations, overall ignition delay is defined as an increase in the temperature of \SI{400}{\K}
+over the initial temperature. \Cref{fig:conv-comparison} compares the CONV simulations to the RCM
+simulations shown in \cref{fig:simulation-comparison}. From \cref{fig:conv-comparison}, it can be
+seen that the curvature in the RCM simulations at \(P_C = \SI{15}{\bar}\), \(\phi =\)
+\numlist{0.25;0.5} is related to the NTC region of the overall ignition delay, while the lack of
+curvature in the \(P_C = \SI{15}{\bar}\), \(\phi = \num{2.0}\) and the \(P_C = \SI{30}{\bar}\),
+\(\phi = \num{1.0}\) simulations is because those lie on the low-temperature side of the predicted
+NTC. It is clear from \cref{fig:conv-comparison} that the model of \citet{Dievart2013} predicts the
+NTC to occur at too high a temperature relative to the experiments.
 
 To elucidate the underlying reasons for the disagreement between the \citet{Dievart2013} model and
-the data, we constructed an additional model using RMG (see
+the data, we conduct simulations with an additional model constructed using RMG (see
 \cref{sec:reaction-mechanism-generator}). As can be seen in \cref{fig:simulation-comparison}a, the
 agreement between the RMG model (dashed lines) and the experimental data is similar to the
-\citet{Dievart2013} model for the \SI{15}{\bar}, \(\phi =\) \numlist{0.25; 0.5} data. Moreover, the
-RMG model predicts a similar NTC region as temperature is decreasing. For the \SI{15}{\bar}, \(\phi =
-2.0\) data, the RMG model tends to over-predict the low-temperature overall ignition delays (i.e.,
-those to the right of the experimental NTC region on the Arrhenius plot), and does not predict the
-NTC region found experimentally. As before, the trend at \SI{30}{\bar}, \(\phi=1.0\) is similar to
-the \SI{15}{\bar}, \(\phi=2.0\) data; the RMG model over-predicts the low-temperature ignition
-delays and does not predict the experimental NTC region. Finally, as in the \citet{Dievart2013}
-model, two-stage ignition is predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
+\citet{Dievart2013} model for the \(P_C = \SI{15}{\bar}\), \(\phi =\) \numlist{0.25; 0.5} data.
+Moreover, the RMG model predicts a similar NTC region as temperature is decreasing. For the \(P_C =
+\SI{15}{\bar}\), \(\phi = 2.0\) data, the RMG model tends to over-predict the low-temperature
+overall ignition delays (i.e., those to the right of the experimental NTC region on the Arrhenius
+plot), and does not predict the NTC region found experimentally. As before, the trend at \(P_C =
+\SI{30}{\bar}\), \(\phi=1.0\) is similar to the \(P_C = \SI{15}{\bar}\), \(\phi=2.0\) data; the RMG
+model over-predicts the low-temperature overall ignition delays and does not predict the
+experimental NTC region. Finally, as in the \citet{Dievart2013} model, two-stage ignition is
+predicted for all of the \(\phi=1.0\) and \(\phi=2.0\) data shown in
 \cref{fig:simulation-comparison}. However, the first-stage ignition delays are
@@
 showed substantial heat release during the compression stroke (i.e., the simulations are much too
-reactive), and so these conditions aren't compared in \cref{fig:simulation-comparison}. We note
+reactive), and so these conditions are not compared in \cref{fig:simulation-comparison}. We note
 again that the model by \citet{Dievart2013} was validated for MV combustion only by comparison to
@@
 The third source of error may lie in the estimation of the thermodynamic properties of the species,
-particularly the fuel radicals. In the work of \citet{Dievart2013}, the program THERM
-\cite{Ritter1991} is used to estimate thermodynamic values using the group additivity method. In the
-RMG model constructed in this work, RMG itself estimates the thermodynamic properties of the
+particularly the fuel radicals. In the work of \citet{Dievart2013}, the program
+THERM~\cite{Ritter1991} was used to estimate thermodynamic values using the group additivity method.
+In the RMG model constructed in this work, RMG itself estimates the thermodynamic properties of the
 molecules also using the group additivity method. Nonetheless, the two models have differing
@@
 process. The following analysis is conducted for a constant volume, adiabatic simulation with
-initial temperature and pressure of \SI{700}{\K}, \SI{30}{\bar}, and for the stoichiometric
-equivalence ratio. The rates of production of the species have been integrated until the time of
-\SI{20}{\percent} fuel consumption. The results of this analysis are shown in \cref{tab:mv-radicals}
-for the two models. The percentages shown in \cref{tab:mv-radicals} are the percent of the fuel
-consumed to form a particular fuel radical by all the reactions that can form that radical, and the
-radicals are labeled according to the convention in \cref{fig:mv-structure}.
+initial temperature and pressure of \SI{700}{\K} and \SI{30}{\bar}, respectively, and for the
+stoichiometric equivalence ratio. The rates of production of the species have been integrated until
+the time of \SI{20}{\percent} fuel consumption. The results of this analysis are shown in
+\cref{tab:mv-radicals} for the two models. The percentages shown in \cref{tab:mv-radicals} are the
+percent of the fuel consumed to form a particular fuel radical by all the reactions that can form
+that radical, and the radicals are labeled according to the convention in \cref{fig:mv-structure}.
 
@@
 To further compare the models with each other, the NASA polynomials representing the thermodynamic
-properties of MV and the 5 fuel radicals from the model of \citet{Dievart2013} and are used to
-replace the equivalent molecules in the RMG model. The results of a path analysis at the same
-condition as the other analysis is shown in \cref{tab:mv-radicals} in the ``RMG switched'' column.
-The results of the analysis of the ``RMG switched'' model show that the radical on the fourth carbon
-is still the most prevalent, despite the heats of formation for the radicals in the ``RMG switched''
-model being identical to the \citet{Dievart2013} model. This suggests that the reaction pathways
-have a substantial impact on the simulation, in addition to the influence of the thermochemistry, as
+properties of MV and the 5 fuel radicals from the model of \citet{Dievart2013} are used to replace
+the equivalent molecules in the RMG model. The results of a path analysis at the same condition as
+the other analysis is shown in \cref{tab:mv-radicals} in the ``RMG switched'' column. The results of
+the analysis of the ``RMG switched'' model show that the radical on the fourth carbon is still the
+most prevalent, despite the heats of formation for the fuel radicals in the ``RMG switched'' model
+being identical to the \citet{Dievart2013} model. This suggests that the reaction pathways have a
+substantial impact on the simulation, in addition to the influence of the thermochemistry, as
 discussed previously. Moreover, since the thermochemistry of the species in a reaction controls the
@@
 In this study, we have measured ignition delays for methyl valerate over a wide range of
-engine-relevant pressures, temperatures, and equivalence ratios. An NTC region of the ignition delay
-and two-stage ignition are recorded for pressures of \SI{15}{\bar} at \(\phi=2.0\) and \SI{30}{\bar}
-at \(\phi=1.0\). A detailed chemical kinetic model available in the literature is unable to
-reproduce the experimental results, so a new model is constructed using the Reaction Mechanism
-Generator software. Although the new model contains many more reactions than the literature model,
-it is still unable to predict the experimental ignition delays satisfactorily. Both models predict
-an NTC region of the ignition delay under conditions where none is found in the experiments, and
-fail to predict the NTC region of ignition delay that is present in the experiments. Possible
-reasons for the discrepancy include missing reaction pathways, incorrect rate estimates, and
-incorrect thermodynamic property estimates. Comparative analysis of the two models failed to
-identify a single source of the error, and further detailed studies are required to improve
-predictions of the ignition delay at these engine-relevant conditions.
+engine-relevant pressures, temperatures, and equivalence ratios. An NTC region of the overall
+ignition delay and two-stage ignition are recorded for pressures of \SI{15}{\bar} at \(\phi=2.0\)
+and \SI{30}{\bar} at \(\phi=1.0\). A detailed chemical kinetic model available in the literature is
+unable to reproduce the experimental results, so a new model is constructed using the Reaction
+Mechanism Generator software. Although the new model contains many more reactions than the
+literature model, it is still unable to predict the experimental ignition delays satisfactorily.
+Both models predict an NTC region of the overall ignition delay under conditions where none is found
+in the experiments, and fail to predict the NTC region of overall ignition delay that is present in
+the experiments. Possible reasons for the discrepancy include missing reaction pathways, incorrect
+rate estimates, and incorrect thermodynamic property estimates. Comparative analysis of the two
+models failed to identify a single source of the error, and further detailed studies are required to
+improve predictions of the ignition delay at these engine-relevant conditions.
 

### 2017-09-27 revision-1/response/response.tex
Final figure updates and response checks

@@
     biodiesel. Nonetheless, to our knowledge, there has been no study to suggest that it is not a
-    good surrogate, probably because very few studies of MV have been conducted, so there is not
-    enough information to judge. However, we think that judging the suitability (or lack thereof)
-    for MV as a surrogate of biodiesel is outside the scope of this work, so we are simply noting
-    that this study provides part of the necessary experimental information so that others may judge
-    for themselves. We also agree with the reviewer that the scope of this work is actually wider
-    than we had noted previously, so we have added a note that this study provides further
-    information on the autoignition chemistry of large methyl esters near line 41:
+    good surrogate, probably because very few studies of MV have been conducted to date, so there is
+    not enough information to judge. However, we think that judging the suitability (or lack
+    thereof) for MV as a surrogate of biodiesel is outside the scope of this work, so we are simply
+    noting that this study provides part of the necessary experimental information so that others
+    may judge for themselves. We also agree with the reviewer that the scope of this work is
+    actually wider than we had noted previously, so we have added a note that this study provides
+    further information on the autoignition chemistry of large methyl esters near line 41:
 
@@
     The reason for this is because of physical limitations of the experimental apparatus and the
-    fuel. In particular, achieving sufficiently low temperatures on our apparatus requires
+    fuel. In particular, achieving sufficiently EOC low temperatures on our apparatus would require
     increasing the reaction chamber volume at the end of compression, which in turn requires
@@
     sufficiently high tank pressure to enable repeated experiments with the EOC volume required to
-    reach temperatures where the ignition delay would be measurable. However, we feel that this
+    reach EOC temperatures where the ignition delay would be measurable. However, we feel that this
     explanation would only confuse the manuscript, so we omit it.
@@
     range such that they start to assume the characteristic curvature as the temperature decreases.
-    In addition to the figure, we have added some text near line 290 to clarify this point.
+    In addition to the figure, we have added some paragraphs near line 293 to clarify this point,
+    but we have not copied the added text here to save space.
     %
@@
     In this context, we are defining low temperature ignition delays as those to the right of the
-    experimental NTC region on the Arrhenius plot. We have added the following text near line 308:
+    experimental NTC region on the Arrhenius plot. We have added the following text near line 319:
 
@@
     addition, we have noted that improper prediction of the thermochemistry of species may affect
-    the RMG model in particular, due to the rate-based algorithm used in RMG, which may miss
-    important reactions if the thermochemistry is not accurate. We have modified the text beginning
-    near line 394, but do not copy the modification here for space.
+    the RMG-generated model in particular due to the rate-based algorithm used in RMG, which may
+    miss important reactions if the thermochemistry is not accurate. We have modified the text
+    beginning near line 402, but do not copy the modification here for space.
     %
@@
     replacement of only a few rates related to low-temperature chemistry, will not be sufficient to
-    significantly improve the performance of the model. In addition, we feel that such ad-hoc
-    replacement of rates without detailed study, even if it should improve the performance, could
-    not be shown to be anything more than fortuitous without further study, which is outside the
-    scope of the present work. Finally, in a separate work, Di\'{e}vart et al. (Combustion and
-    Flame, 159 (5) 1793--1805, 2012) note that ``preliminary tests with up-to-date kinetic
-    parameters [including the work of Hayes and Burgess] for the generic low temperature scheme have
-    shown an enhancement of the fuel reactivity at the end of the NTC region.'' Since the model
-    already over-predicts the reactivity of the RCM data, particularly at the end of the NTC region,
-    further enhancement is not likely to improve agreement.
+    significantly improve the performance of the model. Moreover, we feel that such ad-hoc
+    replacement of rates, even if the performance improves, could not be shown to be anything more
+    than fortuitous without further detailed study, which is outside the scope of the present work.
+    Finally, in a separate work, Di\'{e}vart et al. (Combustion and Flame, 159 (5) 1793--1805, 2012)
+    note that ``preliminary tests with up-to-date kinetic parameters [including the work of Hayes
+    and Burgess] for the generic low temperature scheme have shown an enhancement of the fuel
+    reactivity at the end of the NTC region.'' Since the model already over-predicts the reactivity
+    of the RCM data, particularly at the end of the NTC region, further enhancement is not likely to
+    improve agreement.
 
     Therefore, we have not included these modifications in the manuscript; we have, however, added a
-    note to the end of the discussion that references the work of \citet{Hayes2009} to indicate that
-    some of the necessary detailed work has been started.
+    note to the end of the discussion (near line 418) that references the work of \citet{Hayes2009}
+    to indicate that some of the necessary detailed work has been started.
     %
 \end{response}
-
-\newpage
 
@@
     low vapor pressures. We have added references to these prior studies in the present manuscript
-    near line 105:
+    near line 110:
 

### 2017-09-27 revision-1/revised-paper/methyl-valerate.tex
Final figure updates and response checks

@@
 
+To further understand the model of \citet{Dievart2013}, we have conducted adiabatic, constant volume
+simulations (called CONV simulations), as these simulations are not linked to a particular
+experiment by the volume trace and can be conducted over a wide range of temperatures. In the CONV
+simulations, overall ignition delay is defined as an increase in the temperature of \SI{400}{\K}
+over the initial temperature. The diluent for all of the CONV simulations is pure argon, although
+the RCM simulations described previously consider the diluent mixture associated with each
+experiment (either pure argon or a 50:50 mixture of argon and nitrogen by mole, see
+\cref{tab:mixtures} and the Supplementary Material).
+
 \begin{center}
@@
     \includegraphics[width=\textwidth]{figures/constant-volume-comparison.pdf}
-    \caption{Comparison of simulated ignition delays computed in an adiabatic, constant volume
-    system (dotted lines) and computed using the procedure described in
+    \caption{Comparison of simulated overall ignition delays computed in an adiabatic, constant
+    volume system (dotted lines) and computed using the procedure described in
     \cref{sec:experimental-modeling} (solid lines). a) \SI{15}{\bar}, b) \SI{30}{\bar}.}
@@
 
-To further understand the model of \citet{Dievart2013}, we have conducted adiabatic, constant volume
-simulations (called CONV simulations), as these simulations are not linked to a particular
-experiment by the volume trace and can be conducted over a wide range of temperatures. In the CONV
-simulations, overall ignition delay is defined as an increase in the temperature of \SI{400}{\K}
-over the initial temperature. \Cref{fig:conv-comparison} compares the CONV simulations to the RCM
-simulations shown in \cref{fig:simulation-comparison}. From \cref{fig:conv-comparison}, it can be
-seen that the curvature in the RCM simulations at \(P_C = \SI{15}{\bar}\), \(\phi =\)
-\numlist{0.25;0.5} is related to the NTC region of the overall ignition delay, while the lack of
-curvature in the \(P_C = \SI{15}{\bar}\), \(\phi = \num{2.0}\) and the \(P_C = \SI{30}{\bar}\),
-\(\phi = \num{1.0}\) simulations is because those lie on the low-temperature side of the predicted
-NTC. It is clear from \cref{fig:conv-comparison} that the model of \citet{Dievart2013} predicts the
-NTC to occur at too high a temperature relative to the experiments.
+\Cref{fig:conv-comparison} compares the CONV simulations to the RCM simulations shown in
+\cref{fig:simulation-comparison}. From \cref{fig:conv-comparison}, it can be seen that the curvature
+in the RCM simulations at \(P_C = \SI{15}{\bar}\), \(\phi =\) \numlist{0.25;0.5} is related to the
+NTC region of the overall ignition delay, while the lack of curvature in the \(P_C =
+\SI{15}{\bar}\), \(\phi = \num{2.0}\) and the \(P_C = \SI{30}{\bar}\), \(\phi = \num{1.0}\)
+simulations is because those lie on the low-temperature side of the predicted NTC. It is clear from
+\cref{fig:conv-comparison} that the model of \citet{Dievart2013} predicts the NTC to occur at too
+high a temperature relative to the experiments.
 
@@
 
+\Cref{tab:heat-of-formation} shows that the heats of formation of the fuel and radicals 3, 4, 5, and
+M are quite similar between the two mechanisms. However, the heat of formation of the second
+radical, the one closest to the methyl ester group, has a significantly lower heat of formation in
+the model by \citet{Dievart2013} than in the RMG model. Note that it is expected that the second
+radical will be somewhat more stable than the other radicals, due to the influence of the methyl
+ester group on the adjacent carbon atom.
+
 \begin{center}
@@
 
-\Cref{tab:heat-of-formation} shows that the heats of formation of the fuel and radicals 3, 4, 5, and
-M are quite similar between the two mechanisms. However, the heat of formation of the second
-radical, the one closest to the methyl ester group, has a significantly lower heat of formation in
-the model by \citet{Dievart2013} than in the RMG model. Note that it is expected that the second
-radical will be somewhat more stable than the other radicals, due to the influence of the methyl
-ester group on the adjacent carbon atom.
-
 This difference in heats of formation affects the pathways that consume the fuel. By conducting a
@@
 that radical, and the radicals are labeled according to the convention in \cref{fig:mv-structure}.
+
+At the relatively low temperature and high pressure condition of this analysis, all of the fuel is
+consumed by \ce{H}-atom abstractions to form the fuel radicals shown. It can be seen that the two
+models have quite different distributions of products from the first \ce{H}-abstraction reactions.
+The model of \citet{Dievart2013} predicts that \ce{H}-abstraction from the second carbon is the most
+prevalent, while the RMG model predicts that the radical on the fourth carbon in the chain will be
+primarily formed. This is in line with the heats of formation in \cref{tab:heat-of-formation}, where
+the most stable radical (i.e., the radical with the smallest heat of formation) is most likely to be
+formed in each model.
 
@@
 \end{center}
-
-At the relatively low temperature and high pressure condition of this analysis, all of the fuel is
-consumed by \ce{H}-atom abstractions to form the fuel radicals shown. It can be seen that the two
-models have quite different distributions of products from the first \ce{H}-abstraction reactions.
-The model of \citet{Dievart2013} predicts that \ce{H}-abstraction from the second carbon is the most
-prevalent, while the RMG model predicts that the radical on the fourth carbon in the chain will be
-primarily formed. This is in line with the heats of formation in \cref{tab:heat-of-formation}, where
-the most stable radical (i.e., the radical with the smallest heat of formation) is most likely to be
-formed in each model.
 

### 2017-09-27 revision-1/revised-paper/methyl-valerate.tex
Final updates of the revision!

@@
     \(\tau=\SI{27.56\pm0.89}{\ms}\), and \(\tau_1=\SI{16.60\pm0.46}{\ms}\). The non-reacting
-    pressure traces by replacing \ce{O2} with \ce{N2} is also shown for reference.}
+    pressure trace by replacing \ce{O2} with \ce{N2} is also shown for reference.}
     \label{fig:ign-delay-def}
@@
 ignition delays as the temperature decreases, as judged by the increasing curvature of the
-simulations, although such a trend is not observed apparent in the figure. However, at \(\phi =
+simulations, although such a trend is not observed for the experimental data. However, at \(\phi =
 2.0\), the model does not predict the presence of an NTC region, although one is present in the
